---
title: "CellChat.byOrigin2.LN_P"
---

```{r knitr, include = FALSE}
DOCNAME = "CellChat.byOrigin2"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(ggplotify)
library(ggcorrplot)
library(ggpubr)

# patch
library(patchwork)
library(cowplot)
theme_set(theme_cowplot(font_size = 12,
                        rel_small = 10/12,
                        rel_tiny = 8/12,
                        rel_large = 12/12,
                        font_family = "Arial"))
# heatmap
library(pheatmap)
library(ComplexHeatmap)

# fonts
library(extrafont)
#font_import()
#font_import(paths = "/T02Data/niuyw/.local/share/fonts")
loadfonts(device = "pdf", quiet = TRUE)

# color
library(ggsci)

# cellchat
library(CellChat)
library(NMF)
library(ggalluvial)

# Seurat
library(Seurat)

options(stringsAsFactors = FALSE)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load rds

```{r load-rds-srat, cache=FALSE, eval=FALSE}
seurat = readRDS(here::here('output/04.rm_cells/seurat.rds'))
seurat
```

```{r load-rds, cache=FALSE}
cellchat = readRDS(here::here("output", DOCNAME, 'cellchat.LN_P.rds'))
cellchat
```


## Create a CellChat object

```{r create-cellchat, eval=FALSE}
srat_sub = subset(seurat, Origin2_n == "LN_P")
data.input <- GetAssayData(srat_sub, assay = "RNA", slot = "data") # normalized data matrix
labels <- srat_sub$cellType2
meta <- data.frame(group = labels, row.names = names(labels)) # create a dataframe of the cell labels

# drop unused levels, otherwise it would throw an error
meta$group = droplevels(meta$group, exclude = setdiff(levels(meta$group), unique(meta$group)))

# create
cellchat <- createCellChat(object = data.input, meta = meta, group.by = "group")
```

Set the ligand-receptor interaction database

```{r set-database, fig.width=8, fig.height=4}
CellChatDB <- CellChatDB.human # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)

# set the used database in the object
cellchat@DB <- CellChatDB
```

Preprocessing the expression data

```{r Preprocessing, eval=FALSE}
# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database

cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
# project gene expression data onto PPI network (optional)
cellchat <- projectData(cellchat, PPI.human)
```

## Inference of cell-cell communication network

Compute the communication probability and infer cellular communication network

```{r computeCommunProb, eval=FALSE}
cellchat <- computeCommunProb(cellchat)
# Filter out the cell-cell communication if there are only few number of cells in certain cell groups
cellchat <- filterCommunication(cellchat, min.cells = 10)
```

Infer the cell-cell communication at a signaling pathway level

```{r computeCommunProbPathway, eval=FALSE}
cellchat <- computeCommunProbPathway(cellchat)
```

Calculate the aggregated cell-cell communication network

```{r aggregateNet, eval=FALSE}
cellchat <- aggregateNet(cellchat)
```

We can also visualize the aggregated cell-cell communication network. For example, showing the number of interactions or the total interaction strength (weights) between any two cell groups using circle plot.

```{r netVisual_circle, fig.width=10, fig.height=5}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1, 2), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

Due to the complicated cell-cell communication network, we can examine the signaling sent from each cell group. Here we also control the parameter edge.weight.max so that we can compare edge weights between differet networks.

```{r netVisual_circle-each, fig.width=12, fig.height=12}
mat <- cellchat@net$weight
par(mfrow = c(3, 4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```

## Visualization of cell-cell communication network {.tabset}

Visualize each signaling pathway using circle plot

```{r vis-each-signaling-pathway, fig.width=5, fig.height=5}
# Access all the signaling pathways showing significant communications
pathways.show.all <- cellchat@netP$pathways

src_list <- lapply(pathways.show.all, function(cluster){
    src <- c(
        "### {{cluster}}",
        "```{r vis-each-signaling-pathway-{{cluster}}}",
        "netVisual_aggregate(cellchat, signaling = '{{cluster}}', layout = 'circle')",
        "plotGeneExpression(cellchat, signaling = '{{cluster}}')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


## Systems analysis of cell-cell communication network

To facilitate the interpretation of the complex intercellular communication networks, CellChat quantitively measures networks through methods abstracted from graph theory, pattern recognition and manifold learning.

### Identify signaling roles of cell groups

CellChat allows ready identification of dominant senders, receivers, mediators and influencers in the intercellular communication network by computing several network centrality measures for each cell group. 

#### Compute and visualize the network centrality scores {.tabset}

```{r netAnalysis_computeCentrality}
# Compute the network centrality scores
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") # the slot 'netP' means the inferred intercellular communication network of signaling pathways
# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups


src_list <- lapply(pathways.show.all, function(cluster){
    src <- c(
        "##### {{cluster}}",
        "```{r netAnalysis_computeCentrality-{{cluster}}}",
        "netAnalysis_signalingRole_network(cellchat, signaling = '{{cluster}}', width = 8, height = 2.5, font.size = 10)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


#### Visualize the dominant senders (sources) and receivers (targets) in a 2D space

We also provide another intutive way to visualize the dominant senders (sources) and receivers (targets) in a 2D space using scatter plot.

```{r netAnalysis_signalingRole_scatter, fig.width=8, fig.height=7}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
gg1 <- netAnalysis_signalingRole_scatter(cellchat)
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
gg1
```

#### Identify signals contributing most to outgoing or incoming signaling of certain cell groups

We can also answer the question on which signals contributing most to outgoing or incoming signaling of certain cell groups.

```{r netAnalysis_signalingRole_heatmap, fig.width=10, fig.height=10}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing")
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming")
ht1 + ht2
```

### Identify global communication patterns

In addition to exploring detailed communications for individual pathways, an important question is how multiple cell groups and signaling pathways coordinate to function. CellChat employs a pattern recognition method to identify the global communication patterns.


#### Identify and visualize outgoing communication pattern of secreting cells

Outgoing patterns reveal how the sender cells (i.e. cells as signal source) coordinate with each other as well as how they coordinate with certain signaling pathways to drive communication.

Here we run selectK to infer the number of patterns.

```{r selectK-outgoing, fig.width=8, fig.height=6}
selectK(cellchat, pattern = "outgoing")
```

Both Cophenetic and Silhouette values begin to drop suddenly when the number of outgoing patterns is 4.

```{r identifyCommunicationPatterns-outgoing, fig.width=10, fig.width=8}
nPatterns = 4
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing", k = nPatterns)
```


```{r netAnalysis_river-outgoing}
# river plot
netAnalysis_river(cellchat, pattern = "outgoing")
#> Please make sure you have load `library(ggalluvial)` when running this function
```

```{r netAnalysis_dot-outgoing, fig.width=8, fig.height=5}
# dot plot
netAnalysis_dot(cellchat, pattern = "outgoing")
```

#### Identify and visualize incoming communication pattern of target cells

Incoming patterns show how the target cells (i.e. cells as signal receivers) coordinate with each other as well as how they coordinate with certain signaling pathways to respond to incoming signals.


```{r selectK-incoming, fig.width=8, fig.height=6}
selectK(cellchat, pattern = "incoming")
```

Both Cophenetic and Silhouette values begin to drop suddenly when the number of incoming patterns is 4.

```{r identifyCommunicationPatterns-incoming}
nPatterns = 4
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "incoming", k = nPatterns)
```


```{r netAnalysis_river-incoming}
# river plot
netAnalysis_river(cellchat, pattern = "incoming")
#> Please make sure you have load `library(ggalluvial)` when running this function
```

```{r netAnalysis_dot-incoming, fig.width=8, fig.height=5}
# dot plot
netAnalysis_dot(cellchat, pattern = "incoming")
```

### Manifold and classification learning analysis of signaling networks

Further, CellChat is able to quantify the similarity between all significant signaling pathways and then group them based on their cellular communication network similarity. Grouping can be done either based on the functional or structural similarity.


Functional similarity: High degree of functional similarity indicates major senders and receivers are similar, and it can be interpreted as the two signaling pathways or two ligand-receptor pairs exhibit similar and/or redundant roles. The functional similarity analysis requires the same cell population composition between two datasets.

Structural similarity: A structural similarity was used to compare their signaling network structure, without considering the similarity of senders and receivers.

#### Identify signaling groups based on their functional similarity

```{r computeNetSimilarity-functional, fig.width=8, fig.height=7}
cellchat <- computeNetSimilarity(cellchat, type = "functional")
cellchat <- netEmbedding(cellchat, type = "functional")
#> Manifold learning of the signaling networks for a single dataset
cellchat <- netClustering(cellchat, type = "functional")
#> Classification learning of the signaling networks for a single dataset
# Visualization in 2D-space
netVisual_embedding(cellchat, type = "functional", label.size = 3.5)
```

#### Identify signaling groups based on structure similarity

```{r computeNetSimilarity-structural, fig.width=8, fig.height=7}
cellchat <- computeNetSimilarity(cellchat, type = "structural")
cellchat <- netEmbedding(cellchat, type = "structural")
#> Manifold learning of the signaling networks for a single dataset
cellchat <- netClustering(cellchat, type = "structural")
#> Classification learning of the signaling networks for a single dataset
# Visualization in 2D-space
netVisual_embedding(cellchat, type = "structural", label.size = 3.5)
```

## Save

```{r save, eval=FALSE}
saveRDS(cellchat, file = here::here("output", DOCNAME, 'cellchat.LN_P.rds'))
```

