---
title: "vdj.tcr.rep_ov_uniq.byPatient"
---

```{r knitr, include = FALSE}
DOCNAME = "vdj.tcr.rep_ov_uniq.byPatient"
#dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

Load packages

```{r libaries, cache = FALSE, warning=FALSE, message=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(igraph)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(font_size=8) +
  theme(panel.background=element_rect(fill="white"),
        line = element_line(colour="black", lineend="round"),
        text = element_text(size=6, colour="black"),
        plot.title = element_text(size=6),
        axis.text = element_text(colour="black"),
        axis.title = element_text(size = 6, colour = "black"),
        axis.ticks = element_line(colour="black"),
        axis.ticks.length = unit(.1,"cm"),
        strip.background = element_blank(),
        axis.text.x=element_text(angle=45, hjust=1))
geom.text.size = 6/(14/5)
```

```{r source, cache = FALSE, message=FALSE, warning=FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
devtools::load_all('/T02Data/niuyw1/Rscir')
```

分析每个病人中，不同样本之间 unique shared clonotypes


## Set up

Load data

```{r load-data, cache=FALSE, message=FALSE, warning=FALSE}
metadata = read_csv(here::here("output", "04.rm_cells", "seurat_tcells.cellmeta.csv"))
clones = read_csv('/niuyw-usb-disk/Projects/scESCA/cellranger/TCR/200121_6samples.clonotype_tcr.csv')
```

Merge metadata with clonotype

```{r merge-meta-clonotypes}
clones = clones %>%
  filter(clonotype != 'None') %>%  # cells with alpha-beta pair
  mutate(clonotype_size = as.numeric(clonotype_size)) %>%
  mutate(clonal_expansion = case_when(
    clonotype_size == 1 ~ 'No',
    TRUE ~ 'Clonal'
  ))

d_4_p = metadata %>%
  left_join(clones, by = 'barcode') %>%
  filter(!is.na(clonotype))

head(d_4_p)
```

## Heatmap {.tabset}

By patient

```{r , results = "hide"}
tmp = unique(d_4_p$Patient)
src_list <- lapply(tmp, function(s){
    src <- c(
        "### {{s}} {.tabset}",
        "",
        "```{r  fig.height=8, fig.width=8}",
        "tmp = d_4_p %>% filter(Patient == '{{s}}') %>% group_by(clonotype, Origin) %>%",
        "summarise(n = n()) %>% mutate(n = n_distinct(clonotype, Origin)) %>%",
        "dplyr::select(clonotype, n) %>%  mutate(shared_by2 = ifelse(n == 2, 'Yes', 'No')) %>%",
        "dplyr::select(clonotype, shared_by2)",
        "df = d_4_p %>% filter(Patient == '{{s}}') %>% left_join(tmp, by = 'clonotype') %>% filter(shared_by2 == 'Yes')",
        "ov = RepOverlap(df, 'Origin', fraction = FALSE)",
        "PlotRepOverlap(ov, fraction = FALSE, title = '{{s}}')",
        "```",
        "CD8",
        "```{r  fig.height=8, fig.width=8}",
        "tmp = d_4_p %>% filter(Patient == '{{s}}', level_1=='CD8') %>% group_by(clonotype, Origin) %>%",
        "summarise(n = n()) %>% mutate(n = n_distinct(clonotype, Origin)) %>%",
        "dplyr::select(clonotype, n) %>%  mutate(shared_by2 = ifelse(n == 2, 'Yes', 'No')) %>%",
        "dplyr::select(clonotype, shared_by2)",
        "df = d_4_p %>% filter(Patient == '{{s}}', level_1=='CD8') %>% left_join(tmp, by = 'clonotype') %>% filter(shared_by2 == 'Yes')",
        "ov = RepOverlap(df, 'Origin', fraction = FALSE)",
        "PlotRepOverlap(ov, fraction = FALSE, title = '{{s}}')",
        "```",
        "",
        "CD4",
         "```{r  fig.height=8, fig.width=8}",
        "tmp = d_4_p %>% filter(Patient == '{{s}}', level_1=='CD4') %>% group_by(clonotype, Origin) %>%",
        "summarise(n = n()) %>% mutate(n = n_distinct(clonotype, Origin)) %>%",
        "dplyr::select(clonotype, n) %>%  mutate(shared_by2 = ifelse(n == 2, 'Yes', 'No')) %>%",
        "dplyr::select(clonotype, shared_by2)",
        "df = d_4_p %>% filter(Patient == '{{s}}', level_1=='CD4') %>% left_join(tmp, by = 'clonotype') %>% filter(shared_by2 == 'Yes')",
        "ov = RepOverlap(df, 'Origin', fraction = FALSE)",
        "PlotRepOverlap(ov, fraction = FALSE, title = '{{s}}')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Network {.tabset}

By patient

```{r , results = "hide"}
tmp = unique(d_4_p$Patient)
src_list <- lapply(tmp, function(s){
    src <- c(
        "### {{s}} {.tabset}",
        "```{r  fig.height=8, fig.width=8}",
        "tmp = d_4_p %>% filter(Patient == '{{s}}') %>% group_by(clonotype, Origin) %>%",
        "summarise(n = n()) %>% mutate(n = n_distinct(clonotype, Origin)) %>%",
        "dplyr::select(clonotype, n) %>%  mutate(shared_by2 = ifelse(n == 2, 'Yes', 'No')) %>%",
        "dplyr::select(clonotype, shared_by2)",
        "df = d_4_p %>% filter(Patient == '{{s}}') %>% left_join(tmp, by = 'clonotype') %>% filter(shared_by2 == 'Yes')",
        "ov = RepOverlap(df, 'Origin', fraction = FALSE)",
        "a = igraph::as_data_frame(graph_from_adjacency_matrix(ov, weighted = TRUE)) %>% filter(!is.na(weight))",
        "net = graph_from_data_frame(a, directed = FALSE)",
        "plot(net, edge.width=edge_attr(net)$weight)",
        "````",
        "",
        "CD8",
        "```{r  fig.height=8, fig.width=8}",
        "tmp = d_4_p %>% filter(Patient == '{{s}}', level_1=='CD8') %>% group_by(clonotype, Origin) %>%",
        "summarise(n = n()) %>% mutate(n = n_distinct(clonotype, Origin)) %>%",
        "dplyr::select(clonotype, n) %>%  mutate(shared_by2 = ifelse(n == 2, 'Yes', 'No')) %>%",
        "dplyr::select(clonotype, shared_by2)",
        "df = d_4_p %>% filter(Patient == '{{s}}', level_1=='CD8') %>% left_join(tmp, by = 'clonotype') %>% filter(shared_by2 == 'Yes')",
        "ov = RepOverlap(df, 'Origin', fraction = FALSE)",
        "a = igraph::as_data_frame(graph_from_adjacency_matrix(ov, weighted = TRUE)) %>% filter(!is.na(weight))",
        "net = graph_from_data_frame(a, directed = FALSE)",
        "plot(net, edge.width=edge_attr(net)$weight)",
        "```",
        "",
        "CD4",
        "```{r  fig.height=8, fig.width=8}",
        "tmp = d_4_p %>% filter(Patient == '{{s}}', level_1=='CD4') %>% group_by(clonotype, Origin) %>%",
        "summarise(n = n()) %>% mutate(n = n_distinct(clonotype, Origin)) %>%",
        "dplyr::select(clonotype, n) %>%  mutate(shared_by2 = ifelse(n == 2, 'Yes', 'No')) %>%",
        "dplyr::select(clonotype, shared_by2)",
        "df = d_4_p %>% filter(Patient == '{{s}}', level_1=='CD4') %>% left_join(tmp, by = 'clonotype') %>% filter(shared_by2 == 'Yes')",
        "ov = RepOverlap(df, 'Origin', fraction = FALSE)",
        "a = igraph::as_data_frame(graph_from_adjacency_matrix(ov, weighted = TRUE)) %>% filter(!is.na(weight))",
        "net = graph_from_data_frame(a, directed = FALSE)",
        "plot(net, edge.width=edge_attr(net)$weight)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Session info
