---
title: "vdj.tcr.gene_usage"
---

```{r knitr, include = FALSE}
DOCNAME = "vdj.tcr.gene_usage"
#dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = TRUE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(patchwork)
library(cowplot)
library(ComplexHeatmap)
library(pheatmap)
library(ggpubr)
library(ggrepel)
library(ggcorrplot)
library(circlize)
theme_set(theme_cowplot())

# rigde
library(ggridges)

# color
library(ggsci)
library(RColorBrewer)

# Seurat
library(Seurat)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
devtools::load_all('/T02Data/niuyw1/Rscir')
```

分析 TCR 的 gene usage

## Set up

Load data

```{r load-data, cache=FALSE, message=FALSE, warning=FALSE}
seurat = readRDS(here::here('output/04.rm_cells/seurat_tcells.rds'))
Idents(seurat) <- 'seurat_clusters'

# export cellmeta
vars = c("UMAP_1", "UMAP_2", "seurat_clusters", 'cellType', 'Patient', 'Tissue',
         'Origin2_n', "Origin3", 'Metastasis_n', "Origin", 'Source',
         "Origin2", "Metastasis",
         "level_1", "level_2", "level_3")
metadata <- FetchData(seurat,
                      vars = vars) %>%
  rownames_to_column('barcode')

clones = read_csv('/niuyw-usb-disk/Projects/scESCA/cellranger/TCR/200121_6samples.clonotype_tcr.csv')
```

## Neat data

Merge metadata with clonotypes

```{r merge-meta-clonotypes}
clones = clones %>%
  filter(clonotype != 'None') %>%  # cells with alpha-beta pair
  mutate(clonotype_size = as.numeric(clonotype_size)) %>%
  mutate(clonal_expansion = case_when(
    clonotype_size == 1 ~ 'No',
    TRUE ~ 'Clonal'
  ))

d_4_p = metadata %>%
  left_join(clones, by = 'barcode') %>%
  filter(!is.na(clonotype))

head(d_4_p)
```

## Overall

计算所有基因在每个 cluster 中的使用比例，然后观察其相关性

All clusters

```{r overall-usage, fig.width=8, fig.height=7, cache=FALSE}
mat = d_4_p %>%
  dplyr::select(barcode, level_3, ends_with(c("_v_gene", "_j_gene"))) %>%
  pivot_longer(cols = ends_with("gene"), names_to = "tmp", values_to = "gene") %>%
  group_by(level_3, gene) %>%
  summarise(n = n()) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols = "gene",
              names_from = "level_3", values_from = "prop",
              values_fill = 0) %>%
  column_to_rownames("gene")

pheatmap(cor(mat, method = "spearman"), clustering_method = "average")
```

CD4

```{r overall-usage-cd4, fig.width=6, fig.height=5, cache=FALSE}
mat_cd4 = mat %>%
  dplyr::select(starts_with("CD4"))

pheatmap(cor(mat_cd4, method = "spearman"), clustering_method = "average")
```

CD8

```{r overall-usage-cd8, fig.width=6, fig.height=5, cache=FALSE}
mat_cd8 = mat %>%
  dplyr::select(starts_with("CD8"))

pheatmap(cor(mat_cd8, method = "spearman"), clustering_method = "average")
```

## TRA Gene usage {.tabset}

### V gene {.tabset}

#### Cluster {.tabset}

```{r TRA-V-Cluster, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(seurat_clusters == '0') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)
```

```{r TRA-V-gene-usage-each-cluster, results = "hide"}
clusters = levels(d_4_p$seurat_clusters)
src_list <- lapply(clusters, function(s){
    src <- c(
        "##### Cluster {{s}} {.unnumbered}",
        "```{r TRA-V-gene-usage-cluster-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(seurat_clusters == {{s}}) %>% Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### level_3 {.tabset}

```{r TRA-V-level_3, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(level_3 == 'CD4-C1-Tcm') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)
```

```{r TRA-V-gene-usage-each-level_3, results = "hide"}
level_3 = levels(d_4_p$level_3)
src_list <- lapply(level_3, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRA-V-gene-usage-level_3-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(level_3 == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Tissue {.tabset}

```{r TRA-V-Tissue, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(Tissue == 'PBMC') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)
```

```{r TRA-V-gene-usage-each-Tissue, results = "hide"}
tissues = levels(d_4_p$Tissue)
src_list <- lapply(tissues, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRA-V-gene-usage-tissue-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Tissue == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Origin2_n {.tabset}

```{r TRA-V-Origin2_n, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(Origin2_n == 'PBMC') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)
```

```{r TRA-V-gene-usage-each-Origin2_n, results = "hide"}
tissues = levels(d_4_p$Origin2_n)
src_list <- lapply(tissues, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRA-V-gene-usage-Origin2_n-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Origin2_n == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Origin {.tabset}

```{r TRA-V-Origin, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(Origin == 'PBMC2') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)
```

```{r TRA-V-gene-usage-each-Origin, results = "hide"}
origins = levels(d_4_p$Origin)
src_list <- lapply(origins, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRA-V-gene-usage-origin-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Origin == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRA_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


### J gene {.tabset}

#### Cluster {.tabset}

```{r TRA-J-gene-usage-each-cluster, results = "hide"}
clusters = levels(d_4_p$seurat_clusters)
src_list <- lapply(clusters, function(s){
    src <- c(
        "##### Cluster {{s}} {.unnumbered}",
        "```{r TRA-J-gene-usage-cluster-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(seurat_clusters == {{s}}) %>% Rscir::PlotGroupAbundance(groupby = 'TRA_j_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


#### level_3 {.tabset}

```{r TRA-J-level_3, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(level_3 == 'CD4-C1-Tcm') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRA_j_gene', fraction = T)
```

```{r TRA-J-gene-usage-each-level_3, results = "hide"}
clusters = levels(d_4_p$level_3)
src_list <- lapply(clusters, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRA-J-gene-usage-level_3-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(level_3 == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRA_j_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Tissue {.tabset}

```{r TRA-J-gene-usage-each-Tissue, results = "hide"}
tissues = levels(d_4_p$Tissue)
src_list <- lapply(tissues, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRA-J-gene-usage-tissue-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Tissue == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRA_j_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Origin {.tabset}

```{r TRA-J-gene-usage-each-Origin, results = "hide"}
origins = levels(d_4_p$Origin)
src_list <- lapply(origins, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRA-J-gene-usage-origin-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Origin == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRA_j_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


## TRB Gene usage {.tabset}

### V gene {.tabset}

#### Cluster {.tabset}

```{r TRB-V-Cluster, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(seurat_clusters == '0') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)
```

```{r TRB-V-gene-usage-each-cluster, results = "hide"}
clusters = levels(d_4_p$seurat_clusters)
src_list <- lapply(clusters, function(s){
    src <- c(
        "##### Cluster {{s}} {.unnumbered}",
        "```{r TRB-V-gene-usage-cluster-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(seurat_clusters == {{s}}) %>% Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### level_3 {.tabset}

```{r TRB-V-level_3, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(level_3 == 'CD4-C1-Tcm') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)
```

```{r TRB-V-gene-usage-each-level_3, results = "hide"}
clusters = levels(d_4_p$level_3)
src_list <- lapply(clusters, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRB-V-gene-usage-level_3-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(level_3 == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Tissue {.tabset}

```{r TRB-V-Tissue, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(Tissue == 'PBMC') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)
```

```{r TRB-V-gene-usage-each-Tissue, results = "hide"}
tissues = levels(d_4_p$Tissue)
src_list <- lapply(tissues, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRB-V-gene-usage-tissue-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Tissue == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Origin2_n {.tabset}

```{r TRB-V-Origin2_n, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(Origin2_n == 'PBMC1') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)
```

```{r TRB-V-gene-usage-each-Origin2_n, results = "hide"}
tissues = levels(d_4_p$Origin2_n)
src_list <- lapply(tissues, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRB-V-gene-usage-Origin2_n-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Origin2_n == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Origin {.tabset}

```{r TRB-V-Origin, fig.height=4, fig.width=8, eval=FALSE}
d_4_p %>%
  filter(Origin == 'PBMC2') %>%
  Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)
```

```{r TRB-V-gene-usage-each-Origin, results = "hide"}
origins = levels(d_4_p$Origin)
src_list <- lapply(origins, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRB-V-gene-usage-origin-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Origin == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRB_v_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

### J gene {.tabset}

#### Cluster {.tabset}

```{r TRB-J-gene-usage-each-cluster, results = "hide"}
clusters = levels(d_4_p$seurat_clusters)
src_list <- lapply(clusters, function(s){
    src <- c(
        "##### Cluster {{s}} {.unnumbered}",
        "```{r TRB-J-gene-usage-cluster-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(seurat_clusters == {{s}}) %>% Rscir::PlotGroupAbundance(groupby = 'TRB_j_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### level_3 {.tabset}

```{r TRB-J-gene-usage-each-level_3, results = "hide"}
clusters = levels(d_4_p$level_3)
src_list <- lapply(clusters, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRB-J-gene-usage-level_3-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(level_3 == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRB_j_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Tissue {.tabset}

```{r TRB-J-gene-usage-each-Tissue, results = "hide"}
tissues = levels(d_4_p$Tissue)
src_list <- lapply(tissues, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRB-J-gene-usage-tissue-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Tissue == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRB_j_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Origin2_n {.tabset}

```{r TRB-J-gene-usage-each-Origin2_n, results = "hide"}
tissues = levels(d_4_p$Origin2_n)
src_list <- lapply(tissues, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRB-J-gene-usage-Origin2_n-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Origin2_n == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRB_j_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

#### Origin {.tabset}

```{r TRB-J-gene-usage-each-Origin, results = "hide"}
origins = levels(d_4_p$Origin)
src_list <- lapply(origins, function(s){
    src <- c(
        "##### {{s}} {.unnumbered}",
        "```{r TRB-J-gene-usage-origin-{{s}}, fig.height=4, fig.width=8}",
        "d_4_p %>% filter(Origin == '{{s}}') %>% Rscir::PlotGroupAbundance(groupby = 'TRB_j_gene', fraction = T)",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


## Spectratype

### TRA {.tabset}

#### Cluster

```{r Spectratype-TRA-Cluster, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRA_cdr3', target_col = 'seurat_clusters', fraction = T) +
  scale_fill_manual(values = cluster_color_maps)
```

#### Tissue

```{r Spectratype-TRA-Tissue, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRA_cdr3', target_col = 'Tissue', fraction = T) +
  scale_fill_manual(values = tissue_color_maps)
```

#### Origin2_n

```{r Spectratype-TRA-Origin2_n, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRA_cdr3', target_col = 'Origin2_n', fraction = T) +
  scale_fill_manual(values = origin2_color_maps)
```

#### Origin

```{r Spectratype-TRA-Origin, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRA_cdr3', target_col = 'Origin', fraction = T) +
  scale_fill_manual(values = origin_color_maps)
```

#### Metastasis

```{r Spectratype-TRA-Metastasis, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRA_cdr3', target_col = 'Metastasis', fraction = T) +
  scale_fill_manual(values = metastasis_color_maps)
```

### TRB {.tabset}

#### Cluster

```{r Spectratype-TRB-Cluster, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRB_cdr3', target_col = 'seurat_clusters', fraction = T) +
  scale_fill_manual(values = cluster_color_maps)
```

#### Tissue

```{r Spectratype-TRB-Tissue, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRB_cdr3', target_col = 'Tissue', fraction = T) +
  scale_fill_manual(values = tissue_color_maps)
```

#### Origin2_n

```{r Spectratype-TRB-Origin2_n, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRB_cdr3', target_col = 'Origin2_n', fraction = T) +
  scale_fill_manual(values = origin2_color_maps)
```

#### Origin

```{r Spectratype-TRB-Origin, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRB_cdr3', target_col = 'Origin', fraction = T) +
  scale_fill_manual(values = origin_color_maps)
```

#### Metastasis

```{r Spectratype-TRB-Metastasis, fig.height=5, fig.width=7}
Rscir::PlotSpectratype(d_4_p, 'TRB_cdr3', target_col = 'Metastasis', fraction = T) +
  scale_fill_manual(values = metastasis_color_maps)
```



