---
title: "vdj.tcr.STARTRAC"
---

```{r knitr, include = FALSE}
DOCNAME = "vdj.tcr.STARTRAC"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(patchwork)
library(cowplot)
library(ComplexHeatmap)
library(pheatmap)
library(ggpubr)
library(ggrepel)
library(ggcorrplot)
library(circlize)
theme_set(theme_cowplot())

# rigde
library(ggridges)

# color
library(ggsci)
library(RColorBrewer)

# STARTRAC
library("Startrac")
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
devtools::load_all('/T02Data/niuyw1/Rscir')
```

使用 STARTRAC 分析 VDJ

## Set up

Load data

```{r load-data, cache=FALSE, message=FALSE, warning=FALSE}
metadata = read_csv(here::here("output/04.rm_cells/seurat_tcells.cellmeta.csv"))
clones = read_csv('/niuyw-usb-disk/Projects/scESCA/cellranger/TCR/200121_6samples.clonotype_tcr.csv')

out = readRDS(here::here("output", DOCNAME, "STARTRAC.out.rds"))
```

## Neat data

Merge metadata with clonotypes

```{r neat}
# in.dat
in.dat = metadata %>%
  left_join(clones, by = 'barcode') %>%
  filter(!is.na(clonotype)) %>%
  dplyr::select(Cell_Name = barcode, clone.id = clonotype,
                patient = Patient, majorCluster = level_3, loc = Origin2_n)

head(in.dat)
```

## Run

```{r run, message=FALSE, warning=FALSE, eval=FALSE}
out <- Startrac.run(in.dat, proj="scESCC", cores=8, verbose=F)
```

## Plot all {.tabset}

### by patient

```{r fig.width=8, fig.height=8}
plot(out, index.type="cluster.all", byPatient=T)
```

### Not by patient

```{r fig.width=8, fig.height=8}
plot(out, index.type = "cluster.all", byPatient = F)
```

## expansion {.tabset}

### by patient

sort by mean

```{r fig.width=8, fig.height=3}
tmp_df = out@cluster.data %>%
  filter(aid != "scESCC") %>%
  group_by(majorCluster) %>%
  summarise(m = mean(expa, na.rm=TRUE))

# sort
tmp_cd8 = tmp_df %>%
  filter(str_detect(majorCluster, "CD8")) %>%
  arrange(-m)
tmp_cd4 = tmp_df %>%
  filter(str_detect(majorCluster, "CD4")) %>%
  arrange(-m)
tmp_treg = tmp_df %>%
  filter(str_detect(majorCluster, "Treg")) %>%
  arrange(-m)
tmp_gdt = tmp_df %>%
  filter(str_detect(majorCluster, "γδT")) %>%
  arrange(-m)
tmp_nkt = tmp_df %>%
  filter(str_detect(majorCluster, "NK")) %>%
  arrange(-m)
tmp_un = tmp_df %>%
  filter(str_detect(majorCluster, "Unknown")) %>%
  arrange(-m)
tmp_df = rbind(tmp_cd8, tmp_cd4, tmp_treg, tmp_gdt, tmp_nkt, tmp_un)

out@cluster.data %>%
  filter(aid != "scESCC") %>%
  mutate(majorCluster = factor(majorCluster, levels = tmp_df$majorCluster)) %>%
  ggplot(aes(x = majorCluster, y = expa, color  = majorCluster)) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = t_level_3_color) +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        legend.position = "none")
```

```{r}
tmp_df %>%
  dplyr::rename(expa = m) %>%
  mutate(type = case_when(
    expa > 0.1 ~ "+++",
    expa > 0.05 & expa <= 0.1 ~ "++",
    expa > 0.005 & expa <= 0.05 ~ "+",
    expa <= 0.005 ~ "-"
  )) %>%
  DT::datatable()
```

### Not by patient

```{r fig.width=8, fig.height=3}
d4p = out@cluster.data %>%
  filter(aid == "scESCC") %>%
  mutate(m = expa)

# sort
tmp_cd8 = d4p %>%
  filter(str_detect(majorCluster, "CD8")) %>%
  arrange(-m)
tmp_cd4 = d4p %>%
  filter(str_detect(majorCluster, "CD4")) %>%
  arrange(-m)
tmp_treg = d4p %>%
  filter(str_detect(majorCluster, "Treg")) %>%
  arrange(-m)
tmp_gdt = d4p %>%
  filter(str_detect(majorCluster, "γδT")) %>%
  arrange(-m)
tmp_nkt = d4p %>%
  filter(str_detect(majorCluster, "NK")) %>%
  arrange(-m)
tmp_un = d4p %>%
  filter(str_detect(majorCluster, "Unknown")) %>%
  arrange(-m)
tmp_df = rbind(tmp_cd8, tmp_cd4, tmp_treg, tmp_gdt, tmp_nkt, tmp_un)

d4p %>%
  mutate(majorCluster = factor(majorCluster, levels = tmp_df$majorCluster)) %>%
  ggplot(aes(x = majorCluster, y = m, fill = majorCluster)) +
  geom_col() +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = t_level_3_color) +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        legend.position = "none")
```

```{r}
d4p %>%
  mutate(type = case_when(
    expa > 0.1 ~ "+++",
    expa > 0.05 & expa <= 0.1 ~ "++",
    expa > 0.005 & expa <= 0.05 ~ "+",
    expa <= 0.005 ~ "-"
  )) %>%
  DT::datatable()
```

## expansion and proliferative freq. {.tabset}

### Not by patient

```{r fig.width=6, fig.height=6}
df_cellNum = metadata %>%
  group_by(level_3) %>%
  summarise(cellNum = log2(n())) %>%
  dplyr::rename(majorCluster = level_3)

metadata %>%
  group_by(level_3, Phase) %>%
  summarise(n = n()) %>%
  mutate(freq = n/sum(n)) %>%
  filter(Phase == "G2M") %>%
  dplyr::rename(majorCluster = level_3) %>%
  left_join(out@cluster.data %>%
              filter(aid == "scESCC"), by = "majorCluster") %>%
  left_join(df_cellNum) %>%
  ggplot(aes(x = freq, y = expa, color = majorCluster, size = cellNum)) +
  geom_point() +
  scale_color_manual(values = t_level_3_color) +
  geom_text_repel(aes(label = majorCluster), max.overlaps = 20) +
  labs(x = "Freq. of prolif.", size = "log2(cellNum)") +
  guides(color = FALSE) +
  theme(legend.position = "top")
```

### by patient

```{r fig.width=6, fig.height=6}
df_cellNum = metadata %>%
  group_by(level_3) %>%
  summarise(cellNum = log2(n())) %>%
  dplyr::rename(majorCluster = level_3)

dat_expa = out@cluster.data %>%
  filter(aid != "scESCC") %>%
  group_by(majorCluster) %>%
  summarise(m_expa = mean(expa, na.rm = TRUE))

dat_freqProlif = metadata %>%
  group_by(Patient, level_3, Phase) %>%
  summarise(n = n()) %>%
  mutate(freq = n/sum(n)) %>%
  filter(Phase == "G2M", n >= 10) %>% # at least 10 cells
  ungroup() %>%
  group_by(level_3) %>% #filter(level_3 == "CD8-C12-Tex")
  summarise(m_freq = mean(freq)) %>%
  dplyr::rename(majorCluster = level_3)

dat_expa %>%
  left_join(dat_freqProlif) %>%
  left_join(df_cellNum) %>%
  ggplot(aes(x = m_freq, y = m_expa, color = majorCluster, size = cellNum)) +
  geom_point() +
  scale_color_manual(values = t_level_3_color) +
  geom_text_repel(aes(label = majorCluster), max.overlaps = 20) +
  labs(x = "Freq. of prolif.", size = "log2(cellNum)") +
  guides(color = FALSE) +
  theme(legend.position = "top")
```

## expa. vs. migr. {.tabset}

### Not by patient

```{r fig.width=6, fig.height=5}
out@cluster.data %>%
  filter(aid == "scESCC") %>%
  ggplot(aes(x = expa, y = migr, color = majorCluster)) +
  geom_point() +
  geom_text_repel(aes(label = majorCluster)) +
  scale_color_manual(values = t_level_3_color) +
  #labs(x = "") +
  theme(legend.position = "none")
```

### by patient

```{r fig.width=6, fig.height=5}
out@cluster.data %>%
  filter(aid != "scESCC") %>%
  group_by(majorCluster) %>%
  summarise(m_expa = mean(expa, na.rm = TRUE),
            m_migr = mean(migr, na.rm = TRUE)) %>%
  ggplot(aes(x = m_expa, y = m_migr, color = majorCluster)) +
  geom_point() +
  geom_text_repel(aes(label = majorCluster)) +
  scale_color_manual(values = t_level_3_color) +
  theme(legend.position = "none")
```

## migration {.tabset}

### by patient

sort by mean

```{r fig.width=8, fig.height=3}
tmp_df = out@cluster.data %>%
  filter(aid != "scESCC") %>%
  group_by(majorCluster) %>%
  summarise(m = mean(migr, na.rm=TRUE))

# sort
tmp_cd8 = tmp_df %>%
  filter(str_detect(majorCluster, "CD8")) %>%
  arrange(-m)
tmp_cd4 = tmp_df %>%
  filter(str_detect(majorCluster, "CD4")) %>%
  arrange(-m)
tmp_treg = tmp_df %>%
  filter(str_detect(majorCluster, "Treg")) %>%
  arrange(-m)
tmp_gdt = tmp_df %>%
  filter(str_detect(majorCluster, "γδT")) %>%
  arrange(-m)
tmp_nkt = tmp_df %>%
  filter(str_detect(majorCluster, "NK")) %>%
  arrange(-m)
tmp_un = tmp_df %>%
  filter(str_detect(majorCluster, "Unknown")) %>%
  arrange(-m)
tmp_df = rbind(tmp_cd8, tmp_cd4, tmp_treg, tmp_gdt, tmp_nkt, tmp_un)

out@cluster.data %>%
  filter(aid != "scESCC") %>%
  mutate(majorCluster = factor(majorCluster, levels = tmp_df$majorCluster)) %>%
  ggplot(aes(x = majorCluster, y = migr, color  = majorCluster)) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = t_level_3_color) +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        legend.position = "none")
```

```{r}
tmp_df %>%
  dplyr::rename(migr = m) %>%
  mutate(type = case_when(
    migr > 0.5 ~ "+++",
    migr > 0.2 & migr <= 0.5 ~ "++",
    migr > 0.1 & migr <= 0.2 ~ "+",
    migr <= 0.1 ~ "-"
  )) %>%
  DT::datatable()
```

### Not by patient

```{r fig.width=8, fig.height=3}
d4p = out@cluster.data %>%
  filter(aid == "scESCC") %>%
  mutate(m = migr)

# sort
tmp_cd8 = d4p %>%
  filter(str_detect(majorCluster, "CD8")) %>%
  arrange(-m)
tmp_cd4 = d4p %>%
  filter(str_detect(majorCluster, "CD4")) %>%
  arrange(-m)
tmp_treg = d4p %>%
  filter(str_detect(majorCluster, "Treg")) %>%
  arrange(-m)
tmp_gdt = d4p %>%
  filter(str_detect(majorCluster, "γδT")) %>%
  arrange(-m)
tmp_nkt = d4p %>%
  filter(str_detect(majorCluster, "NK")) %>%
  arrange(-m)
tmp_un = d4p %>%
  filter(str_detect(majorCluster, "Unknown")) %>%
  arrange(-m)
tmp_df = rbind(tmp_cd8, tmp_cd4, tmp_treg, tmp_gdt, tmp_nkt, tmp_un)

d4p %>%
  mutate(majorCluster = factor(majorCluster, levels = tmp_df$majorCluster)) %>%
  ggplot(aes(x = majorCluster, y = m, fill = majorCluster)) +
  geom_col() +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = t_level_3_color) +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        legend.position = "none")
```

```{r}
d4p %>%
  mutate(type = case_when(
    migr > 0.5 ~ "+++",
    migr > 0.2 & migr <= 0.5 ~ "++",
    migr > 0.1 & migr <= 0.2 ~ "+",
    migr <= 0.1 ~ "-"
  )) %>%
  DT::datatable()
```

## migration and proliferative freq. {.tabset}

### Not by patient

```{r fig.width=6, fig.height=6}
df_cellNum = metadata %>%
  group_by(level_3) %>%
  summarise(cellNum = log2(n())) %>%
  dplyr::rename(majorCluster = level_3)

metadata %>%
  group_by(level_3, Phase) %>%
  summarise(n = n()) %>%
  mutate(freq = n/sum(n)) %>%
  filter(Phase == "G2M") %>%
  dplyr::rename(majorCluster = level_3) %>%
  left_join(out@cluster.data %>%
              filter(aid == "scESCC"), by = "majorCluster") %>%
  left_join(df_cellNum) %>%
  ggplot(aes(x = freq, y = migr, color = majorCluster, size = cellNum)) +
  geom_point() +
  scale_color_manual(values = t_level_3_color) +
  geom_text_repel(aes(label = majorCluster), max.overlaps = 20) +
  labs(x = "Freq. of prolif.", size = "log2(cellNum)") +
  guides(color = FALSE) +
  theme(legend.position = "top")
```

### by patient

```{r fig.width=6, fig.height=6}
df_cellNum = metadata %>%
  group_by(level_3) %>%
  summarise(cellNum = log2(n())) %>%
  dplyr::rename(majorCluster = level_3)

dat_migr = out@cluster.data %>%
  filter(aid != "scESCC") %>%
  group_by(majorCluster) %>%
  summarise(m_migr = mean(migr, na.rm = TRUE))

dat_freqProlif = metadata %>%
  group_by(Patient, level_3, Phase) %>%
  summarise(n = n()) %>%
  mutate(freq = n/sum(n)) %>%
  filter(Phase == "G2M", n >= 10) %>% # at least 10 cells
  ungroup() %>%
  group_by(level_3) %>% #filter(level_3 == "CD8-C12-Tex")
  summarise(m_freq = mean(freq)) %>%
  dplyr::rename(majorCluster = level_3)

dat_migr %>%
  left_join(dat_freqProlif) %>%
  left_join(df_cellNum) %>%
  ggplot(aes(x = m_freq, y = m_migr, color = majorCluster, size = cellNum)) +
  geom_point() +
  scale_color_manual(values = t_level_3_color) +
  geom_text_repel(aes(label = majorCluster), max.overlaps = 20) +
  labs(x = "Freq. of prolif.", size = "log2(cellNum)") +
  guides(color = FALSE) +
  theme(legend.position = "top")
```

## pairwise migration {.tabset}

### Not by patient

```{r fig.width=8, fig.height=20}
plot(out, index.type="pairwise.migr", byPatient=F)
```

heatmap

```{r fig.width=8, fig.height=8}
dat.plot = out@pIndex.migr %>%
  filter(aid == "scESCC") %>%
  dplyr::select(-aid) %>%
  column_to_rownames("majorCluster") %>%
  as.matrix()
yrange = pretty(dat.plot)
col_fun = circlize::colorRamp2(seq(0, max(yrange), length=50),
                               colorRampPalette(rev(brewer.pal(n=7,name="RdBu")))(50),
                               space = "LAB")

Heatmap(dat.plot, col = col_fun, name = "pIndex.migr")
```

CD8/CD4

```{r fig.width=7, fig.height=4}
dat.plot.cd4 = out@pIndex.migr %>%
  filter(aid == "scESCC", str_detect(majorCluster, "CD4")) %>%
  dplyr::select(-aid) %>%
  column_to_rownames("majorCluster") %>%
  as.matrix()
dat.plot.cd4[is.na(dat.plot.cd4)] <- 0

dat.plot.cd8 = out@pIndex.migr %>%
  filter(aid == "scESCC", str_detect(majorCluster, "CD8")) %>%
  dplyr::select(-aid) %>%
  column_to_rownames("majorCluster") %>%
  as.matrix()
dat.plot.cd8[is.na(dat.plot.cd8)] <- 0

common_min = min(c(dat.plot.cd4, dat.plot.cd8))
common_max = max(c(dat.plot.cd4, dat.plot.cd8))
col_fun = circlize::colorRamp2(seq(common_min, common_max, length=50),
                               colorRampPalette(rev(brewer.pal(n=7, name="RdBu")))(50),
                               space = "LAB")
Heatmap(dat.plot.cd8, col = col_fun, name = "pIndex.migr")
Heatmap(dat.plot.cd4, col = col_fun, name = "pIndex.migr")
```

## transition {.tabset}

### by patient

sort by mean

```{r fig.width=8, fig.height=3}
tmp_df = out@cluster.data %>%
  filter(aid != "scESCC") %>%
  group_by(majorCluster) %>%
  summarise(m = mean(tran, na.rm=TRUE))

# sort
tmp_cd8 = tmp_df %>%
  filter(str_detect(majorCluster, "CD8")) %>%
  arrange(-m)
tmp_cd4 = tmp_df %>%
  filter(str_detect(majorCluster, "CD4")) %>%
  arrange(-m)
tmp_treg = tmp_df %>%
  filter(str_detect(majorCluster, "Treg")) %>%
  arrange(-m)
tmp_gdt = tmp_df %>%
  filter(str_detect(majorCluster, "γδT")) %>%
  arrange(-m)
tmp_nkt = tmp_df %>%
  filter(str_detect(majorCluster, "NK")) %>%
  arrange(-m)
tmp_un = tmp_df %>%
  filter(str_detect(majorCluster, "Unknown")) %>%
  arrange(-m)
tmp_df = rbind(tmp_cd8, tmp_cd4, tmp_treg, tmp_gdt, tmp_nkt, tmp_un)

out@cluster.data %>%
  filter(aid != "scESCC") %>%
  mutate(majorCluster = factor(majorCluster, levels = tmp_df$majorCluster)) %>%
  ggplot(aes(x = majorCluster, y = tran, color  = majorCluster)) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = t_level_3_color) +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        legend.position = "none")
```

```{r}
tmp_df %>%
  dplyr::rename(tran = m) %>%
  mutate(type = case_when(
    tran > 0.2 ~ "+++",
    tran > 0.1 & tran <= 0.2 ~ "++",
    tran > 0.05 & tran <= 0.1 ~ "+",
    tran <= 0.05 ~ "-"
  )) %>%
  DT::datatable()
```

### Not by patient

```{r fig.width=8, fig.height=3}
d4p = out@cluster.data %>%
  filter(aid == "scESCC") %>%
  mutate(m = tran)

# sort
tmp_cd8 = d4p %>%
  filter(str_detect(majorCluster, "CD8")) %>%
  arrange(-m)
tmp_cd4 = d4p %>%
  filter(str_detect(majorCluster, "CD4")) %>%
  arrange(-m)
tmp_treg = d4p %>%
  filter(str_detect(majorCluster, "Treg")) %>%
  arrange(-m)
tmp_gdt = d4p %>%
  filter(str_detect(majorCluster, "γδT")) %>%
  arrange(-m)
tmp_nkt = d4p %>%
  filter(str_detect(majorCluster, "NK")) %>%
  arrange(-m)
tmp_un = d4p %>%
  filter(str_detect(majorCluster, "Unknown")) %>%
  arrange(-m)
tmp_df = rbind(tmp_cd8, tmp_cd4, tmp_treg, tmp_gdt, tmp_nkt, tmp_un)

d4p %>%
  mutate(majorCluster = factor(majorCluster, levels = tmp_df$majorCluster)) %>%
  ggplot(aes(x = majorCluster, y = m, fill = majorCluster)) +
  geom_col() +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = t_level_3_color) +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        legend.position = "none")
```

```{r}
d4p %>%
  mutate(type = case_when(
    tran > 0.2 ~ "+++",
    tran > 0.1 & tran <= 0.2 ~ "++",
    tran > 0.05 & tran <= 0.1 ~ "+",
    tran <= 0.05 ~ "-"
  )) %>%
  DT::datatable()
```

## pairwise transition {.tabset}

### Not by patient

```{r fig.width=8, fig.height=8}
plot(out, index.type="pairwise.tran", byPatient=T)
```

CD8/CD4

```{r fig.width=5, fig.height=4}
dat.plot.cd4 = out@pIndex.tran %>%
  filter(aid == "scESCC", str_detect(majorCluster, "CD4")) %>%
  dplyr::select(starts_with("CD4"), majorCluster) %>%
  column_to_rownames("majorCluster") %>%
  as.matrix()
dat.plot.cd4[is.na(dat.plot.cd4)] <- 0

dat.plot.cd8 = out@pIndex.tran %>%
  filter(aid == "scESCC", str_detect(majorCluster, "CD8")) %>%
  dplyr::select(starts_with("CD8"), majorCluster) %>%
  column_to_rownames("majorCluster") %>%
  as.matrix()
dat.plot.cd8[is.na(dat.plot.cd8)] <- 0

common_min = min(c(dat.plot.cd4, dat.plot.cd8))
common_max = max(c(dat.plot.cd4, dat.plot.cd8))
col_fun = circlize::colorRamp2(seq(common_min, common_max, length=50),
                               colorRampPalette(rev(brewer.pal(n=7,name="RdBu")))(50),
                               space = "LAB")
Heatmap(dat.plot.cd8, col = col_fun, name = "pIndex.tran")
Heatmap(dat.plot.cd4, col = col_fun, name = "pIndex.tran")
```

## Save

```{r save, eval=FALSE}
saveRDS(out, file = here::here("output", DOCNAME, "STARTRAC.out.rds"))
```

## Session

