---
title: "vdj.tcr.rep_ov.cd8_tex"
---

```{r knitr, include = FALSE}
DOCNAME = "vdj.tcr.rep_ov.cd4_cd8"
#dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

Load packages

```{r libaries, cache = FALSE, warning=FALSE, message=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(font_size=8) +
  theme(panel.background=element_rect(fill="white"),
        line = element_line(colour="black", lineend="round"),
        text = element_text(size=6, colour="black"),
        plot.title = element_text(size=6),
        axis.text = element_text(colour="black"),
        axis.title = element_text(size = 6, colour = "black"),
        axis.ticks = element_line(colour="black"),
        axis.ticks.length = unit(.1,"cm"),
        strip.background = element_blank(),
        axis.text.x=element_text(angle=45, hjust=1))
geom.text.size = 6/(14/5)
```

```{r source, cache = FALSE, message=FALSE, warning=FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
devtools::load_all('/T02Data/niuyw1/Rscir')
```

分析 TCR 的 clonotype sharing

参考：Yost, K. E., Satpathy, A. T., Wells, D. K., Qi, Y., Wang, C., Kageyama, R., McNamara, K. L., Granja, J. M., Sarin, K. Y., Brown, R. A., Gupta, R. K., Curtis, C., Bucktrout, S. L., Davis, M. M., Chang, A. L. S., & Chang, H. Y. (2019). Clonal replacement of tumor-specific T cells following PD-1 blockade. Nature Medicine, 25(8), 1251–1259. https://doi.org/10.1038/s41591-019-0522-3


## Set up

Load data

```{r load-data, cache=FALSE, message=FALSE, warning=FALSE}
metadata = read_csv(here::here("output", "04.rm_cells", "seurat_tcells.cellmeta.csv")) %>%
  filter(level_1 %in% c("CD4", "CD8", "Treg"))
clones = read_csv('/niuyw-usb-disk/Projects/scESCA/cellranger/TCR/200121_6samples.clonotype_tcr.csv')
```

Merge metadata with clonotype

```{r merge-meta-clonotypes}
clones = clones %>%
  filter(clonotype != 'None') %>%  # cells with alpha-beta pair
  mutate(clonotype_size = as.numeric(clonotype_size)) %>%
  mutate(clonal_expansion = case_when(
    clonotype_size == 1 ~ 'No',
    TRUE ~ 'Clonal'
  ))

d_4_p = metadata %>%
  left_join(clones, by = 'barcode') %>%
  filter(!is.na(clonotype))

head(d_4_p)
```

## C11-Tex among different tissues

CD8-C11-Tex

```{r C11-Tex, fig.width=5, fig.height=5}
df = d_4_p %>% filter(level_3 == "CD8-C11-Tex")
PlotRepOverlap(RepOverlap(df, "Origin4"))
PlotRepOverlap(RepOverlap(df, "Origin4", fraction = F), fraction = F)
```

## Several Tex

```{r ov-Tex, fig.width=4, fig.height=4}
df = d_4_p %>% filter(level_1 == "CD8", level_2 == "Tex")
PlotRepOverlap(RepOverlap(df, "level_3"))
PlotRepOverlap(RepOverlap(df, "level_3", fraction = F), fraction = F)
```

## Session info

