---
title: "CellChat.tissueCompare"
---

```{r knitr, include = FALSE}
DOCNAME = "CellChat.byOrigin2"
#dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(ggplotify)
library(ggcorrplot)
library(ggpubr)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(font_size=8) +
  theme(line = element_line(colour="black", lineend="round"),
        text = element_text(size=6, colour="black"),
        plot.title = element_text(size=6),
        axis.text = element_text(colour="black"),
        axis.title = element_text(size = 6, colour = "black"),
        axis.ticks = element_line(colour="black"),
        axis.ticks.length = unit(.1, "cm"),
        strip.background = element_blank(),
        axis.text.x=element_text(angle=45, hjust=1))
geom.text.size = 6/(14/5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# fonts
library(extrafont)
#font_import()
#font_import(paths = "/T02Data/niuyw/.local/share/fonts")
loadfonts(device = "pdf", quiet = TRUE)

# color
library(ggsci)

# cellchat
library(CellChat)
library(NMF)
library(ggalluvial)

# Seurat
library(Seurat)

options(stringsAsFactors = FALSE)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load cellchat rds

```{r merge-rds, cache=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
origin2_n = c("PBMC1", "PBMC2", "LN_N", "LN_P", "Normal", "Adjacent", "Tumor")
group.new = c("CD4", "Treg", "CD8", "γδT", "NK/NKT",
              "B cells",
              "Mono", "MΦ", "cDC", "tDC", "pDC", "Mast", "CD34+ cells",
              "Epithelia", "Fibroblasts", "Endothelia", "Platelets", "Unknown")
obj.lst = lapply(origin2_n, function(x){
  a = readRDS(here::here("output/CellChat.byOrigin2", paste0("cellchat.", paste0(x, ".rds"))))
  liftCellChat(a, group.new = group.new)
})
names(obj.lst) = origin2_n
cellchat <- mergeCellChat(obj.lst, add.names = origin2_n)
cellchat
saveRDS(cellchat, file = here::here("output", DOCNAME, "cellchat.merged.rds"))
```

```{r load-rds, cache=FALSE, message=FALSE, warning=FALSE}
cellchat = readRDS(here::here("output", DOCNAME, "cellchat.merged.rds"))
cellchat
```

## Total number of interactions and interaction strength

To answer on question on whether the cell-cell communication is enhanced or not, CellChat compares the the total number of interactions and interaction strength of the inferred cell-cell communication networks from different biological conditions.


```{r cmp-total-number-interactions-and-strength, fig.width=8, fig.height=4}
gg1 <- compareInteractions(cellchat, show.legend = F)
gg2 <- compareInteractions(cellchat, show.legend = F, measure = "weight")
gg1 + gg2
```

## Aggregated network

### Number {.tabset}

```{r aggregated-network-count, results='hide'}
num.inter = sapply(cellchat@net, function(x){max(x$count)}, simplify = T)

src_list <- lapply(origin2_n, function(cluster){
    src <- c(
        "#### {{cluster}}",
        "```{r aggregated-network-count-{{cluster}}, fig.width=6, fig.height=5}",
        "obj = cellchat@net[['{{cluster}}']]",
        "mat = obj$count",
        "netVisual_circle(mat, weight.scale = T, edge.weight.max = max(num.inter), label.edge= F, title.name = '{{cluster}}')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

### Strength {.tabset}

```{r aggregated-network-weight, results='hide'}
stren.inter = sapply(cellchat@net, function(x){max(x$weight)}, simplify = T)

src_list <- lapply(origin2_n, function(cluster){
    src <- c(
        "#### {{cluster}}",
        "```{r aggregated-network-weight-{{cluster}}, fig.width=6, fig.height=5}",
        "obj = cellchat@net[['{{cluster}}']]",
        "mat = obj$weight",
        "netVisual_circle(mat, weight.scale = T, edge.weight.max = max(stren.inter), label.edge= F, title.name = '{{cluster}}')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Aggregated network (each) {.tabset}

Weight

```{r aggregated-network-weight-each, results='hide'}
src_list <- lapply(origin2_n, function(cluster){
    src <- c(
        "### {{cluster}}",
        "```{r aggregated-network-weight-each-{{cluster}}, fig.width=12, fig.height=12}",
        "obj = cellchat@net[['{{cluster}}']]",
        "mat = obj$weight",
        "par(mfrow = c(3, 4), xpd=TRUE)",
        "for (i in 1:nrow(mat)) {",
        "mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))",
        "mat2[i, ] <- mat[i, ]",
        "netVisual_circle(mat2, weight.scale = T, edge.weight.max = max(stren.inter), title.name = rownames(mat)[i])}",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


## Differential

### Circle {.tabset}

The differential number of interactions or interaction strength in the cell-cell communication network between two datasets can be visualized using circle plot, where red (or blue) colored edges represent increased (or decreased) signaling in the second dataset compared to the first one.

#### PBMC1 & PBMC2

```{r diff-circle-PBMC1-PBMC2, fig.width=10, fig.height=5}
par(mfrow = c(1,2), xpd=TRUE)
netVisual_diffInteraction(cellchat, comparison = c(1, 2), weight.scale = T)
netVisual_diffInteraction(cellchat, comparison = c(1, 2), weight.scale = T, measure = "weight")
```

#### LN_N & LN_P

```{r diff-circle-LN_N-LN_P, fig.width=10, fig.height=5}
par(mfrow = c(1, 2), xpd=TRUE)
netVisual_diffInteraction(cellchat, comparison = c(3, 4), weight.scale = T)
netVisual_diffInteraction(cellchat, comparison = c(3, 4), weight.scale = T, measure = "weight")
```

#### Normal & Tumor

```{r diff-circle-Normal-Tumor, fig.width=10, fig.height=5}
par(mfrow = c(1, 2), xpd=TRUE)
netVisual_diffInteraction(cellchat, comparison = c(5, 7), weight.scale = T)
netVisual_diffInteraction(cellchat, comparison = c(5, 7), weight.scale = T, measure = "weight")
```

### Heatmap {.tabset}

We can also show differential number of interactions or interaction strength in a greater details using a heatmap. The top colored bar plot represents the sum of column of values displayed in the heatmap (incoming signaling). The right colored bar plot represents the sum of row of values (outgoing signaling). In the colorbar, red (or blue) represents increased (or decreased) signaling in the second dataset compared to the first one.

#### PBMC1 & PBMC2

```{r diff-Heatmap-PBMC1-PBMC2, fig.width=12, fig.height=6}
gg1 <- netVisual_heatmap(cellchat, comparison = c(1, 2))
gg2 <- netVisual_heatmap(cellchat, measure = "weight", comparison = c(1, 2))
gg1 + gg2
```

#### LN_N & LN_P

```{r diff-Heatmap-LN_N-LN_P, fig.width=12, fig.height=6}
gg1 <- netVisual_heatmap(cellchat, comparison = c(3, 4))
gg2 <- netVisual_heatmap(cellchat, measure = "weight", comparison = c(3, 4))
gg1 + gg2
```

#### Normal & Tumor

```{r diff-Heatmap-Normal-Tumor, fig.width=12, fig.height=6}
gg1 <- netVisual_heatmap(cellchat, comparison = c(5, 7))
gg2 <- netVisual_heatmap(cellchat, measure = "weight", comparison = c(5, 7))
gg1 + gg2
```

## Compare the major sources and targets in 2D space

Comparing the outgoing and incoming interaction strength in 2D space allows ready identification of the cell populations with significant changes in sending or receiving signals between different datasets.

```{r cmp-major-sour-target-in-2D, fig.width=12, fig.height=10, message=FALSE, warning=FALSE}
num.link <- sapply(obj.lst, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg <- list()
for (i in 1:length(obj.lst)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(obj.lst[[i]], title = names(obj.lst)[i], weight.MinMax = weight.MinMax)
}
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
patchwork::wrap_plots(plots = gg, ncol = 3)
```

## Visually compare each pathway {.tabset}

Similar to the CellChat analysis of individual dataset, we can visualize the cell-cell communication network using Hierarchy plot.

**NB**: only common pathways could be compared.

```{r vis-each-pathway, results='hide'}
a = lapply(obj.lst, function(x){x@netP$pathways})
pathways.show.all <- Reduce(intersect, a)

src_list <- lapply(pathways.show.all, function(x){
    src <- c(
        "### {{x}}",
        "",
        "```{r vis-each-signaling-pathway-{{x}}, fig.width=12, fig.height=12}",
        "weight.max <- getMaxWeight(obj.lst, slot.name = 'netP', attribute = '{{x}}')",
        "par(mfrow = c(3, 4), xpd=TRUE)",
        "for (i in 1:length(obj.lst)) {",
        "netVisual_aggregate(obj.lst[[i]], signaling = '{{x}}', layout = 'circle',",
        "edge.weight.max = weight.max[1], edge.width.max = 10,",
        "signaling.name = paste('{{x}}', names(obj.lst)[i]))}",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

