---
title: "epithelia.cor_bulk"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r knitr, include = FALSE}
DOCNAME = "epithelia.cor_bulk"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)
#dir.create(here::here("docs/assets", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

We thought it is potentially malignant if an epithelial cell has higher similarity to ESCCs than to normal esophagus, based on the comparison to bulk RNA-Seq profiles.

> For each cell, we computed the correlation between its profile and each bulk profile of ESCC samples and normal tissues. We then tested, for each cell, if it was more similar to the ESCC compared to the normal tissues, by applying a one-sided Wilcoxon ranksum test on the correlation coefficients that were obtained for that cell. Cells that were more similar to the normal tissues (p < 0.05, Wilcoxon ranksum test) were considered as potentially nonmalignant according to this criterion.

参考文献：

- Jerby-Arnon, L., Shah, P., Cuoco, M.S., Rodman, C., Su, M.-J., Melms, J.C., Leeson, R., Kanodia, A., Mei, S., Lin, J.-R., et al. (2018). A Cancer Cell Program Promotes T Cell Exclusion and Resistance to Checkpoint Blockade. Cell 175, 984-997.e24.

## Setup

Load packages

```{r libaries, cache=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(cowplot)
theme_set(theme_cowplot())
library(patchwork)

# Seurat
library(Seurat)

library(SummarizedExperiment)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load rds

```{r load-rds, cache=FALSE}
seurat = readRDS(here::here('output/04.rm_cells/seurat_epithelia.rds'))
seurat
```

Load GTEx exp data

```{r load-GTEx-exp, cache=FALSE}
gtex_data_dir = '/niuyw-usb-disk/Projects/scESCA/GTEx_v8'
gtex_tpm = readRDS(file.path(gtex_data_dir, 'GTEx_v8.tpm.Esophagus.rds'))
dim(gtex_tpm)
```

Load TCGA exp data

```{r prep-TCGA-exp, eval=FALSE}
tcga_data_dir = '/niuyw-usb-disk/Projects/scESCA/TCGA_ESCA'

# clinical
escc_cli = read_delim(file.path(tcga_data_dir, 'esca_cli.txt'), delim = '\t') %>%
  # ESCC only
  filter(histological_type == 'Esophagus Squamous Cell Carcinoma') %>%
  filter(sample_type == "Primary Tumor")

# exp
load(file.path(tcga_data_dir, 'TCGA-ESCA_exp_hg38.rda'))
# to FPKM matrix
esca_fpkm = assay(data)
colnames(esca_fpkm) = str_sub(colnames(esca_fpkm), 1, 15)
# keep ESCC samples
escc_fpkm = esca_fpkm[,intersect(colnames(esca_fpkm), escc_cli$sampleID)]

# ensembl id to gene symbol
id_map = clusterProfiler::bitr(rownames(escc_fpkm), fromType = 'ENSEMBL', toType = 'SYMBOL', OrgDb = 'org.Hs.eg.db')
# remove duplicate
id_map = id_map[!duplicated(id_map$ENSEMBL),]
id_map = id_map[!duplicated(id_map$SYMBOL),]

# prune genes
escc_fpkm = escc_fpkm[id_map$ENSEMBL,]
rownames(escc_fpkm) = id_map$SYMBOL

# write
saveRDS(escc_fpkm, file = file.path(tcga_data_dir, 'escc_fpkm.rds'))
```

```{r load-TCGA-exp, cache=FALSE}
tcga_data_dir = '/niuyw-usb-disk/Projects/scESCA/TCGA_ESCA'
escc_fpkm = readRDS(file.path(tcga_data_dir, 'escc_fpkm.rds'))
dim(escc_fpkm)
```

## Correlation with ESCC

Correlate the expression profiles of each cell to ESCC.

```{r cor-with-ESCC, eval=FALSE}
# genes common
genes_common = intersect(rownames(escc_fpkm), rownames(seurat))

# escc fpkm
escc_exp = escc_fpkm[genes_common,]

# subset seurat
sc_exp = GetAssayData(seurat, slot = 'data')[genes_common,]
sc_exp = as.matrix(expm1(sc_exp)) # to non-log values

# cor
cor_escc = cor(sc_exp, escc_exp, method = "spearman")

# save
saveRDS(cor_escc, file = here::here('output', DOCNAME, 'cor_escc.rds'))
```

```{r cor-with-ESCC-load, cache=FALSE}
cor_escc = readRDS(here::here('output', DOCNAME, 'cor_escc.rds'))
```

## Correlation with GTEx

Correlate the expression profiles of each cell to GTEx.

```{r cor-with-GTEx, eval=FALSE}
# genes common
genes_common = intersect(rownames(gtex_tpm), rownames(seurat))

# escc fpkm
gtex_exp = gtex_tpm[genes_common,]

# subset seurat
sc_exp = GetAssayData(seurat, slot = 'data')[genes_common,]
sc_exp = as.matrix(expm1(sc_exp)) # to non-log values

# cor
cor_gtex = cor(sc_exp, gtex_exp, method = "spearman")

# save
saveRDS(cor_gtex, file = here::here('output', DOCNAME, 'cor_gtex.rds'))
```

```{r cor-with-GTEx-load, cache=FALSE}
cor_gtex = readRDS(here::here('output', DOCNAME, 'cor_gtex.rds'))
```

## wilcox.test

Use one-side wilcox.test to identify malignant cells

```{r test, eval=FALSE}
p_value = sapply(1:nrow(cor_escc), function(i) {
  u = cor_escc[i,]
  v = cor_gtex[i,]
  p = wilcox.test(u, v, alternative = "greater")$p.value
  p
})

df = tibble(barcode = rownames(cor_escc), p = p_value) %>%
  mutate(p_adj = p.adjust(p, method = 'bonferroni')) %>% # use strict p adjust since the sample size is very large
  mutate(malignant = case_when(  # malignant?
    p_adj < 0.01 ~ 'Yes',
    p_adj >= 0.01 ~ 'No'
  ))
  
# metadata
vars = c("UMAP_1", "UMAP_2", "seurat_clusters", "Patient", "Tissue", "Origin", "Origin2", "Origin3", 'Metastasis', 'Drainage', "Source")
metadata <- FetchData(seurat,
                      vars = vars) %>%
  rownames_to_column('barcode')

# join metadata
df = df %>%
  mutate(malignant = factor(malignant, levels = c('No', 'Yes'))) %>%
  left_join(metadata, by = 'barcode')

# save
saveRDS(df, file = here::here('output', DOCNAME, 'epi.cor_with_bulk.rds'))
```

## Results {.tabset}

```{r test-load, cache=FALSE}
df = readRDS(here::here('output', DOCNAME, 'epi.cor_with_bulk.rds'))
table(df$malignant)
```

### UMAP

```{r by-UMAP, fig.height=8, fig.width=10}
# Adding cluster label to center of cluster on UMAP
umap_label <- FetchData(seurat, 
                        vars = c("seurat_clusters", "UMAP_1", "UMAP_2"))  %>%
        group_by(seurat_clusters) %>%
        summarise(x=mean(UMAP_1), y=mean(UMAP_2))

# ptsize
ptsize = get_pt_size(dim(df)[1])

df %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2)) +
  geom_point(aes(color = malignant), size = ptsize) +
  scale_color_manual(values = colors_2) +
  geom_text(data=umap_label, aes(label=seurat_clusters, x, y))
```

### Cluster

Potential malignant cells by seurat_clusters

```{r by-Cluster}
DT::datatable(
  df %>%
  group_by(seurat_clusters, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
)
```

```{r by-Cluster-p, fig.width=8, fig.height=3}
d_4_p = df %>%
  group_by(seurat_clusters, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)

o = d_4_p %>%
  filter(malignant == "Yes") %>%
  arrange(Percent) %>%
  pull(seurat_clusters) %>%
  as.character()

# not in
o = c(setdiff(df$seurat_clusters, o), o)

d_4_p %>%
  mutate(seurat_clusters = factor(seurat_clusters, levels = o)) %>%
  ggplot(aes(x = seurat_clusters, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  geom_hline(aes(yintercept = 75), linetype = 'dashed')
```

### Tissues

Potential malignant cells by Tissue.

```{r by-Tissue}
df %>%
  group_by(Tissue, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
```

```{r by-Tissue-p, fig.width=4, fig.height=3}
df %>%
  group_by(Tissue, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Tissue, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### Patient x Tissue

```{r by-Patient-Tissue}
DT::datatable(
  df %>%
  group_by(Patient, Tissue, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
)
```

```{r by-Patient-Tissue-p, fig.width=8, fig.height=5}
df %>%
  group_by(Patient, Tissue, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Tissue, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  facet_wrap(.~Patient)
```

### Metastasis

```{r by-metastasis, cache=FALSE}
df %>%
  group_by(Metastasis, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
```

```{r by-metastasis-p, fig.width=2.5, fig.height=3}
df %>%
  filter(Metastasis != "None") %>%
  group_by(Metastasis, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Metastasis, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### Patient x Metastasis

```{r by-Patient-Metastasis}
DT::datatable(
  df %>%
  group_by(Patient, Metastasis, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
)
```

```{r by-Patient-Metastasis-p, fig.width=5, fig.height=5}
df %>%
  filter(Metastasis != "None") %>%
  group_by(Patient, Metastasis, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Metastasis, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  facet_wrap(.~Patient)
```

### Drainage

```{r by-Drainage, cache=FALSE}
df %>%
  group_by(Drainage, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
```

```{r by-Drainage-p, fig.width=2.5, fig.height=3}
df %>%
  filter(Drainage != "None") %>%
  group_by(Drainage, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Drainage, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### Patient x Drainage

```{r by-Patient-Drainage}
DT::datatable(
  df %>%
  group_by(Patient, Drainage, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
)
```

```{r by-Patient-Drainage-p, fig.width=5, fig.height=5}
df %>%
  filter(Drainage != "None") %>%
  group_by(Patient, Drainage, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Drainage, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  facet_wrap(.~Patient)
```

### Origin2

PBMC1/PBMC2/LN N/LN P/Normal/Adjacent/Tumor

```{r by-Origin2, cache=FALSE}
df %>%
  group_by(Origin2, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
```

```{r by-Origin2-p, fig.width=4, fig.height=3}
df %>%
  group_by(Origin2, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin2, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### Patient x Origin2

```{r by-Patient-Origin2}
DT::datatable(
  df %>%
  group_by(Patient, Origin2, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
)
```

```{r by-Patient-Origin2-p, fig.width=8, fig.height=5}
df %>%
  group_by(Patient, Origin2, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin2, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  facet_wrap(.~Patient)
```

### Origin3

PBMC1/PBMC2/LN Up/LN Down/Normal/Adjacent/Tumor

```{r by-Origin3, cache=FALSE}
df %>%
  group_by(Origin3, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
```

```{r by-Origin3-p, fig.width=4, fig.height=3}
df %>%
  group_by(Origin3, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin3, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### Patient x Origin3

```{r by-Patient-Origin3}
DT::datatable(
  df %>%
  group_by(Patient, Origin3, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
)
```

```{r by-Patient-Origin3-p, fig.width=8, fig.height=5}
df %>%
  group_by(Patient, Origin3, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin3, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  facet_wrap(.~Patient)
```

### Patient

```{r by-Patient, cache=FALSE}
df %>%
  group_by(Patient, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
```

```{r by-Patient-p, fig.width=4, fig.height=3}
df %>%
  group_by(Patient, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Patient, y = Percent, fill = malignant)) +
  geom_col(color = 'white') +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


