---
title: "00.Fig.overview"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.overview"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# Seurat
library(Seurat)

# vis
# library(scRNAtoolVis)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load rds

```{r load-rds, cache=FALSE}
srat <- readRDS(here::here("output/04.rm_cells/seurat.rds"))
srat
```

## UMAP {.tabset}

### cell type

```{r fig.width=2, fig.height=2}
p.umap.cellType <- DimPlot(srat,
  reduction = "umap", group.by = "cellType",
  label = TRUE, label.size = geom.text.size, raster = TRUE
) +
  scale_color_manual(values = cell_color_maps) +
  labs(title = NULL) +
  my_theme +
  theme(
    legend.position = "none",
    plot.margin = margin(rep(0, 4), unit = "cm")
  )
p.umap.cellType

# saveRDS(p.umap.cellType, file = here::here("output", DOCNAME, "Fig.umap.cellType.rds"))
ggsave2(p.umap.cellType,
  filename = here::here("output", DOCNAME, "Fig.umap.cellType.pdf"),
  height = 2, width = 2, device = cairo_pdf
)
```

### Origin4

```{r fig.width=2.6, fig.height=2}
p.umap.origin4 <- DimPlot(srat,
  reduction = "umap", group.by = "Origin4",
  raster = TRUE
) +
  scale_color_manual(values = origin4_color_maps) +
  labs(title = NULL) +
  my_theme
p.umap.origin4

# saveRDS(p.umap.origin4, file = here::here("output", DOCNAME, "Fig.umap.Origin4.rds"))
ggsave2(p.umap.origin4,
  filename = here::here("output", DOCNAME, "Fig.umap.Origin4.pdf"),
  height = 2.6, width = 2, device = cairo_pdf
)
```

Split

```{r fig.width=7.5, fig.height=1.2}
p.umap.splitByOrigin4 <- DimPlot(srat,
  reduction = "umap",
  split.by = "Origin4", ncol = 8,
  raster = TRUE
) +
  scale_color_manual(values = cell_color_maps) +
  my_theme +
  NoAxes() +
  theme(
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.height = unit(.4, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
    plot.margin = margin(rep(0, 4), unit = "cm"),
    strip.background = element_rect(fill = "#f3f2f1")
  )
p.umap.splitByOrigin4

# saveRDS(p.umap.splitByOrigin4, file = here::here("output", DOCNAME, "Fig.umap.splitByOrigin4.rds"))
ggsave2(p.umap.splitByOrigin4,
  filename = here::here("output", DOCNAME, "Fig.umap.splitByOrigin4.pdf"),
  height = 1.2, width = 7.5, device = cairo_pdf
)
```

### Origin

```{r fig.width=3, fig.height=2}
p.umap.origin <- DimPlot(srat,
  reduction = "umap", group.by = "Origin",
  raster = TRUE
) +
  scale_color_manual(values = origin_color_maps) +
  labs(title = NULL) +
  my_theme
p.umap.origin

# saveRDS(p.umap.origin, file = here::here("output", DOCNAME, "Fig.umap.Origin.rds"))
ggsave2(p.umap.origin,
  filename = here::here("output", DOCNAME, "Fig.umap.Origin.pdf"),
  height = 2, width = 3, device = cairo_pdf
)
```

### Patient

```{r fig.width=2.5, fig.height=2}
p.umap.patient <- DimPlot(srat,
  reduction = "umap", group.by = "Patient", raster = TRUE
) +
  scale_color_manual(values = patient_color_maps) +
  labs(title = NULL) +
  my_theme
p.umap.patient

# saveRDS(p.umap.patient, file = here::here("output", DOCNAME, "Fig.umap.Patient.rds"))
ggsave2(p.umap.patient,
  filename = here::here("output", DOCNAME, "Fig.umap.Patient.pdf"),
  height = 2, width = 2.5, device = cairo_pdf
)
```

Split

```{r fig.width=7.5, fig.height=1.2}
p.umap.splitByPatient <- DimPlot(srat,
  reduction = "umap", split.by = "Patient",
  ncol = 8, raster = TRUE
) +
  scale_color_manual(values = cell_color_maps) +
  my_theme +
  NoAxes() +
  theme(
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.height = unit(.4, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
    plot.margin = margin(rep(0, 4), unit = "cm"),
    strip.background = element_rect(fill = "#f3f2f1")
  )
p.umap.splitByPatient

# saveRDS(p.umap.splitByPatient, file = here::here("output", DOCNAME, "Fig.umap.splitByPatient.rds"))
ggsave2(p.umap.splitByPatient,
  filename = here::here("output", DOCNAME, "Fig.umap.splitByPatient.pdf"),
  height = 1.2, width = 7.5, device = cairo_pdf
)
```

### VDJ

```{r fig.width=2.5, fig.height=2}
srat@meta.data <- srat@meta.data %>%
  rownames_to_column() %>%
  mutate(VDJ = case_when(
    VDJ != "TCR" ~ "None",
    TRUE ~ "TCR"
  )) %>%
  mutate(VDJ = factor(VDJ, levels = c("TCR", "None"))) %>%
  column_to_rownames()
p.umap.vdj <- DimPlot(srat,
  reduction = "umap", group.by = "VDJ", raster = TRUE
) +
  scale_color_manual(values = vdj_color_maps) +
  labs(title = NULL) +
  my_theme
p.umap.vdj

# saveRDS(p.umap.vdj, file = here::here("output", DOCNAME, "Fig.umap.VDJ.rds"))
ggsave2(p.umap.vdj,
  filename = here::here("output", DOCNAME, "Fig.umap.VDJ.pdf"),
  height = 2, width = 2.5, device = cairo_pdf
)
```

### Metastasis_n

```{r fig.width=2.5, fig.height=2}
p.umap.metastasis <- DimPlot(srat,
  reduction = "umap", group.by = "Metastasis_n", raster = TRUE
) +
  scale_color_manual(values = metastasis_color_maps) +
  labs(title = NULL) +
  my_theme
p.umap.metastasis

# saveRDS(p.umap.metastasis, file = here::here("output", DOCNAME, "Fig.umap.Metastasis_n.rds"))
ggsave2(p.umap.metastasis,
  filename = here::here("output", DOCNAME, "Fig.umap.Metastasis_n.pdf"),
  height = 2, width = 2.5, device = cairo_pdf
)
```

## Markers

DotPlot of markers

```{r plot-cell-markers-supp-dot, fig.width=7.5, fig.height=2}
p.marker.dot <- DotPlot(srat, features = c(
  # T
  "CD3D", "CD3G", "CD3E", "CD2",
  # B
  "CD19", "MS4A1", "CD79A", "IGHM",
  # mye
  "LYZ", "FCER1G", "TYROBP", "AIF1",
  # mast
  "TPSAB1", "TPSB2", "MS4A2", "KIT",
  # epi
  "CNFN", "KRT6A", "CRNN", "KRT13",
  # fib
  "DCN", "COL1A1", "COL1A2", "LUM",
  # endo
  "PECAM1", "CLDN5", "CDH5", "VWF",
  # platelets
  "PPBP", "PF4", "ITGA2B", "ITGB3"
), dot.scale = 4) +
  labs(x = NULL, y = NULL) +
  scale_y_discrete(limits = rev) +
  my_theme +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1)
  )
p.marker.dot

# saveRDS(p.marker.dot, file = here::here("output", DOCNAME, "Fig.dotplot.markers.rds"))

ggsave2(p.marker.dot,
  filename = here::here("output", DOCNAME, "Fig.dotplot.markers.pdf"),
  height = 2, width = 7.5, device = cairo_pdf
)
```

## Composition plot

### By Origin4 by cell type

```{r fig.width=2, fig.height=2}
p.byOrigin4.byCelltype <- srat@meta.data %>%
  group_by(Origin4, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100) %>%
  ggplot(aes(x = Origin4, y = Percent, fill = cellType)) +
  geom_bar(stat = "identity") +
  labs(x = NULL, fill = NULL, y = "% of cells") +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  my_theme +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm")
  )
p.byOrigin4.byCelltype
```

### By Origin by cell type

```{r fig.width=3, fig.height=2}
p.byOrigin.byCelltype <- srat@meta.data %>%
  group_by(Origin, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100) %>%
  ggplot(aes(x = Origin, y = Percent, fill = cellType)) +
  geom_bar(stat = "identity", color = "white") +
  labs(x = NULL, fill = NULL, y = "% of cells") +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  my_theme +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm")
  )
p.byOrigin.byCelltype
```

### cellType compositions

```{r cellType-compositions, fig.width=6, fig.height=2.5}
cluster_byOrigin4 <- srat@meta.data %>%
  group_by(cellType, Origin4) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100)
cluster_byPatient <- srat@meta.data %>%
  group_by(cellType, Patient) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100)
cluster_cellNum <- srat@meta.data %>%
  group_by(cellType) %>%
  summarise(n = n())

p1 <- ggplot(cluster_byOrigin4, aes(x = cellType, y = Percent, fill = Origin4)) +
  geom_bar(stat = "identity", color = "white") +
  scale_fill_manual(values = origin4_color_maps) +
  labs(fill = NULL, x = NULL) +
  coord_flip() +
  guides(fill = guide_legend(ncol = 1, byrow = TRUE)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_discrete(limits = rev) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  my_theme +
  theme(legend.position = "top")
p2 <- ggplot(cluster_byPatient, aes(x = cellType, y = Percent, fill = Patient)) +
  geom_bar(stat = "identity", color = "white") +
  scale_fill_manual(values = patient_color_maps) +
  labs(fill = NULL, x = NULL) +
  coord_flip() +
  guides(fill = guide_legend(ncol = 1, byrow = TRUE)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_discrete(limits = rev) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  my_theme +
  theme(legend.position = "top")
p3 <- ggplot(cluster_cellNum, aes(x = cellType, y = log2(n), fill = "blue")) +
  geom_bar(stat = "identity", color = "#00008f") +
  scale_fill_manual(values = "#8f8fb1") +
  labs(fill = NULL, x = NULL) +
  coord_flip() +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_discrete(limits = rev) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  my_theme +
  theme(legend.position = "top")

p.cellType.comp <- p1 + p2 + p3 +
  plot_layout(ncol = 3)
p.cellType.comp

ggsave2(p.cellType.comp,
  filename = here::here("output", DOCNAME, "Fig.cellTypeComposition.pdf"),
  height = 2.5, width = 6, device = cairo_pdf
)
```

### Sample composition

Source

```{r sample-compositions, fig.width=7.8, fig.height=3}
p <- PlotSampleFraction(seurat_obj = srat, "cellType", cell_color_maps)
p

# saveRDS(p, file = here::here("output", DOCNAME, "Fig.sampleComposition.rds"))
ggsave2(p,
  filename = here::here("output", DOCNAME, "Fig.sampleComposition.pdf"),
  height = 3, width = 7.8, device = cairo_pdf
)
```

## HC of Origin4

Heatmap by ComplexHeatmap

```{r plot-top-markers-complex, fig.height=2, fig.width=1.5, warning=FALSE, message=FALSE}
# Get average expression of variable features across tissue
# change identity to orig.ident
Idents(srat) <- "Origin4"

# get average expression across original ident
var_ave_byOrigin4 = AverageExpression(srat, assays = "RNA", features = VariableFeatures(srat), slot = "data")

# restore cell identity
Idents(srat) <- "cellType"

# scale
mat_scale <- pheatmap:::scale_rows(as.matrix(var_ave_byOrigin4$RNA))

# color
x <- ceiling(max(abs(mat_scale)))
col_fun <- circlize::colorRamp2(c(-x, 0, x), c("blue", "white", "red"))

ht <- Heatmap(mat_scale,
  name = "Exp.",
  col = col_fun,
  cluster_columns = TRUE, cluster_rows = TRUE,
  clustering_method_columns = "ward.D2",
  show_row_names = FALSE,
  column_names_gp = gpar(fontsize = 6),
  heatmap_legend_param = list( # direction = "horizontal",
    labels_gp = gpar(fontsize = 6),
    title_gp = gpar(fontsize = 6),
    # title_position = "leftcenter",
    grid_height = unit(2, "mm"),
    grid_width = unit(2, "mm")
  ),
  column_dend_height = unit(2, "mm"),
  show_row_dend = FALSE
)
draw(ht, heatmap_legend_side = "right")
p.Origin4.heat <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))

# save
ggsave2(p.Origin4.heat,
  filename = here::here("output", DOCNAME, "Fig.HC_Origin4.pdf"),
  height = 2, width = 1.5, device = cairo_pdf
)
```

## N3-N5

肿瘤组织内异质性

```{r n3-n5, fig.height=3, fig.width=2.5}
p.cellByN3N4N5 <- srat@meta.data %>%
  filter(Tissue == "Tumor") %>%
  group_by(Patient, Origin, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100) %>%
  ggplot(aes(x = Origin, y = Percent, fill = cellType)) +
  geom_col() +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(. ~ Patient) +
  my_theme +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.height = unit(.4, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
    plot.margin = margin(rep(0, 4), unit = "cm"),
    strip.background = element_rect(fill = "#f3f2f1")
  )
p.cellByN3N4N5
```

## Metastatic

转移阳性的病人与阴性病人

```{r Metastatic, fig.height=3, fig.width=3}
p.cellByMetastatic <- srat@meta.data %>%
  group_by(Origin4, Metastatic, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100) %>%
  ggplot(aes(x = Metastatic, y = Percent, fill = cellType)) +
  geom_col() +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(. ~ Origin4, ncol = 4) +
  my_theme +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.height = unit(.4, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
    plot.margin = margin(rep(0, 4), unit = "cm"),
    strip.background = element_rect(fill = "#f3f2f1")
  )
p.cellByMetastatic
```

## LN Drainage

引流方向

```{r LN-drainage-by-Metastasis-p, fig.height=3, fig.width=2}
p.cellByDrainage <- srat@meta.data %>%
  filter(Tissue == "LN") %>%
  group_by(Metastasis_n, Drainage, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100) %>%
  ggplot(aes(x = Drainage, y = Percent, fill = cellType)) +
  geom_col() +
  # geom_bar(stat="identity", position="dodge", width=0.8) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, fill = NULL) +
  facet_wrap(. ~ Metastasis_n) +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.cellByDrainage
```

## Align plots

### Main

Expression similarity of Origin4

```{r fig.width=8, fig.height=1.5}
lapply(levels(srat$Origin4), function(t) {
  tissue_origin <- readRDS(here::here("output/exp_similarity.cross_tissue.Origin4", paste0(t, ".tissue_origin.rds")))
  tissue_origin <- tissue_origin[!str_detect(tissue_origin$NN_type, "Epi_"), ]
  plot_sum <- t(as.data.frame(prop.table(table(tissue_origin$NN_tissue)) * 100))
  plot_sum <- as.data.frame(t(plot_sum))
  colnames(plot_sum) <- c("Tissue", "Percent")
  plot_sum$Percent <- as.numeric(as.character(plot_sum$Percent))
  plot_sum$Tissue <- factor(plot_sum$Tissue, levels = c(
    "prePBMC", "postPBMC", "nLN", "m_nLN", "m_pLN",
    "Normal", "Adjacent", "Tumor"
  ))
  p <- ggplot(data = plot_sum, aes(x = Tissue, y = Percent)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    scale_y_continuous(expand = c(0, 0)) +
    labs(x = NULL, y = NULL, title = t) +
    my_theme +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
  p
}) %>%
  wrap_plots(nrow = 1)

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.main.exp_similarity.pdf"),
  width = 8, height = 1.5, device = cairo_pdf
)
```

UMAP of cellType, UMAP of Origin4, Origin4 by cellType, UMAP of Patient

```{r fig.width=7.8, fig.height=1.8}
p.row2.Origin4 <- plot_grid(p.umap.cellType, p.umap.origin4, p.byOrigin4.byCelltype, p.umap.patient,
  rel_widths = c(1.9, 2.6, 1.9, 2.4), nrow = 1
)
p.row2.Origin4

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.main.Origin4_cellType.pdf"),
  height = 1.8, width = 7.8, device = cairo_pdf
)
```

UMAP of epi, fib, endo (epi as cluster)

```{r fig.width=7.5, fig.height=2.5}
srat.epi <- readRDS(here::here("output/04.rm_cells", "seurat_epithelia.rds"))
srat.endo <- readRDS(here::here("output/04.rm_cells", "seurat_endothelia.rds"))
srat.fib <- readRDS(here::here("output/04.rm_cells", "seurat_fibroblasts.rds"))

p.umap.endo <- DimPlot(srat.endo,
  reduction = "umap", group.by = "level_2",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE
) +
  scale_color_manual(values = endo_level_2_color) +
  labs(title = NULL) +
  my_theme +
  NoLegend()

p.umap.fib <- DimPlot(srat.fib,
  reduction = "umap", group.by = "level_3",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE
) +
  scale_color_manual(values = fib_level_3_color) +
  labs(title = NULL) +
  my_theme +
  NoLegend()

p.umap.epi <- DimPlot(srat.epi,
  reduction = "umap", group.by = "seurat_clusters",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE
) +
  scale_color_manual(values = cluster_color_maps) +
  labs(title = NULL) +
  my_theme +
  NoLegend()

p.umap.epi + p.umap.fib + p.umap.endo
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.main.umap_stromal.pdf"),
  height = 2.5, width = 7.5, device = cairo_pdf
)
```

UMAP of epi, fib, endo (epi as malignant)

```{r fig.width=7.6, fig.height=2.5}
p.umap.endo <- DimPlot(srat.endo,
  reduction = "umap", group.by = "level_2",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE
) +
  scale_color_manual(values = endo_level_2_color) +
  labs(title = NULL) +
  my_theme +
  NoLegend()

p.umap.fib <- DimPlot(srat.fib,
  reduction = "umap", group.by = "level_3",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE
) +
  scale_color_manual(values = fib_level_3_color) +
  labs(title = NULL) +
  my_theme +
  NoLegend()

p.umap.epi <- DimPlot(srat.epi,
  reduction = "umap", group.by = "malignant",
  label.size = geom.text.size,
  raster = TRUE
) +
  scale_color_manual(values = epi_malig_colors) +
  labs(title = NULL) +
  my_theme

p.umap.epi + p.umap.fib + p.umap.endo
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.main.main.umap_stromal.epi_malig.pdf"),
  height = 2.5, width = 7.6, device = cairo_pdf
)
```

### Supp

Row 1

```{r fig.width=7.5, fig.height=2}
ps.row1 <- p.umap.patient + p.umap.vdj + p.umap.metastasis +
  plot_layout(nrow = 1)
ps.row1

ggsave2(ps.row1,
  filename = here::here("output", DOCNAME, "Fig.supp.r1.pdf"),
  height = 2, width = 7.5, device = cairo_pdf
)
```

Row 2

```{r fig.width=7.5, fig.height=2}
p.marker.dot
```

Row 3

```{r fig.width=7.5, fig.height=1.2}
p.umap.splitByOrigin4
```

Row 4

```{r fig.width=7.5, fig.height=2.5}
ps.row4 <- p.Origin4.heat + p.cellType.comp +
  plot_layout(widths = c(1.5, 6))
ps.row4

ggsave2(ps.row4,
  filename = here::here("output", DOCNAME, "Fig.supp.r4.pdf"),
  height = 2.5, width = 7.5, device = cairo_pdf
)
```

Row 5

```{r fig.width=7.5, fig.height=3}
ps.row5 <- p.cellByN3N4N5 + p.cellByMetastatic + p.cellByDrainage +
  plot_layout(
    widths = c(3, 3, 1.5),
    guides = "collect"
  )
ps.row5

ggsave2(ps.row5,
  filename = here::here("output", DOCNAME, "Fig.supp.r5.pdf"),
  height = 3, width = 7.5, device = cairo_pdf
)
```

## Session info
