---
title: "00.Fig.cellchat"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.cellchat"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
theme_set(theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
))
my_theme <- theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# cellchat
library(CellChat)
library(NMF)
library(ggalluvial)

# Seurat
library(Seurat)

options(stringsAsFactors = FALSE)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

## Visually compare each pathway (cellType2)

Load cellchat rds

```{r load-rds, cache=FALSE, message=FALSE, warning=FALSE}
origin2_n = c("PBMC1", "PBMC2", "LN_N", "LN_P", "Normal", "Adjacent", "Tumor")
obj.lst = lapply(origin2_n, function(x){
  a = readRDS(here::here("output/CellChat.byOrigin2", paste0("cellchat.", paste0(x, ".rds"))))
  a = updateCellChat(a)
  a = subsetCellChat(a, idents.use = c("Unknown"), invert = TRUE)
  return(a)
})
names(obj.lst) = origin2_n
```

## plot

```{r vis-each-pathway, fig.width=21, fig.height=3}
pathway_to_show <- c("CD6", "ALCAM", "CCL", "SELL", "COLLAGEN")

pdf(here::here("output", DOCNAME, "Fig.netVisual_aggregate.selected_pathways.pdf"), height = 3, width = 21)
p_lst <- lapply(pathway_to_show, FUN = function(x){
  weight.max <- getMaxWeight(obj.lst, slot.name = 'netP', attribute = x)
  par(mfrow = c(1, 7), xpd=TRUE)
  for (i in 1:length(obj.lst)) {
    netVisual_aggregate(obj.lst[[i]], signaling = x, layout = 'circle',
                        edge.weight.max = weight.max[1], edge.width.max = 10,
                        signaling.name = paste(x, names(obj.lst)[i])
                        )
    }
})
invisible(dev.off())
```

## plot (test colors)

```{r vis-each-pathway, fig.width=21, fig.height=3}
pathway_to_show <- c("CD6", "ALCAM", "CCL", "SELL", "COLLAGEN")

pdf(here::here("output", DOCNAME, "Fig.netVisual_aggregate.selected_pathways-2.pdf"), height = 3, width = 21)
p_lst <- lapply(pathway_to_show, FUN = function(x){
  weight.max <- getMaxWeight(obj.lst, slot.name = 'netP', attribute = x)
  par(mfrow = c(1, 7), xpd=TRUE)
  for (i in 1:length(obj.lst)) {
    netVisual_aggregate(obj.lst[[i]], signaling = x, layout = 'circle',
                        edge.weight.max = weight.max[1], edge.width.max = 10,
                        signaling.name = paste(x, names(obj.lst)[i])
                        )
    }
})
invisible(dev.off())
```


## Compare the major sources and targets in 2D space (cellType3)

```{r}
origin2_n = c("PBMC1", "PBMC2", "LN_N", "LN_P", "Normal", "Adjacent", "Tumor")
obj.lst = lapply(origin2_n, function(x){
  a = readRDS(here::here(paste0("output/CellChat.byOrigin2.", paste0(x, ".cellType3/cellchat.rds"))))
  a = updateCellChat(a)
  #a = subsetCellChat(a, idents.use = c("Unknown"), invert = TRUE)
  return(a)
})
names(obj.lst) = origin2_n
```

Comparing the outgoing and incoming interaction strength in 2D space allows ready identification of the cell populations with significant changes in sending or receiving signals between different datasets.

```{r cmp-major-sour-target-in-2D, fig.width=12, fig.height=22, message=FALSE, warning=FALSE}
num.link <- sapply(obj.lst, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(unlist(num.link)), max(unlist(num.link))) # control the dot size in the different datasets
gg <- list()
for (i in 1:length(obj.lst)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(obj.lst[[i]], title = names(obj.lst)[i], weight.MinMax = weight.MinMax) +
  xlim(0, 210) +
  ylim(0, 250)
}
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
patchwork::wrap_plots(plots = gg, ncol = 2) +
  plot_layout(guides = "collect")

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.netAnalysis_signalingRole_scatter.pdf"),
  width = 12, height = 22
)
```












## Session info

