---
title: "00.Fig.epithelia"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.epithelia"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)
library(circlize)

# color
library(ggsci)

# Seurat
library(Seurat)
#library(scRNAtoolVis)

# enrichment
library(clusterProfiler)
library(org.Hs.eg.db)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

## Load data

Load srat

```{r load-rds, cache=FALSE}
srat <- readRDS(here::here("output/04.rm_cells/seurat_epithelia.rds"))
Idents(srat) <- "seurat_clusters"
srat
```

Load metadata

```{r load-meta, cache=FALSE, message=FALSE, warning=FALSE}
metadata <- read_csv(here::here("output/04.rm_cells/seurat_epithelia.cellmeta.csv")) %>%
    mutate(cellType3 = level_1)
sam_info <- read_csv(here::here("data/sample_info.3.csv"))
```

Load malignant percent of each cluster

```{r load-malignant, cache=FALSE, message=FALSE, warning=FALSE}
cluster_malig <- readRDS(here::here('output', "epithelia.cor_bulk", 'epi.cor_with_bulk.rds')) %>%
  group_by(seurat_clusters, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ungroup() %>%
  complete(seurat_clusters, malignant, fill = list(n = 0, Percent=0)) %>%
  filter(malignant == "Yes") %>%
  arrange(Percent)
```

## UMAP {.tabset}

### Clusters

```{r fig.width=3, fig.height=3, cache=FALSE}
p.umap.cellType <- DimPlot(srat,
  reduction = "umap", group.by = "seurat_clusters",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE
) +
  scale_color_manual(values = cluster_color_maps) +
  labs(title = NULL) +
  my_theme +
  NoLegend()
p.umap.cellType

# saveRDS(p.umap.cellType, file = here::here("output", DOCNAME, "Fig.umap.cellType.rds"))
ggsave2(p.umap.cellType,
  filename = here::here("output", DOCNAME, "Fig.umap.cellType.pdf"),
  height = 3, width = 3, device = cairo_pdf
)
```

### malignant

```{r fig.width=2.5, fig.height=2, cache=FALSE}
p.umap.malignant <- DimPlot(srat,
  reduction = "umap", group.by = "malignant", raster = TRUE
) +
  scale_color_manual(values = epi_malig_colors) +
  labs(title = NULL) +
  my_theme
p.umap.malignant

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.umap.malignant.pdf"),
  height = 2, width = 2.5, device=cairo_pdf
)
```

## Malig./Non-malig.

By cluster by malignant

```{r fig.width=4.5, fig.height=2.6, cache=FALSE}
p.byClusterd.byMalignant = readRDS(here::here('output', "epithelia.cor_bulk", 'epi.cor_with_bulk.rds')) %>%
  group_by(seurat_clusters, malignant) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ungroup() %>%
  complete(seurat_clusters, malignant, fill = list(n = 0, Percent=0)) %>%
  mutate(seurat_clusters = factor(seurat_clusters, levels = cluster_malig$seurat_clusters)) %>%
  ggplot(aes(x = seurat_clusters, y = Percent, fill = malignant)) +
  geom_col() +
  scale_fill_manual(values = colors_2) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(fill = NULL, y = "% of cells", x = "clusters") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  geom_hline(yintercept = c(3, 70), linetype = 'dashed') +
  my_theme +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm")
  )
p.byClusterd.byMalignant

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.malig_percent.pdf"),
  height = 2.6, width = 4.5, device=cairo_pdf
)
```

### Comparison all groups

#### DEG

```{r load-deg, message=FALSE, warning=FALSE}
deg <- read_csv(here::here('output', "06.epi.cmp_malig", 'epi.cmp_malig.markers.csv')) %>%
  dplyr::rename(avg_log2FC = avg_logFC) %>%
  mutate(cluster = factor(cluster, levels = c("NonMalig.", "PreMalig.", "Malig.")))
```

Plot Volcano

```{r, fig.height=3, fig.width=3}
p.volcano = jjVolcano(diffData = deg, cluster_size = 3, topGeneN = 5,
          pSize = .4, base_size = 6, size = geom.text.size,
          tile.col = epi_malig_colors,
          legend.position = c(.2, .1)) +
  labs(x = NULL)
p.volcano

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.deg_malig.volcano.pdf"),
  height = 3, width = 3, device = cairo_pdf
)
```

#### GO {.tabset}

```{r enrich, message=FALSE, warning=FALSE, fig.width=3, fig.height=3}
cg.BP = readRDS(here::here('output', "06.epi.cmp_malig", "epi.cmp_malig.GO-BP.rds"))

p.enrich_bp = dotplot(cg.BP) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 40)) + 
  scale_size(range = c(0, 4)) +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.enrich_bp

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.deg_malig.BP.pdf"),
  height = 3, width = 3, device = cairo_pdf
)
```

### TFs

All TFs tested

```{r load-de-tf-glm, message=FALSE, warning=FALSE}
# load
de_tf.glm_patient = readRDS(here::here("output", "06.epi.cmp_malig", "epi.cmp_malig.TFs.glm_patient.rds"))
```

Plot

```{r fig.width=2, fig.height=3}
# t_mat
t_mat = de_tf.glm_patient$t_mat %>%
  as.data.frame()

# select rows to show
selected_row = de_tf.glm_patient$df %>%
  filter(p_adj < 0.05) %>% # keep only those with p_adj < 0.05
  group_by(cluster) %>%
  top_n(5, wt = t) %>%
  pull(gene) %>% unique()

# add RORA/RORC/BARX2, SOX4
selected_row = c("RORA", "RORC", "BARX2", "SOX4", selected_row)

# highligh
highlight_rows = "SOX4"

plot_heatmap_glm(as.matrix(t_mat[selected_row,]), cluster_columns = FALSE,
                 highlight_rows = highlight_rows)

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.deg_malig.TFs.pdf"),
  height = 3, width = 2, device = cairo_pdf
)
```



## Selected genes

### markers

ward.D2 linkage

```{r fig.width=4.5, fig.height=3}
markers <- c(
  # lowDiff
  "EPCAM", "KRT5", "KRT14", "KRT15",
  # highDiff
  "KRT4", "KRT13", "ECM1", "SPRR3", "CNFN", "CRNN",
  # basal layer/tumor cell
  "TP63", "SOX2", "ITGA6", "SCPEP1",
  # TF
  "BARX2", "RORA", "RORC",
  # prolif
  "MKI67"
)

mat <- AverageExpression(srat,
  assays = "RNA", features = markers,
  slot = "data", verbose = F
)$RNA
mat <- log1p(as.matrix(mat))
# scale
mat_scale <- pheatmap:::scale_rows(mat)

# color
x <- floor(max(abs(mat_scale)))
col_fun <- circlize::colorRamp2(c(-x, 0, x), c("#476fa9", "#ffffff", "#ca3226"))
legend_name <- "Exp."

ht <- ComplexHeatmap::Heatmap(mat_scale,
  name = legend_name,
  col = col_fun,
  cluster_columns = TRUE,
  clustering_method_columns = "ward.D2",
  cluster_rows = TRUE,
  clustering_method_rows = "ward.D2",
  #column_dend_height = unit(10, "mm"),
  show_row_dend = FALSE,
  show_row_names = TRUE, show_column_names = TRUE,
  row_names_side = "right",
  row_names_gp = gpar(fontsize = 6, fontface = "italic"),
  column_names_gp = gpar(fontsize = 6),
  column_names_rot = 60,
  column_title = NULL,
  heatmap_legend_param = list(
    title_position = "topcenter",
    direction = "vertical",
    labels_gp = gpar(fontsize = 6),
    title_gp = gpar(fontsize = 6),
    grid_height = unit(1.5, "mm"),
    grid_width = unit(1.5, "mm")
  )
)
# draw(ht)
p.markers.heatmap <- plot_grid(grid.grabExpr(draw(ht)))
p.markers.heatmap

# save
saveRDS(p.markers.heatmap, file = here::here("output", DOCNAME, "Fig.markers.heatmap.rds"))

ggsave2(p.markers.heatmap,
  filename = here::here("output", DOCNAME, "Fig.markers.heatmap.pdf"),
  height = 3, width = 4.5, device = cairo_pdf
)
```

### SOX2/TP63

```{r fig.width=2, fig.height=3}
p.dot_SOX2 = DotPlot(srat, features = c("SOX2", "TP63", "KRT14", "KRT5", "KRT4", "KRT13", "IVL"),
        group.by = "malignant") +
  labs(x = NULL, y = NULL) +
  coord_flip() +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.dot_SOX2

ggsave2(p.dot_SOX2,
  filename = here::here("output", DOCNAME, "Fig.markers.dotplot_SOX2.pdf"),
  height = 3, width = 2, device = cairo_pdf
)
```

## monocle2

Load TF activity

```{r}
aucell_res_path = '../../pySCENIC/200227_6samples'

# SCENIC TFs
aucell_tf = read_csv(file.path(aucell_res_path, 'seurat_epithelia.TF.auc_mtx.csv'))
aucell_tf = aucell_tf %>% 
  mutate(Regulon = str_replace(Regulon, "[(][+][)]", "")) %>% 
  filter(Regulon %in% c("RORA", "RORC", "BARX2", "SOX4")) %>% 
  column_to_rownames(var = 'Regulon')

# t
aucell_tf = as.data.frame(t(aucell_tf)) %>%
  rownames_to_column("barcode")
```

Load cds directly

```{r load-rds-random1, cache=FALSE}
# monocle
library(monocle)
cds1 = readRDS(here::here('output/04.rm_cells/monocle2_epithelia.random1.rds'))
srat1 = readRDS(here::here('output/04.rm_cells/seurat_epithelia.random1.rds'))

cds1
```

Malignant

```{r plot-Malignant_1-random1, fig.height=2.5, fig.width=2.5}
p.monocle2_malig = plot_cell_trajectory(cds1, color_by = "malignant", cell_size = 2) +
  scale_color_manual(values = epi_malig_colors) +
  my_theme
p.monocle2_malig

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.monocle2_malig.pdf"),
  height = 2.5, width = 2.5, device = cairo_pdf
)
```

Gene expression

```{r gene-exp-random1, fig.height=1.6, fig.width=5}
features = c("RORA", "RORC", "BARX2")

p.monocle2_gene = FeaturePlot(srat1, features = features, reduction = "dt", ncol = 3) & my_theme
p.monocle2_gene

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.monocle2_RORA_exp.pdf"),
  height = 1.6, width = 5, device = cairo_pdf
)
```

TF activity

```{r tf-exp-random1, fig.height=1.5, fig.width=5}
dat = FetchData(srat1, vars = c("DT_1", "DT_2")) %>%
  rownames_to_column("barcode") %>%
  left_join(aucell_tf, by = "barcode")

p1 = dat %>%
  ggplot(aes(x = DT_1, y = DT_2, color = RORA)) +
  geom_point(size = .5) +
  my_theme
p2 = dat %>%
  ggplot(aes(x = DT_1, y = DT_2, color = RORC)) +
  geom_point(size = .5) +
  my_theme
p3 = dat %>%
  ggplot(aes(x = DT_1, y = DT_2, color = BARX2)) +
  geom_point(size = .5) +
  my_theme
p.monocle2_tf = p1 + p2 + p3 +
  plot_layout(ncol = 3)
p.monocle2_tf

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.monocle2_RORA_TF.pdf"),
  height = 1.5, width = 5, device = cairo_pdf
)
```

## CytoTRACE

```{r fig.width=5, fig.height=1.6}
result = readRDS(here::here("output", "CytoTRACE.epithelia", "CytoTRACE.res.rds"))

d4p = tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(srat@meta.data %>% rownames_to_column("barcode"), by = 'barcode') 

p1 = d4p %>%
  filter(malignant == "NonMalig.") %>%
  ggplot(aes(x=reorder(seurat_clusters, -value, median), y=value, fill=seurat_clusters)) +
  geom_boxplot_jitter(outlier.jitter.width = 0.01, outlier.size = .5) +
  scale_fill_manual(values = cluster_color_maps) +
  labs(y= 'Predicted order', x = NULL) +
  my_theme +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 60, hjust = 1))
p2 = d4p %>%
  filter(malignant == "PreMalig.") %>%
  ggplot(aes(x=reorder(seurat_clusters, -value, median), y=value, fill=seurat_clusters)) +
  geom_boxplot_jitter(outlier.jitter.width = 0.01, outlier.size = .5) +
  scale_fill_manual(values = cluster_color_maps) +
  labs(y= NULL, x = NULL) +
  my_theme +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 60, hjust = 1))
p3 = d4p %>%
  filter(malignant == "Malig.") %>%
  ggplot(aes(x=reorder(seurat_clusters, -value, median), y=value, fill=seurat_clusters)) +
  geom_boxplot_jitter(outlier.jitter.width = 0.01, outlier.size = .5) +
  scale_fill_manual(values = cluster_color_maps) +
  labs(y= NULL, x = NULL) +
  my_theme +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 60, hjust = 1))

p.CytoTRACE = p1 + p2 + p3 +
  plot_layout(widths = c(9, 7, 18))
p.CytoTRACE

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.CytoTRACE.byMalig.pdf"),
  height = 1.6, width = 5, device = cairo_pdf
)
```


## Session info

