---
title: "00.Fig.myeloid.mac"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.myeloid.mac"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
#library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
theme_set(theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
))
my_theme <- theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# pvclust
library(pvclust)
library(dendextend)

# Seurat
library(Seurat)

# monocle
library(monocle)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
devtools::load_all(here::here("code/Rscir"))
```

Load rds

```{r load-rds, cache=FALSE}
srat <- readRDS(here::here("output/04.rm_cells/seurat_mac.rds"))
Idents(srat) <- "level_3"
srat
```

Load cell meta

```{r load-cell-meta, cache=FALSE, message=FALSE, warning=FALSE}
cell_meta <- read_csv(here::here("output/04.rm_cells/seurat_mac.cellmeta.csv")) %>%
  dplyr::select(barcode, Origin2_n, seurat_clusters, level_2, level_3)
head(cell_meta)

sam_info <- read_csv(here::here("data/sample_info.3.csv")) %>%
  mutate(cell.filtered = as.numeric(cell.filtered))
```

## Freq

### Fraction changes

```{r fig.width=2, fig.height=2}
df_fc <- srat@meta.data %>%
  group_by(level_3, Source) %>%
  summarise(n = n()) %>%
  mutate(n = as.numeric(n)) %>%
  left_join(sam_info, "Source") %>%
  mutate(value = (n / cell.filtered) * 100) %>%
  dplyr::select(Origin2_n, Source, level_3, value) %>% 
  group_by(Origin2_n, level_3) %>%
  summarise(Percent = mean(value)) %>% 
    mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "prePBMC", "postPBMC", "nLN", "pLN", "Normal", "Adjacent", "Tumor"
    ))
  )

p.frac_change <- df_fc %>% 
  ggplot(aes(x = Origin2_n, y = Percent, color = level_3, group = level_3)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = mye_level_3_color[as.character(unique(df_fc$level_3))]) +
  labs(x = NULL, color = NULL) +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.frac_change

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.freq_change_mac_level3.pdf"),
  width = 2, height = 2
)
```

### Roe Origin2_n

```{r Origin2_n-enrichment, fig.width=2.5, fig.height=2.5, cache=FALSE}
cluster_cellNum = srat@meta.data %>% 
  group_by(level_3) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum = srat@misc$sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 = srat@meta.data %>%
  group_by(level_3, Origin2_n) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp = setdiff(levels(cluster_byOrigin2$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x = tibble(level_3 = levels(cluster_cellNum$level_3), Origin2_n = t, o = 0)
    cluster_byOrigin2 = bind_rows(cluster_byOrigin2, x)
  }
}

r_oe = cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = 'Origin2_n') %>%
  left_join(cluster_cellNum, by = 'level_3') %>%
  mutate(e = (origin.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(level_3, Origin2_n, r) %>% 
  spread(Origin2_n, r, fill = 0) %>% 
  column_to_rownames("level_3")

# change order
r_oe = r_oe[rev(intersect(levels(srat$level_3), unique(srat$level_3))),]

# use ggplot for better control
p.roe = plot_origin2_enrich(r_oe)
p.roe

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.Roe_mac_level3.pdf"),
  width = 2.5, height = 2.5
)
```


## HC of level3

Heatmap by ComplexHeatmap

```{r plot-top-markers-complex, fig.height=2, fig.width=1.5, warning=FALSE, message=FALSE}
# get average expression across original ident
var_ave_byCluster = AverageExpression(srat, assays = "RNA", features = VariableFeatures(srat), slot = "data")

# scale
mat_scale <- pheatmap:::scale_rows(as.matrix(var_ave_byCluster$RNA))

# color
x <- ceiling(max(abs(mat_scale)))
col_fun <- circlize::colorRamp2(c(-x, 0, x), c("blue", "white", "red"))

ht <- Heatmap(mat_scale,
  name = "Exp.",
  col = col_fun,
  cluster_columns = TRUE, cluster_rows = TRUE,
  clustering_method_columns = "complete",
  show_row_names = FALSE,
  column_names_gp = gpar(fontsize = 6),
  heatmap_legend_param = list( # direction = "horizontal",
    labels_gp = gpar(fontsize = 6),
    title_gp = gpar(fontsize = 6),
    # title_position = "leftcenter",
    grid_height = unit(2, "mm"),
    grid_width = unit(2, "mm")
  ),
  column_dend_height = unit(2, "mm"),
  show_row_dend = FALSE
)
draw(ht, heatmap_legend_side = "right")
p <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p

# save
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.HC_Level3.pdf"),
  height = 2, width = 1.5, device = cairo_pdf
)
```

## Functional signatures

Signature from:

- Gonzalez, H., Mei, W., Robles, I., Hagerling, C., Allen, B. M., Hauge Okholm, T. L., Nanjaraj, A., Verbeek, T., Kalavacherla, S., van Gogh, M., Georgiou, S., Daras, M., Phillips, J. J., Spitzer, M. H., Roose, J. P., & Werb, Z. (2022). Cellular architecture of human brain metastases. Cell, 185(4), 729-745.e20. https://doi.org/10.1016/j.cell.2021.12.043 (Table S4)

- Leruste, A., Tosello, J., Ramos, R. N., Tauziède-Espariat, A., Brohard, S., Han, Z.-Y., Beccaria, K., Andrianteranagna, M., Caudana, P., Nikolic, J., Chauvin, C., Niborski, L. L., Manriquez, V., Richer, W., Masliah-Planchon, J., Grossetête-Lalami, S., Bohec, M., Lameiras, S., Baulande, S., … Bourdeaut, F. (2019). Clonally Expanded T Cells Reveal Immunogenicity of Rhabdoid Tumors. Cancer Cell, 36(6), 597-612.e8. https://doi.org/10.1016/j.ccell.2019.10.008 (Table S4)

### By level 3

```{r , fig.height=2.5, fig.width=3.5}
markers <- list(
  # M1
  M1 = filter_features(srat, c("CCR7", "IL2RA", "IL15RA", "IL7R", "CXCL11", "CCL19", "CXCL10", "CXCL9", "TNF", "CCL5", "CCL15", "IL12B", "IL15", "TNFSF10", "IL6", "CCL20", "NAMPT", "TYMP", "BCL2A1", "FAS", "BIRC3", "GADD45G", "XAF1", "SLC7A5", "SLCO5A1", "SLC2A6", "SLC31A2", "IDO1", "PLA1A", "OASL", "CHI3L2", "HSD11B1", "AK3", "SPHK1", "PFKFB3", "PSME2", "PFKP", "PSMB9", "PSMA2", "OAS2", "PTX3", "VCAN", "APOL3", "IGFBP4", "APOL1", "PDGFA", "EDN1", "APOL2", "INHBA", "APOL6", "HESX1", "IRF1", "ATF3", "IRF7")),
  # M2
  M2 = filter_features(srat, c("P2RY13", "LPAR6", "TGFBR2", "HRH1", "TLR5", "CD302", "MSR1", "CXCR4", "CLEC7A", "P2RY14", "CD209", "CLEC10A", "MS4A6A", "CD36", "MS4A4A", "MRC1", "IGF1", "CCL23", "CCL18", "CCL13", "SLCO2B1", "SLC4A7", "SLC38A6", "CTSC", "HEXB", "LIPA", "ADK", "HNMT", "TPST2", "CERK", "HS3ST2", "LTA4H", "CA2", "ALOX15", "HS3ST1", "TGFBI", "SELENOP", "CHN2", "FN1", "FGL2", "GAS7", "EGR2", "MAF")),
  # angiogenesis
  angiogenesis = filter_features(srat, c("CCL3", "VEGFA", "SND1", "MTDH", "ITGA5", "TNF", "CXCL3", "ANXA3", "HMGB1", "HIF1A", "SEMA4D", "LRG1", "CHIL1")),
  # TAM activation
  TAM_acti = filter_features(srat, c("APOE", "SPP1", "FN1", "CD276", "CD9", "TREM2", "LGALS3", "CD59", "C1QB", "CHIT1", "CCL18", "CD63", "MARCO", "C2", "CD81", "A2M", "CD84", "IFI27", "NRP2", "CD68")),
  # ECM remodeling
  ECM_remodel = filter_features(srat, c("ADAMDEC1", "CTSC", "CTSB", "RGCC", "CTSS", "CTSZ", "ADAM17", "ADAM10", "ADAM8", "CTSL", "CTSA", "CTSS", "LYZ", "MMP9", "MMP14")),
  # immunosuppression
  immunosuppression = filter_features(srat, c("HAVCR2", "FCGR2B", "IL4RA", "CD274", "HIF1A", "IL15")),
  # proinflammatory cytokines
  proinflammatory_cytokines = filter_features(srat, c("PTGS2", "CXCL2", "IL8", "CXCL5", "CXCL1", "IL1B", "IL1A", "CSF3", "IL6")),
  # FcR and complement
  FcR_complement = filter_features(srat, c("FCER1G", "FCGR2A", "FCGR3A", "C1QC", "C1QA", "C1QB", "C5AR1", "C3")),
  # antigenic presentation
  antigenic_presentation = filter_features(srat, c("CD1C", "CD1E", "CD74", "HLA-DQA1", "HLA-DQA2", "HLA-DQB1", "HLA-DQB2", "HLA-DPA1", "HLA-DPB1", "HLA-DRA", "HLA-DRB1", "HLA-DRB5")),
    # hypoxia
  Hypoxia = filter_features(srat, c("AL2", "ALDOA", "AM", "BCL2", "BCL2L1", "FOS", "JUN", "SRC", "CBP", "CCR6", "CEBPB", "ENOLAS", "FAS", "FASLG", "FKBP52", "GAPDH", "GLUT1", "GLUT3", "CSF2", "IFNG", "IFNB1", "IL13", "IL1R1", "IL2RA", "IL4", "IL5", "IL6R", "IL8", "IL10R", "IL15R", "ITGA6", "ITK", "JUN D", "CLB54", "MAP3K5", "MAP2K1", "MAP2K2", "MIF", "NFATC1", "NFKB2", "NFKBIE", "P53", "TGM6", "TNF", "TNFRSF1B", "TRADD", "TRAIL", "TRAP1", "VEGF"))
)

set.seed(1)
srat = AddModuleScore(srat, features = markers, name = 'cluster')
program_score = srat@meta.data %>%
  dplyr::select(level_3, starts_with('cluster')) %>%
  pivot_longer(starts_with("cluster"), names_to = "cluster", values_to = "value") %>%
  group_by(cluster, level_3) %>%
  summarise(m = mean(value)) %>%
  pivot_wider(names_from = "cluster", values_from = "m") %>%
  column_to_rownames("level_3")

program_score = program_score[,paste0("cluster", 1:length(markers))]
colnames(program_score) = names(markers)
program_score = t(program_score)

mat <- pheatmap:::scale_rows(program_score)

ht <- Heatmap(mat,
  name = "z-score",
  cluster_columns = TRUE,
  cluster_rows = TRUE,
  show_column_dend = F,
  show_row_dend = F,
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8),
  column_names_rot = 60,
  heatmap_legend_param = list( # direction = "horizontal",
    labels_gp = gpar(fontsize = 8),
    title_gp = gpar(fontsize = 8),
    # title_position = "leftcenter",
    grid_height = unit(2, "mm"),
    grid_width = unit(2, "mm")
  )
)

p <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.mac_functional_signature_level3.heatmap.pdf"),
  height = 2.5, width = 3.5, device=cairo_pdf
)
```

### By level 3 By Origin2

Use `AddModuleScore` to get program scores of each cell

```{r , fig.width=7.5, fig.height=3.5}
gs_lst <- list(
  # TCA cycle
  TCA_cycle = filter_features(srat, c("ACLY", "ACO1", "ACO2", "CS", "DLAT", "DLD", "DLST", "FH", "IDH1", "IDH2", "IDH3A", "IDH3B", "IDH3G", "MDH1", "MDH1B", "MDH2", "OGDH", "PC", "PCK1PCK2", "PDHA1", "PDHB", "SDHA", "SDHB", "SDHD", "SUCLA2", "SUCLG1", "SUCLG2")),
  # Glycolysis
  Glycolysis = filter_features(srat, c("ALDOA", "ALDOB", "ALDOC", "BPGM", "ENO1", "ENO2", "GALM", "GCK", "GPI", "HK2", "HK3", "PFKL", "PGAM2", "PGK1", "PGK2", "PGM1", "PGM2", "PGM3", "PKLR", "TPI1")),
  # Pentose Phosphate Pathway
  Pentose_phos = filter_features(srat, c("G6PD", "6PD", "PGLS", "PRPS1", "PRPS1L1", "PRPS2", "RBKS", "RPE", "RPIA", "TALDO1", "TKT")),
  # Glycogen Metabolism
  Glycogen_metabo = filter_features(srat, c("GBE1", "GYS1", "GYS2", "UGP2", "AGL", "PGM1", "PGM2", "PGM3", "PYGL", "PYGM", "GSK3A", "GSK3B", "PHKA1", "PHKB", "PHKG1", "PHKG2")),
  # Lipid mediators
  Lipid_mediators = filter_features(srat, c("ADIPOQ", "CD137", "AgRP", "ANGPT1", "ANGPT2", "ANGPTL3", "ANGPTL4", "ANGPTL6", "BAFF", "BMP4", "CTSD", "CTSS", "CCL2", "CCL5", "RARRES2", "CFD", "CRP", "CXCL8", "CD26", "ESM1", "S100A12", "AHSG", "FGF1", "FGF2", "FGF21", "FGF23", "LGALS3", "GCG", "GH1", "HGF", "ICAM1", "IGFBP1", "IGFBP2", "IGFBP3", "IGFBP4", "IGFBP6", "IGFBP7", "IL1b", "IL6", "IL10", "IL11", "INS", "LEP", "LIF", "NGAL", "MCSF", "MIF", "MPO", "GHRL", "GHSR", "LIPE", "LPL", "AGT", "CEBPA", "CFD", "FABP4", "FASN", "IRS2", "KLF15", "PPARG", "PPARGC1A", "RETN", "SRT3", "SLC2A4", "SREBF1", "ACACB", "AXIN1", "CCND1", "CDK4", "CEBPB", "CEBPD", "DKK1", "E2F1", "FABP4", "FASN")),
  # Glucose Deprivation
  Glucose_deprivation = filter_features(srat, c("NRN1", "FAM129A", "1L23A", "SPRY1", "GADD45B", "HSPA1A/B", "GPNMB", "HAVCR2", "SOAT2", "ADAMTS6")),
  # G1/S
  G1S = filter_features(srat, c("BRCA1", "BRCA2", "CCNE1", "CCNE2", "CCNG2", "CDC25A", "CDC45", "CDC6", "CDKN1A", "CDKN2C", "CDKN3", "DHFR", "E2F1", "E2F5", "H1F0", "H1FNT", "H1FOO", "H1FX", "H2AFB1", "H2AFB2", "H2AFB3", "H2AFJ", "H2AFV", "H2AFVP1", "H2AFX", "H2AFY", "H2AFY2", "H2AFZ", "H2AFZP1", "H2AFZP2", "H2AFZP3", "H2AFZP4", "H2AFZP5", "H2AFZP6", "H2BFM", "H2BFS", "H2BFWT", "H2BFXP", "H3F3A", "H3F3AP1", "H3F3AP2", "H3F3B", "H3F3C", "HIST1H1A", "HIST1H1B", "HIST1H1C", "HIST1H1D", "HIST1H1E", "HIST1H1PS1", "HIST1H1PS2", "HIST1H1T", "HIST1H2AA", "HIST1H2AB", "HIST1H2AC", "HIST1H2AD", "HIST1H2AE", "HIST1H2AG", "HIST1H2AH", "HIST1H2AI", "HIST1H2AJ", "HIST1H2AK", "HIST1H2AL", "HIST1H2AM", "HIST1H2APS1", "HIST1H2APS2", "HIST1H2APS3", "HIST1H2APS4", "HIST1H2APS5", "HIST1H2BA", "HIST1H2BB", "HIST1H2BC", "HIST1H2BD", "HIST1H2BE", "HIST1H2BF", "HIST1H2BG", "HIST1H2BH", "HIST1H2BI", "HIST1H2BJ", "HIST1H2BK", "HIST1H2BL", "HIST1H2BM", "HIST1H2BN", "HIST1H2BO", "HIST1H2BPS1", "HIST1H2BPS2", "HIST1H3A", "HIST1H3B", "HIST1H3C", "HIST1H3D", "HIST1H3E", "HIST1H3F", "HIST1H3G", "HIST1H3H", "HIST1H3I", "HIST1H3J", "HIST1H3PS1", "HIST1H4A", "HIST1H4B", "HIST1H4C", "HIST1H4D", "HIST1H4E", "HIST1H4F", "HIST1H4G", "HIST1H4H", "HIST1H4I", "HIST1H4J", "HIST1H4K", "HIST1H4L", "HIST1H4PS1", "HIST2H2AA3", "HIST2H2AA4", "HIST2H2AB", "HIST2H2AC", "HIST2H2BA", "HIST2H2BB", "HIST2H2BC", "HIST2H2BD", "HIST2H2BE", "HIST2H2BF", "HIST2H3A", "HIST2H3C", "HIST2H3D", "HIST2H3PS2", "HIST2H4A", "HIST2H4B", "HIST3H2A", "HIST3H2BA", "HIST3H2BB", "HIST3H3", "HIST4H4", "MCM2", "MCM6", "MSH2", "NASP", "NPAT", "PCNA", "RRM1", "RRM2", "SLBP", "TYMS")),
  # G2M
  G2M = filter_features(srat, c("AURKA", "BIRC5", "BUB1", "BUB1B", "CCNA2", "CCNB1", "CCNB2", "CCNF", "CDC20", "CDC25B", "CDC25C", "CDK1", "CDKN2D", "CENPA", "CENPF", "CKS2", "KIF20A", "PLK1", "RACGAP1", "TOP2A")),
  # hypoxia
  Hypoxia = filter_features(srat, c("AL2", "ALDOA", "AM", "BCL2", "BCL2L1", "FOS", "JUN", "SRC", "CBP", "CCR6", "CEBPB", "ENOLAS", "FAS", "FASLG", "FKBP52", "GAPDH", "GLUT1", "GLUT3", "CSF2", "IFNG", "IFNB1", "IL13", "IL1R1", "IL2RA", "IL4", "IL5", "IL6R", "IL8", "IL10R", "IL15R", "ITGA6", "ITK", "JUN D", "CLB54", "MAP3K5", "MAP2K1", "MAP2K2", "MIF", "NFATC1", "NFKB2", "NFKBIE", "P53", "TGM6", "TNF", "TNFRSF1B", "TRADD", "TRAIL", "TRAP1", "VEGF")),
  # type I interferon response
  Type1_interferon = filter_features(srat, c("IRF1", "IFIH1", "IFITM3", "DDX58", "IFI44L", "IFI6", "IFITM2", "NAMPT", "OASL", "RTP4", "TREX1", "ADAR", "FAM46C", "LY6E", "MCOLN2", "APOBEC3G", "IL15", "ISG15", "MX1", "TLR3")),
  # Type II interferon response
  Type2_interferon = filter_features(srat, c("IFNG", "CXCL9", "CXCL10", "STAT1", "CD274", "P21", "MYC", "SMAD7", "IRF1")),
  # pro-fibrotic signature
  proFibrotic = filter_features(srat, c("CCL22", "CSF1", "CHIT1", "FCMR", "SIGLEC15", "CTSK", "COL22A1", "CHI3L1", "SPP1", "SPARC", "MMP9", "MMP7", "GPC4", "TIMP3", "MATK", "LIPA", "PALLD", "FDX1", "LPL")),
  # pro-fibrotic/Phagocytosis
  phagocytosis = filter_features(srat, GOfuncR::get_anno_genes("GO:0006909") %>% pull(gene)),
  # M1
  M1 = filter_features(srat, c("CCR7", "IL2RA", "IL15RA", "IL7R", "CXCL11", "CCL19", "CXCL10", "CXCL9", "TNF", "CCL5", "CCL15", "IL12B", "IL15", "TNFSF10", "IL6", "CCL20", "NAMPT", "TYMP", "BCL2A1", "FAS", "BIRC3", "GADD45G", "XAF1", "SLC7A5", "SLCO5A1", "SLC2A6", "SLC31A2", "IDO1", "PLA1A", "OASL", "CHI3L2", "HSD11B1", "AK3", "SPHK1", "PFKFB3", "PSME2", "PFKP", "PSMB9", "PSMA2", "OAS2", "PTX3", "VCAN", "APOL3", "IGFBP4", "APOL1", "PDGFA", "EDN1", "APOL2", "INHBA", "APOL6", "HESX1", "IRF1", "ATF3", "IRF7")),
  # M2
  M2 = filter_features(srat, c("P2RY13", "LPAR6", "TGFBR2", "HRH1", "TLR5", "CD302", "MSR1", "CXCR4", "CLEC7A", "P2RY14", "CD209", "CLEC10A", "MS4A6A", "CD36", "MS4A4A", "MRC1", "IGF1", "CCL23", "CCL18", "CCL13", "SLCO2B1", "SLC4A7", "SLC38A6", "CTSC", "HEXB", "LIPA", "ADK", "HNMT", "TPST2", "CERK", "HS3ST2", "LTA4H", "CA2", "ALOX15", "HS3ST1", "TGFBI", "SELENOP", "CHN2", "FN1", "FGL2", "GAS7", "EGR2", "MAF")),
  # angiogenesis
  angiogenesis = filter_features(srat, c("CCL3", "VEGFA", "SND1", "MTDH", "ITGA5", "TNF", "CXCL3", "ANXA3", "HMGB1", "HIF1A", "SEMA4D", "LRG1", "CHIL1")),
  # TAM activation
  TAM_acti = filter_features(srat, c("APOE", "SPP1", "FN1", "CD276", "CD9", "TREM2", "LGALS3", "CD59", "C1QB", "CHIT1", "CCL18", "CD63", "MARCO", "C2", "CD81", "A2M", "CD84", "IFI27", "NRP2", "CD68")),
  # ECM remodeling
  ECM_remodel = filter_features(srat, c("ADAMDEC1", "CTSC", "CTSB", "RGCC", "CTSS", "CTSZ", "ADAM17", "ADAM10", "ADAM8", "CTSL", "CTSA", "CTSS", "LYZ", "MMP9", "MMP14")),
  # immunosuppression
  immunosuppression = filter_features(srat, c("HAVCR2", "FCGR2B", "IL4RA", "CD274", "HIF1A", "IL15")),
  # proinflammatory cytokines
  proinflammatory_cytokines = filter_features(srat, c("PTGS2", "CXCL2", "IL8", "CXCL5", "CXCL1", "IL1B", "IL1A", "CSF3", "IL6")),
  # FcR and complement
  FcR_complement = filter_features(srat, c("FCER1G", "FCGR2A", "FCGR3A", "C1QC", "C1QA", "C1QB", "C5AR1", "C3")),
  # antigenic presentation
  antigenic_presentation = filter_features(srat, c("CD1C", "CD1E", "CD74", "HLA-DQA1", "HLA-DQA2", "HLA-DQB1", "HLA-DQB2", "HLA-DPA1", "HLA-DPB1", "HLA-DRA", "HLA-DRB1", "HLA-DRB5"))
)

set.seed(1)
srat = AddModuleScore(srat, features = gs_lst, name = 'program')

mat_all = srat@meta.data %>%
  #rownames_to_column("barcode") %>%
  dplyr::select(level_3, Origin2_n, starts_with('program')) %>%
  pivot_longer(starts_with("program"), names_to = "program", values_to = "value") %>%
  group_by(level_3, Origin2_n, program) %>%
  summarise(m = mean(value)) %>%
  pivot_wider(names_from = "program", values_from = "m")
mat_all = droplevels(mat_all)

mat_lst = split(mat_all, mat_all$level_3)
mat_lst = lapply(mat_lst, function(x){
  m = x %>%
    ungroup() %>%
    dplyr::select(-level_3) %>%
    column_to_rownames("Origin2_n")
  m = m[,paste0("program", 1:length(gs_lst))]
  colnames(m) = names(gs_lst)
  m = t(m)
  pheatmap:::scale_rows(m)
})
#mat_lst <- mat_lst[c("Tn", "Teff", "Tex", "Tcm", "Tem", "Trm")]

# plot
value_range <- range(mat_lst)
p_lst <- lapply(1:length(mat_lst), function(i){
  mat <- mat_lst[i]
  df <- reshape2::melt(mat) %>% 
    mutate(
      Var2 = factor(Var2, levels = levels(srat$Origin2_n)),
      Var1 = factor(Var1, levels = rev(names(gs_lst)))
    )
  df %>% 
    ggplot(aes(x = Var2, y = Var1, fill = value)) +
    geom_tile() +
    scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red",
    limits = value_range
  ) +
    labs(title = names(mat_lst)[i]) +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    text = element_text(size = 6)
  )
})
p_lst[[length(p_lst)]] <- p_lst[[length(p_lst)]] +
  theme(axis.text.y = element_text())

p_lst %>% 
  wrap_plots(nrow = 1, guides = "collect", widths = c(7, 6, 6, 7, 4, 4, 4))

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.mac_functional_signature.heatmap.pdf"),
  height = 3.5, width = 7.5, device=cairo_pdf
)
```

## Selected genes

### 组织驻留

```{r , fig.height=2, fig.width=2}
p <- StackedVlnPlot(srat,
  features = c(
    "LYVE1", "MARCO", "FOLR2"
  ),
  color.use = mye_level_3_color,
  angle.x = 60
)
p

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.mac_exp_tissue_resident.pdf"),
  width = 2, height = 2
)
```

### Don't eat me

```{r , fig.height=2, fig.width=2}
p <- StackedVlnPlot(srat,
  features = c(
    "SIRPA", "LILRB1", "LILRB2"
  ),
  color.use = mye_level_3_color,
  angle.x = 60
)
p

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.mac_exp_dontEatMe.pdf"),
  width = 2, height = 2
)
```

## SOX4

```{r}
srat_mps <- readRDS(here::here('output/04.rm_cells/seurat_mps.rds'))
srat_mps
```


```{r fig.width=2, fig.height=1.8}
FeaturePlot(srat_mps, features = "SOX4", order = T, pt.size = .001) +
  labs(title = NULL) +
  my_theme

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.monocle3_genes.mono_mac.SOX4.pdf"),
  width = 2, height = 1.8
)
```


## Session info


