---
title: "00.Fig.tcells"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.tcells"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)
library(UpSetR)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(
  font_size = 6,# Overall font size
  rel_small = 6 / 6,# axis tick labels
  rel_tiny = 6 / 6,# caption
  rel_large = 6 / 6 # title
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# Seurat
library(Seurat)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load meta data

```{r load-meta, message=FALSE, warning=FALSE}
meta.data <- read_csv(here::here("output/04.rm_cells/seurat_tcells.cellmeta.csv"))
sam_info <- read_csv(here::here("data/sample_info.3.csv")) %>%
  mutate(cell.filtered = as.numeric(cell.filtered))
```

Load rds

```{r load-rds, cache=FALSE}
srat = readRDS(here::here('output/04.rm_cells/seurat_tcells.rds'))
Idents(srat) <- 'level_3'
srat
```

## UMAP {.tabset}

### Level 3

```{r fig.width=2.5, fig.height=3.5, cache=FALSE}
p.umap.celltype = DimPlot(srat, reduction = 'umap', group.by = 'level_3') + 
  scale_color_manual(values = t_level_3_color) +
  labs(title = NULL) +
  my_theme +
  guides(color = guide_legend(ncol = 4, byrow = TRUE)) +
  theme(legend.margin = margin(rep(0, 4), unit = "cm"),
        legend.spacing.y = unit(0, "cm"),
        legend.key.height = unit(.4, "cm"),
        legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
        legend.position = "bottom")
p.umap.celltype

#saveRDS(p.umap.celltype, file = here::here("output", DOCNAME, "Fig.umap.cellType.rds"))
ggsave2(p.umap.celltype, filename = here::here("output", DOCNAME, "Fig.umap.cellType.pdf"),
        height = 3.5, width = 2.5, device=cairo_pdf)
```

With label

```{r fig.width=2.5, fig.height=2.5, cache=FALSE}
p.umap.celltype = DimPlot(srat, reduction = 'umap', group.by = 'level_3',
                          label = TRUE, label.size = geom.text.size,
                          raster = TRUE) + 
  scale_color_manual(values = t_level_3_color) +
  labs(title = NULL) +
  my_theme +
  theme(legend.position = "none")
p.umap.celltype

ggsave2(filename = here::here("output", DOCNAME, "Fig.umap.cellType.labelled.pdf"),
        height = 2.5, width = 2.5, device=cairo_pdf)
```

## Classification markers

Heatmap

```{r fig.width=4.25, fig.height=4}
markers = list(
  Nai = c('CCR7', 'SELL', 'TCF7', 'LEF1'),
  Mem = c('CD44', 'LMNA', 'ANXA1', 'GPR183'),
  #Trm = c("CD69", "ITGAE", "IL2RB"),
  Cyto = c('ICOS', 'CD27', 'CD28', "CXCR5"),
  Check = c("HAVCR2", "LAG3", "PDCD1", "ENTPD1", "LAYN", "CTLA4"),
  Costi = c("RGS5", "MCAM", "PDGFRB", "COL4A1", "COL4A2"),
  Treg = c("FOXP3", "IL2RA"),
  Gdt = c("TRGV4", "TRAV3", "ZNF683", "TRGV2"),
  NKT = c("KLRB1", "NKG7")
)

mat = AverageExpression(srat, assays = "RNA", features = unlist(markers),
                        slot = "data", verbose = F)$RNA
mat = log1p(as.matrix(mat))
# scale
mat_scale = pheatmap:::scale_rows(mat)

# no Unknown
mat_scale = mat_scale[,-31]

# change order
mat_scale <- mat_scale[,c("CD4-C2-Tn", "CD4-C3-Tn", "CD4-C4-Tn", "CD4-C6-Tn", "CD4-C7-Tn", "CD4-C9-Tn", "CD4-C10-Tn", "CD4-C1-Tcm", "CD4-C5-Tcm", "CD4-C12-Tcm", "CD4-C11-Tem", "CD4-C8-Tfh", "Treg-C1", "Treg-C2", "Treg-C3", "CD8-C8-Tn", "CD8-C1-Tcm", "CD8-C9-Trm", "CD8-C3-Tem", "CD8-C4-Teff", "CD8-C5-Teff", "CD8-C6-Teff", "CD8-C10-Teff", "CD8-C2-Tex", "CD8-C7-Tex", "CD8-C11-Tex", "CD8-C12-Tex", "γδT-C1", "γδT-C2", "NK/NKT")]

# color
x = ceiling(max(abs(mat_scale)))
col_fun = circlize::colorRamp2(c(-x, 0, x), c("#476fa9", "#ffffff", "#ca3226"))
legend_name = 'Exp.'

ht <- ComplexHeatmap::Heatmap(mat_scale, name = legend_name,
                              col = col_fun,
                              cluster_columns = FALSE, cluster_rows = FALSE,
                              show_row_names = TRUE, show_column_names = TRUE,
                              row_names_side = 'right',
                              row_names_gp = gpar(fontsize = 6, fontface = "italic"),
                              column_names_gp = gpar(fontsize = 6),
                              column_names_rot = 60,
                              row_split = factor(
                                rep(names(markers), sapply(markers, length)),
                                levels = names(markers)
                              ),
                              row_title_gp = gpar(fontsize = 6),
                              column_split = factor(c(rep(c("CD4"), 12),
                                                      rep(c("Treg"), 3),
                                                      rep(c("CD8"), 12),
                                                      rep(c("γδT"), 2),
                                                      "NK/NKT"),
                                                    levels = c("CD4", "Treg", "CD8", "γδT", "NK/NKT")),
                              column_title = NULL,
                              left_annotation = rowAnnotation(empty = anno_empty(border = FALSE,
                                                                                 width = unit(1, "mm")), 
                                                              foo = anno_block(gp = gpar(fill = "gray",
                                                                                         col = "gray"),
                                                                               width = unit(0.5, "mm"))),
                              heatmap_legend_param = list(
                                title_position = "topcenter",
                                direction = 'vertical',
                                labels_gp = gpar(fontsize = 6),
                                title_gp = gpar(fontsize = 8),
                                grid_height = unit(3, "mm"),
                                grid_width = unit(3, "mm")
                              ))
#draw(ht)
p.classification_markers.heatmap = plot_grid(grid.grabExpr(draw(ht)))
p.classification_markers.heatmap

# save
saveRDS(p.classification_markers.heatmap, file = here::here("output", DOCNAME, "Fig.markers.heatmap.rds"))

ggsave2(p.classification_markers.heatmap,
  filename = here::here("output", DOCNAME, "Fig.markers.heatmap.pdf"),
  height = 4, width = 4.25, device=cairo_pdf
)
```

## Classification markers (small)

Heatmap

```{r fig.width=3, fig.height=2.5}
ht <- ComplexHeatmap::Heatmap(mat_scale, name = legend_name,
                              col = col_fun,
                              cluster_columns = FALSE, cluster_rows = FALSE,
                              show_row_names = TRUE, show_column_names = TRUE,
                              row_names_side = 'right',
                              row_names_gp = gpar(fontsize = 4, fontface = "italic"),
                              column_names_gp = gpar(fontsize = 4),
                              column_names_rot = 60,
                              row_split = factor(
                                rep(names(markers), sapply(markers, length)),
                                levels = names(markers)
                              ),
                              row_title_gp = gpar(fontsize = 4),
                              column_split = factor(c(rep(c("CD4"), 12),
                                                      rep(c("Treg"), 3),
                                                      rep(c("CD8"), 12),
                                                      rep(c("γδT"), 2),
                                                      "NK/NKT"),
                                                    levels = c("CD4", "Treg", "CD8", "γδT", "NK/NKT")),
                              column_title = NULL,
                              left_annotation = rowAnnotation(empty = anno_empty(border = FALSE,
                                                                                 width = unit(1, "mm")), 
                                                              foo = anno_block(gp = gpar(fill = "gray",
                                                                                         col = "gray"),
                                                                               width = unit(0.5, "mm"))),
                              heatmap_legend_param = list(
                                title_position = "topcenter",
                                direction = 'vertical',
                                labels_gp = gpar(fontsize = 4),
                                title_gp = gpar(fontsize = 6),
                                grid_height = unit(3, "mm"),
                                grid_width = unit(3, "mm")
                              ))
#draw(ht)
p.classification_markers.heatmap = plot_grid(grid.grabExpr(draw(ht)))
p.classification_markers.heatmap

# save
saveRDS(p.classification_markers.heatmap, file = here::here("output", DOCNAME, "Fig.markers.heatmap.small.rds"))

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.markers.heatmap.small.pdf"),
  height = 2.5, width = 3, device=cairo_pdf
)
```

## Fraction changes

```{r fig.width=4, fig.height=2}
# level2
df_fc_cd8 <- meta.data %>%
  filter(level_1 == "CD8") %>% 
  group_by(level_2, Source) %>%
  summarise(n = n()) %>%
  mutate(n = as.numeric(n)) %>%
  left_join(sam_info, "Source") %>%
  mutate(value = (n / cell.filtered) * 100) %>%
  dplyr::select(Origin2_n, Source, level_2, value) %>% 
  group_by(Origin2_n, level_2) %>%
  summarise(Percent = mean(value)) %>% 
    mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "prePBMC", "postPBMC", "nLN", "pLN", "Normal", "Adjacent", "Tumor"
    )),
    level_2 = factor(level_2, levels = c("Tn", "Tcm", "Tem", "Trm", "Teff", "Tex"))
  )

df_fc_cd4 <- meta.data %>%
  filter(level_1 %in% c("CD4", "Treg")) %>% 
  group_by(level_2, Source) %>%
  summarise(n = n()) %>%
  mutate(n = as.numeric(n)) %>%
  left_join(sam_info, "Source") %>%
  mutate(value = (n / cell.filtered) * 100) %>%
  dplyr::select(Origin2_n, Source, level_2, value) %>% 
  group_by(Origin2_n, level_2) %>%
  summarise(Percent = mean(value)) %>% 
    mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "prePBMC", "postPBMC", "nLN", "pLN", "Normal", "Adjacent", "Tumor"
    )),
    level_2 = factor(level_2, levels = c("Tn", "Tcm", "Tem", "Tfh", "Treg"))
  )

p1 <- df_fc_cd8 %>% 
  ggplot(aes(x = Origin2_n, y = Percent, color = level_2, group = level_2)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = t_level_2_color[as.character(unique(df_fc_cd8$level_2))]) +
  labs(x = NULL, color = NULL) +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p2 <- df_fc_cd4 %>% 
  ggplot(aes(x = Origin2_n, y = Percent, color = level_2, group = level_2)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = t_level_2_color[as.character(unique(df_fc_cd4$level_2))]) +
  labs(x = NULL, color = NULL) +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p1 + p2

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.freq_change_t_level2.pdf"),
  width = 3, height = 2
)
```

### Level 2

Neat

```{r}
tb.fc.level_2 <- meta.data %>%
  filter(level_1 != "Unknown") %>% 
  mutate(level_2_new = paste(level_1, level_2, sep = "-")) %>% 
  group_by(level_2_new, Source) %>%
  summarise(n = n()) %>%
  mutate(n = as.numeric(n)) %>%
  left_join(sam_info, "Source") %>%
  mutate(value = (n / cell.filtered) * 100) %>%
  # 贲门左1 和左2 都命名为贲门
  mutate(Origin = case_when(
    Origin %in% c("LN-LP1", "LN-LP2") ~ "LN-LP",
    TRUE ~ as.character(Origin)
  )) %>%
  mutate(Tissue = factor(Tissue, levels = c("PBMC", "LN", "Normal", "Adjacent", "Tumor"))) %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "prePBMC", "postPBMC", "nLN", "pLN",
      "Normal", "Adjacent", "Tumor"
    ))
  )
```

cmp

```{r}
group_2_p <- ggpubr::compare_means(value ~ Origin2_n, data = tb.fc.level_2, group.by = "level_2_new") %>% 
  filter(p < 0.05) %>% 
  pull(level_2_new) %>% 
  unique()
group_2_p
```

Plot

```{r fig.width=6, fig.height=5}
plst <- lapply(group_2_p[-1], function(x) {
  PlotFracChange(tb.fc.level_2 %>% filter(level_2_new == x),
    x = "Origin2_n",
    title = x
  ) +
    my_theme +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
})
plst %>%
  wrap_plots(ncol = 4, guides = "collect")

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.freq_change_t_level2.supp.pdf"),
  width = 6, height = 5
)
```

## Roe Origin4

```{r fig.width=2, fig.height=4, warning=FALSE}
p.roe.t = readRDS(here::here("output", "freq.enrichment_cellType3", "p.roe.t.rds"))
p.roe.t

ggsave2(p.roe.t, filename = here::here("output", DOCNAME, "Fig.roe_t.pdf"),
        height = 4, width = 2)
```

## VDJ

### Tissue-shared

Load data

```{r load-clone, cache=FALSE, message=FALSE, warning=FALSE}
metadata = read_csv(here::here("output/04.rm_cells/seurat_tcells.cellmeta.csv")) %>%
  filter(level_1 %in% c("CD4", "CD8", "Treg"))
clones = read_csv('/work/home/project/scESCA/cellranger/TCR/200121_6samples.clonotype_tcr.csv')
```

Merge metadata with clonotypes

```{r merge-meta-clonotypes}
clones = clones %>%
  filter(clonotype != 'None') %>%  # cells with alpha-beta pair
  mutate(clonotype_size = as.numeric(clonotype_size)) %>%
  mutate(clonal_expansion = case_when(
    clonotype_size == 1 ~ 'No',
    TRUE ~ 'Clonal'
  ))

# d_4_p
d_4_p = metadata %>%
  left_join(clones, by = 'barcode') %>%
  filter(!is.na(clonotype)) %>%
  mutate(seurat_clusters = as.factor(seurat_clusters)) %>%
  mutate(Tissue = factor(Tissue, levels = c("PBMC", "LN", "Normal", "Adjacent", "Tumor"))) %>%
  mutate(Origin = factor(Origin, levels = c("PBMC1", "PBMC2",
                                            'LN-LR', 'LN-RR', 'LN-UTP1', 'LN-UTP2', 'LN-ITP',
                                            'LN-LTP', 'LN-LP', 'LN-LP1', 'LN-LP2', 'LN-LG',
                                            "Normal", "Adjacent",
                                            "Tumor_core", "Tumor_invasion", "Tumor_middle"))) %>%
  mutate(Origin2_n = factor(Origin2_n, levels = c("prePBMC", "postPBMC", "nLN", "pLN", "Normal", "Adjacent", "Tumor")),
         Origin4 = factor(Origin4, levels = c("prePBMC", "postPBMC", "nLN", "m_nLN", "m_pLN", "Normal", "Adjacent", "Tumor")),
         Origin3 = factor(Origin3, levels = c("PBMC1", "PBMC2", "LN_Up", "LN_Down", "Normal", "Adjacent", "Tumor")),
         Metastatic = factor(Metastatic, levels = c("None", "N", "P")))
```

9047 clonotypes were tissue-shared (!!!! Origin2_n).

```{r tissue_shared, cache=FALSE}
tmp = d_4_p %>%
  group_by(clonotype, Origin2_n) %>%
  summarise(n = n()) %>%
  mutate(n = n_distinct(clonotype, Origin2_n)) %>%
  dplyr::select(clonotype, n) %>%
  mutate(tissue_shared = ifelse(n > 1, 'Yes', 'No')) %>%
  dplyr::select(clonotype, tissue_shared)
d_4_p = d_4_p %>%
  left_join(tmp, by = 'clonotype')
table(tmp$tissue_shared)
```

Comparison between fraction of clonal cells in each cluster (x axis) and percentage of cells with TCRs shared across tissues (y axis).

```{r clonal-inter-tissue-level_3, fig.height=3, fig.width=4}
a = d_4_p %>%
  group_by(level_3, clonal_expansion) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  filter(clonal_expansion == 'Clonal')
b = d_4_p %>%
  unite('z', tissue_shared, clonal_expansion, remove = FALSE) %>%
  group_by(level_3, z) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  filter(z == 'Yes_Clonal')

a %>%
  left_join(b, by = 'level_3') %>%
  ggplot(aes(x = Percent.x, y = Percent.y, label = level_3)) +
  geom_point(color = '#006485', size = 3) +
  geom_text_repel(max.overlaps = 20) +
  labs(x = 'Clonal cells in each cluster (%)',
       y = 'Inter-tissue clonal cells in each cluster (%)',
       title = 'Inter-tissue clonal cells') +
  my_theme

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.vdj_inter_tissue.pdf"),
  width = 4, height = 3
)
```

### overlap by Origin2

```{r ov-Origin2_n, fig.width=4, fig.height=4}
PlotRepOverlap(RepOverlap(d_4_p, "Origin2_n"))

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.VDJ_overlap_Origin2_n.pdf"),
  height = 4, width = 4, device = cairo_pdf
)
```

### C11-Tex

```{r fig.width=3, fig.height=2}
x <- RepOverlap(d_4_p, "level_3")["CD8-C11-Tex",] * 100
data.frame(cell = names(x), value = x) %>% 
  na.omit() %>% 
  ggplot(aes(x = reorder(cell, -value), y = value, fill = cell)) +
  geom_col() +
  scale_fill_manual(values = t_level_3_color) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = "% shared clonotypes from C11-Tex to others") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 60, hjust = 1))

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.VDJ_overlap_C11-Tex_to_others.pdf"),
  width = 3, height = 2, device = cairo_pdf
)
```

Upset plot

```{r fig.width=7, fig.height=3}
cell_groups <- c("CD8-C11-Tex", "Treg-C2", "CD8-C5-Teff", "Treg-C1", "CD4-C8-Tfh", "CD8-C7-Tex", "CD8-C2-Tex", "Treg-C3")

df <- d_4_p %>% 
  filter(level_3 %in% cell_groups)

x <- split(df$clonotype, df$level_3)
upset(fromList(x), nsets = 8, nintersects = NA,
      scale.intersections	= "log2",
      sets.bar.color = t_level_3_color[cell_groups],
      point.size = 1, line.size = 0.3,
      text.scale = 0.8,
      mb.ratio = c(0.5, 0.5),
      number.angles = -15)

pdf(file = here::here("output", DOCNAME, "Fig.VDJ_overlap_C11-Tex_to_others.upsetPlot.pdf"),
    width = 7, height = 3)
upset(fromList(x), nsets = 8, nintersects = NA,
      scale.intersections	= "log2",
      sets.bar.color = t_level_3_color[cell_groups],
      point.size = 1, line.size = 0.3,
      text.scale = 0.8,
      mb.ratio = c(0.5, 0.5),
      number.angles = -15)
invisible(dev.off())
```









## Session info

