---
title: "00.Fig.cell_cell_freq"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.cell_cell_freq"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
# library(ggpubr)
library(ggrepel)
library(ggrastr)
library(ggtree)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)
library(dendextend)

# color
library(ggsci)

# Seurat
library(Seurat)

# vis
# library(scRNAtoolVis)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
# source(here::here("code/plot.R"))
```

## HC of cell freq

Load metadata

```{r load-meta, cache=FALSE, message=FALSE, warning=FALSE}
metadata <- read_csv(here::here("output/04.rm_cells/seurat.cellmeta.csv"))
sam_info <- read_csv(here::here("data/sample_info.3.csv"))
```

Neat

```{r}
# n mat
mat <- metadata %>%
  group_by(cellType3, Origin2_n) %>%
  summarise(o = n()) %>%
  ungroup() %>%
  complete(cellType3, Origin2_n, fill = list(o = 0)) %>%
  mutate(o = as.numeric(o)) %>%
  pivot_wider(id_cols = "Origin2_n", names_from = "cellType3", values_from = "o") %>%
  column_to_rownames("Origin2_n")
```

HC and Plot

```{r , fig.height=14, fig.width=3}
cor_dat <- cor(mat, method = "spearman")
dist_dat <- as.dist(1 - cor_dat)

hc <- hclust(dist_dat, method = "average")

dend <- as.dendrogram(hc)
# rev
dend <- rev(dend)
labels_sorted <- rev(hc$labels[hc$order])

# label color
label_colors <- metadata %>%
  dplyr::select(cellType, cellType3, -barcode) %>%
  distinct() %>%
  mutate(color = cell_color_maps[cellType]) %>%
  column_to_rownames("cellType3")

# Set the plot margin: bottom, left, top & right
par(
  mar = c(0, 0, 0, 10) + 0.1,
  xpd = NA
) # allow content to go into outer margin
dend <- dend %>%
  set("labels_cex", .5) %>% # label size
  set("labels_col", value = label_colors[labels_sorted, ]$color) %>% # label color
  set("branches_lwd", 2) %>%
  set("branches_k_color", k = 5, value = c(
    "#999999", "#E69F00", "#56B4E9", "#009E73",
    "#F0E442", "#0072B2", "#D55E00", "#CC79A7"
  )) %>% # branch color
  raise.dendrogram(.6)
dend %>%
  plot(horiz = TRUE, axes = FALSE)
colored_bars(colors = label_colors[labels_sorted,]$color, horiz=TRUE,
             rowLabels = c("cell"), y_shift = .8, cex.rowLabels = 0.5)

# ggplot
ggd1 <- as.ggdend(dend)
p.dend <- ggplot(ggd1, horiz = TRUE) # reproducing the above plot in ggplot2 :)
# p.dend
```

## ggtree tree

```{r, fig.height=14, fig.width=3}
# Split the tree into 5 clusters
clusters <- cutree(hc, k = 5)

p.gr <- ggtree(hc) +
  # geom_tree() +
  geom_tiplab(size = 2)
p.gr
```

## Roe

```{r fig.width=2.4, fig.height=12, warning=FALSE}
cluster_cellNum <- metadata %>%
  group_by(cellType3) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  group_by(cellType3, Origin2_n) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o)) %>%
  ungroup() %>%
  complete(cellType3, Origin2_n, fill = list(o = 0))

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin2_n") %>%
  left_join(cluster_cellNum, by = "cellType3") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType3, Origin2_n, r) %>%
  spread(Origin2_n, r, fill = 0) %>%
  column_to_rownames("cellType3")

# data for plot
r_oe[r_oe > 10] <- 10
d4p <- r_oe %>%
  rownames_to_column("cellType3") %>%
  pivot_longer(cols = !cellType3, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "prePBMC", "postPBMC",
      "nLN", "pLN",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType3 = factor(cellType3, levels = labels_sorted)
  )

# cell Num
df.cell_num <- cluster_byOrigin2

# plot
p.roe <- ggplot(d4p, aes(x = Origin2_n, y = cellType3, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.cell_num,
    aes(x = Origin2_n, y = cellType3, size = o),
    shape = 1,
    inherit.aes = FALSE
  ) +
  # scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "Cells") +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "top",
    text = element_text(size = 6),
    panel.border = element_rect(colour = "black", fill = NA, size = .5)
  )
p.roe
```

```{r , fig.height=14, fig.width=4, eval=FALSE}
p.roe <- p.roe + aplot::ylim2(p.dend)

plot_grid(p.dend, p.roe, align = "h")
```


## Functional signature

Load rds

```{r load-rds, cache=FALSE, eval=FALSE}
srat <- readRDS(here::here("output/04.rm_cells/seurat.rds"))
Idents(srat) <- "cellType3"
srat
```

Signature from:

- Azizi E, Carr AJ, Plitas G, Cornish AE, Konopacki C, Prabhakaran S, Nainys J, Wu K, Kiseliovas V, Setty M, et al. 2018. Single-Cell Map of Diverse Immune Phenotypes in the Breast Tumor Microenvironment. Cell 0. https://www.cell.com/cell/abstract/S0092-8674(18)30723-2 (Accessed July 2, 2018).

- Leruste, A., Tosello, J., Ramos, R. N., Tauziède-Espariat, A., Brohard, S., Han, Z.-Y., Beccaria, K., Andrianteranagna, M., Caudana, P., Nikolic, J., Chauvin, C., Niborski, L. L., Manriquez, V., Richer, W., Masliah-Planchon, J., Grossetête-Lalami, S., Bohec, M., Lameiras, S., Baulande, S., … Bourdeaut, F. (2019). Clonally Expanded T Cells Reveal Immunogenicity of Rhabdoid Tumors. Cancer Cell, 36(6), 597-612.e8. https://doi.org/10.1016/j.ccell.2019.10.008 (Table S4)


```{r , eval=FALSE}
gene_signatures <- list(
  # TCA cycle
  TCA_cycle = filter_features(srat, c("ACLY", "ACO1", "ACO2", "CS", "DLAT", "DLD", "DLST", "FH", "IDH1", "IDH2", "IDH3A", "IDH3B", "IDH3G", "MDH1", "MDH1B", "MDH2", "OGDH", "PC", "PCK1PCK2", "PDHA1", "PDHB", "SDHA", "SDHB", "SDHD", "SUCLA2", "SUCLG1", "SUCLG2")),
  # Glycolysis
  Glycolysis = filter_features(srat, c("ALDOA", "ALDOB", "ALDOC", "BPGM", "ENO1", "ENO2", "GALM", "GCK", "GPI", "HK2", "HK3", "PFKL", "PGAM2", "PGK1", "PGK2", "PGM1", "PGM2", "PGM3", "PKLR", "TPI1")),
  # Pentose Phosphate Pathway
  Pentose_phos = filter_features(srat, c("G6PD", "6PD", "PGLS", "PRPS1", "PRPS1L1", "PRPS2", "RBKS", "RPE", "RPIA", "TALDO1", "TKT")),
  # Lipid mediators
  Lipid_mediators = filter_features(srat, c("ADIPOQ", "CD137", "AgRP", "ANGPT1", "ANGPT2", "ANGPTL3", "ANGPTL4", "ANGPTL6", "BAFF", "BMP4", "CTSD", "CTSS", "CCL2", "CCL5", "RARRES2", "CFD", "CRP", "CXCL8", "CD26", "ESM1", "S100A12", "AHSG", "FGF1", "FGF2", "FGF21", "FGF23", "LGALS3", "GCG", "GH1", "HGF", "ICAM1", "IGFBP1", "IGFBP2", "IGFBP3", "IGFBP4", "IGFBP6", "IGFBP7", "IL1b", "IL6", "IL10", "IL11", "INS", "LEP", "LIF", "NGAL", "MCSF", "MIF", "MPO", "GHRL", "GHSR", "LIPE", "LPL", "AGT", "CEBPA", "CFD", "FABP4", "FASN", "IRS2", "KLF15", "PPARG", "PPARGC1A", "RETN", "SRT3", "SLC2A4", "SREBF1", "ACACB", "AXIN1", "CCND1", "CDK4", "CEBPB", "CEBPD", "DKK1", "E2F1", "FABP4", "FASN")),
  # G1/S
  G1S = filter_features(srat, c("BRCA1", "BRCA2", "CCNE1", "CCNE2", "CCNG2", "CDC25A", "CDC45", "CDC6", "CDKN1A", "CDKN2C", "CDKN3", "DHFR", "E2F1", "E2F5", "H1F0", "H1FNT", "H1FOO", "H1FX", "H2AFB1", "H2AFB2", "H2AFB3", "H2AFJ", "H2AFV", "H2AFVP1", "H2AFX", "H2AFY", "H2AFY2", "H2AFZ", "H2AFZP1", "H2AFZP2", "H2AFZP3", "H2AFZP4", "H2AFZP5", "H2AFZP6", "H2BFM", "H2BFS", "H2BFWT", "H2BFXP", "H3F3A", "H3F3AP1", "H3F3AP2", "H3F3B", "H3F3C", "HIST1H1A", "HIST1H1B", "HIST1H1C", "HIST1H1D", "HIST1H1E", "HIST1H1PS1", "HIST1H1PS2", "HIST1H1T", "HIST1H2AA", "HIST1H2AB", "HIST1H2AC", "HIST1H2AD", "HIST1H2AE", "HIST1H2AG", "HIST1H2AH", "HIST1H2AI", "HIST1H2AJ", "HIST1H2AK", "HIST1H2AL", "HIST1H2AM", "HIST1H2APS1", "HIST1H2APS2", "HIST1H2APS3", "HIST1H2APS4", "HIST1H2APS5", "HIST1H2BA", "HIST1H2BB", "HIST1H2BC", "HIST1H2BD", "HIST1H2BE", "HIST1H2BF", "HIST1H2BG", "HIST1H2BH", "HIST1H2BI", "HIST1H2BJ", "HIST1H2BK", "HIST1H2BL", "HIST1H2BM", "HIST1H2BN", "HIST1H2BO", "HIST1H2BPS1", "HIST1H2BPS2", "HIST1H3A", "HIST1H3B", "HIST1H3C", "HIST1H3D", "HIST1H3E", "HIST1H3F", "HIST1H3G", "HIST1H3H", "HIST1H3I", "HIST1H3J", "HIST1H3PS1", "HIST1H4A", "HIST1H4B", "HIST1H4C", "HIST1H4D", "HIST1H4E", "HIST1H4F", "HIST1H4G", "HIST1H4H", "HIST1H4I", "HIST1H4J", "HIST1H4K", "HIST1H4L", "HIST1H4PS1", "HIST2H2AA3", "HIST2H2AA4", "HIST2H2AB", "HIST2H2AC", "HIST2H2BA", "HIST2H2BB", "HIST2H2BC", "HIST2H2BD", "HIST2H2BE", "HIST2H2BF", "HIST2H3A", "HIST2H3C", "HIST2H3D", "HIST2H3PS2", "HIST2H4A", "HIST2H4B", "HIST3H2A", "HIST3H2BA", "HIST3H2BB", "HIST3H3", "HIST4H4", "MCM2", "MCM6", "MSH2", "NASP", "NPAT", "PCNA", "RRM1", "RRM2", "SLBP", "TYMS")),
  # G2M
  G2M = filter_features(srat, c("AURKA", "BIRC5", "BUB1", "BUB1B", "CCNA2", "CCNB1", "CCNB2", "CCNF", "CDC20", "CDC25B", "CDC25C", "CDK1", "CDKN2D", "CENPA", "CENPF", "CKS2", "KIF20A", "PLK1", "RACGAP1", "TOP2A")),
  # hypoxia
  Hypoxia = filter_features(srat, c("AL2", "ALDOA", "AM", "BCL2", "BCL2L1", "FOS", "JUN", "SRC", "CBP", "CCR6", "CEBPB", "ENOLAS", "FAS", "FASLG", "FKBP52", "GAPDH", "GLUT1", "GLUT3", "CSF2", "IFNG", "IFNB1", "IL13", "IL1R1", "IL2RA", "IL4", "IL5", "IL6R", "IL8", "IL10R", "IL15R", "ITGA6", "ITK", "JUN D", "CLB54", "MAP3K5", "MAP2K1", "MAP2K2", "MIF", "NFATC1", "NFKB2", "NFKBIE", "P53", "TGM6", "TNF", "TNFRSF1B", "TRADD", "TRAIL", "TRAP1", "VEGF")),
  # Pro inflammatory
  Proinflammatory = filter_features(srat, c("IL1A", "IL1B", "IFNG", "TBX21", "CCL3", "CCL4", "PRF1", "GZMA", "GZMB", "GZMK", "GZMH", "CD8A", "FASLG", "CCL2", "CCL20", "IL2", "IL6", "IL12A", "IL17A", "IL23A", "PTGS2", "TLR4")),
  # Anti-inflammatory
  Antiinflammatory = filter_features(srat, c("TIGIT", "IDO1", "LGALS3", "PDCD1", "FOXP3", "ENTPD1", "CD274", "CSF2", "CTLA4", "CXCL12", "CXCL5", "IL8", "MIF", "PTGS2", "VEGFA")),
  # type I interferon response
  Type1_interferon = filter_features(srat, c("IRF1", "IFIH1", "IFITM3", "DDX58", "IFI44L", "IFI6", "IFITM2", "NAMPT", "OASL", "RTP4", "TREX1", "ADAR", "FAM46C", "LY6E", "MCOLN2", "APOBEC3G", "IL15", "ISG15", "MX1", "TLR3")),
  # Type II interferon response
  Type2_interferon = filter_features(srat, c("IFNG", "CXCL9", "CXCL10", "STAT1", "CD274", "P21", "MYC", "SMAD7", "IRF1")),
  # angiogenesis
  angiogenesis = filter_features(srat, c("CCL3", "VEGFA", "SND1", "MTDH", "ITGA5", "TNF", "CXCL3", "ANXA3", "HMGB1", "HIF1A", "SEMA4D", "LRG1", "CHIL1")),
  # ECM remodeling
  ECM_remodel = filter_features(srat, c("ADAMDEC1", "CTSC", "CTSB", "RGCC", "CTSS", "CTSZ", "ADAM17", "ADAM10", "ADAM8", "CTSL", "CTSA", "CTSS", "LYZ", "MMP9", "MMP14")),
  # pro-fibrotic signature
  proFibrotic = filter_features(srat, c("CCL22", "CSF1", "CHIT1", "FCMR", "SIGLEC15", "CTSK", "COL22A1", "CHI3L1", "SPP1", "SPARC", "MMP9", "MMP7", "GPC4", "TIMP3", "MATK", "LIPA", "PALLD", "FDX1", "LPL"))
)

set.seed(1)
srat <- AddModuleScore(srat, features = gene_signatures, name = "program")
program_score <- srat@meta.data %>%
  dplyr::select(cellType3, starts_with("program")) %>%
  pivot_longer(starts_with("program"), names_to = "program", values_to = "value") %>%
  group_by(program, cellType3) %>%
  summarise(m = mean(value)) %>%
  pivot_wider(names_from = "program", values_from = "m") %>%
  column_to_rownames("cellType3")

program_score <- program_score[, paste0("program", 1:length(gene_signatures))]
colnames(program_score) <- names(gene_signatures)
program_score <- t(program_score)

saveRDS(program_score, file = here::here("output", DOCNAME, "mat_functaionalSignature.rds"))
```

```{r , fig.height=12, fig.width=3}
mat_functaionalSignature <- readRDS(here::here("output", DOCNAME, "mat_functaionalSignature.rds"))
# mat <- pheatmap:::scale_rows(mat_functaionalSignature)
mat <- mat_functaionalSignature

ht <- Heatmap(t(mat)[rev(labels_sorted), ],
  name = "z-score",
  cluster_columns = TRUE,
  cluster_rows = FALSE,
  show_column_dend = F,
  show_row_dend = F,
  row_names_gp = gpar(fontsize = 6),
  column_names_gp = gpar(fontsize = 6),
  column_names_rot = 60,
  heatmap_legend_param = list( # direction = "horizontal",
    labels_gp = gpar(fontsize = 6),
    title_gp = gpar(fontsize = 6),
    # title_position = "leftcenter",
    grid_height = unit(2, "mm"),
    grid_width = unit(2, "mm")
  )
)

p.pathway <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p.pathway
```

ggplot

```{r, fig.height=12, fig.width=3}
pg.pathway <- t(mat)[rev(labels_sorted), ] %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  pivot_longer(cols = TCA_cycle:proFibrotic, names_to = "name", values_to = "value") %>%
  mutate(
    rowname = factor(rowname, levels = labels_sorted),
    name = factor(name, levels = c(
      "TCA_cycle", "Glycolysis", "Pentose_phos", "Lipid_mediators", "G1S", "G2M", "Hypoxia", "Proinflammatory", "Antiinflammatory", "Type1_interferon", "Type2_interferon", "angiogenesis", "ECM_remodel", "proFibrotic"
    ))
  ) %>%
  ggplot(aes(x = name, y = rowname, fill = value)) +
  geom_tile() +
  scale_y_discrete(position = "right") +
  scale_fill_gradient(
    low = "white", high = "red"
  ) +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "top",
    text = element_text(size = 6),
    panel.border = element_rect(colour = "black", fill = NA, size = .5)
  )

pg.pathway
```

## Cytokine

```{r , eval=FALSE}
cytokine <- filter_features(srat, c(
  "CCL3", "CCL4", "CCL5", "CCL20", "CCL19", "CCL21", "CXCL8", "CXCL9", "CXCL10", "CXCL13", "TGFB1", "IL6", "IL10", "IFNG"
))

exp <- AverageExpression(srat, assays = "RNA", features = cytokine, slot = "data", verbose = FALSE)
mat_cytokine <- log1p(exp$RNA)

saveRDS(mat_cytokine, file = here::here("output", DOCNAME, "mat_cytokine.rds"))
```

```{r , fig.height=12, fig.width=3}
mat_cytokine <- readRDS(file = here::here("output", DOCNAME, "mat_cytokine.rds"))
mat <- mat_cytokine

ht <- Heatmap(t(mat)[rev(labels_sorted), ],
  name = "exp.",
  cluster_columns = TRUE,
  cluster_rows = FALSE,
  show_column_dend = F,
  show_row_dend = F,
  row_names_gp = gpar(fontsize = 6),
  column_names_gp = gpar(fontsize = 6),
  column_names_rot = 60,
  heatmap_legend_param = list( # direction = "horizontal",
    labels_gp = gpar(fontsize = 6),
    title_gp = gpar(fontsize = 6),
    # title_position = "leftcenter",
    grid_height = unit(2, "mm"),
    grid_width = unit(2, "mm")
  )
)

p.cytokine <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p.cytokine

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.cytokine.pdf"),
  width = 3, height = 12
)
```

ggplot

```{r, fig.height=12, fig.width=3}
mat_cytokine <- readRDS(file = here::here("output", DOCNAME, "mat_cytokine.rds"))
mat <- mat_cytokine

pg.cytokine <- t(mat)[rev(labels_sorted), ] %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  pivot_longer(cols = CCL3:IFNG, names_to = "name", values_to = "value") %>%
  mutate(
    rowname = factor(rowname, levels = labels_sorted),
    name = factor(name, levels = c(
      "CCL3", "CCL4", "CCL5", "CCL20", "CCL19", "CCL21", "CXCL8",
      "CXCL9", "CXCL10", "CXCL13", "TGFB1", "IL6", "IL10", "IFNG"
    ))
  ) %>%
  ggplot(aes(x = name, y = rowname, fill = value)) +
  geom_tile() +
  scale_y_discrete(position = "right") +
  scale_fill_gradient(
    low = "white", high = "red"
  ) +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "top",
    text = element_text(size = 6),
    panel.border = element_rect(colour = "black", fill = NA, size = .5)
  )

pg.cytokine
```

## Cytokine receptor

```{r , eval=FALSE}
cytokine_r <- filter_features(srat, c(
  "CCR5", "CCR6", "CCR7", "CXCR1", "CXCR3", "CXCR5", "TGFBR2", "IL6R", "IL6ST", "IL10R", "TNFRSF1A", "TNFRSF1B", "IFNGR1"
))

exp <- AverageExpression(srat, assays = "RNA", features = cytokine_r, slot = "data", verbose = FALSE)
mat_cytokine_r <- log1p(exp$RNA)

saveRDS(mat_cytokine_r, file = here::here("output", DOCNAME, "mat_cytokine_receptor.rds"))
```

```{r , fig.height=12, fig.width=3}
mat_cytokine_r <- readRDS(file = here::here("output", DOCNAME, "mat_cytokine_receptor.rds"))
mat <- mat_cytokine_r

ht <- Heatmap(t(mat)[rev(labels_sorted), ],
  name = "exp.",
  cluster_columns = TRUE,
  cluster_rows = FALSE,
  show_column_dend = F,
  show_row_dend = F,
  row_names_gp = gpar(fontsize = 6),
  column_names_gp = gpar(fontsize = 6),
  column_names_rot = 60,
  heatmap_legend_param = list( # direction = "horizontal",
    labels_gp = gpar(fontsize = 6),
    title_gp = gpar(fontsize = 6),
    # title_position = "leftcenter",
    grid_height = unit(2, "mm"),
    grid_width = unit(2, "mm")
  )
)

p.cytokine_r <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p.cytokine_r

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.cytokine_r.pdf"),
  width = 3, height = 12
)
```

ggplot

```{r, fig.height=12, fig.width=3}
pg.cytokine_r <- t(mat)[rev(labels_sorted), ] %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  pivot_longer(cols = CCR5:IFNGR1, names_to = "name", values_to = "value") %>%
  mutate(
    rowname = factor(rowname, levels = labels_sorted),
    name = factor(name, levels = c("CCR5", "CCR6", "CCR7", "CXCR1", "CXCR3", "CXCR5", "TGFBR2", "IL6R", "IL6ST", "IL10R", "TNFRSF1A", "TNFRSF1B", "IFNGR1"))
  ) %>%
  ggplot(aes(x = name, y = rowname, fill = value)) +
  geom_tile() +
  scale_y_discrete(position = "right") +
  scale_fill_gradient(
    low = "white", high = "red"
  ) +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "top",
    text = element_text(size = 6),
    panel.border = element_rect(colour = "black", fill = NA, size = .5)
  )

pg.cytokine_r
```

## Immunotherapeutic targets

```{r , eval=FALSE}
markers <- filter_features(srat, c(
  "PDCD1", "CD247", "CTLA4", "LAG3", "HAVCR2", "TNFSF4", "TNFRSF4", "BTLA", "CD47", "SIRPA", "LGALS9"
))

exp <- AverageExpression(srat, assays = "RNA", features = markers, slot = "data", verbose = FALSE)
mat_immunoTarget <- log1p(exp$RNA)

saveRDS(mat_immunoTarget, file = here::here("output", DOCNAME, "mat_immunoTarget.rds"))
```

```{r , fig.height=12, fig.width=3}
mat_immunoTarget <- readRDS(file = here::here("output", DOCNAME, "mat_immunoTarget.rds"))
mat <- mat_immunoTarget

ht <- Heatmap(t(mat)[rev(labels_sorted), ],
  name = "exp.",
  cluster_columns = TRUE,
  cluster_rows = FALSE,
  show_column_dend = F,
  show_row_dend = F,
  row_names_gp = gpar(fontsize = 6),
  column_names_gp = gpar(fontsize = 6),
  column_names_rot = 60,
  heatmap_legend_param = list( # direction = "horizontal",
    labels_gp = gpar(fontsize = 6),
    title_gp = gpar(fontsize = 6),
    # title_position = "leftcenter",
    grid_height = unit(2, "mm"),
    grid_width = unit(2, "mm")
  )
)

p.target <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p.target
```

ggplot

```{r, fig.height=12, fig.width=3}
pg.target <- t(mat)[rev(labels_sorted), ] %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  pivot_longer(cols = PDCD1:LGALS9, names_to = "name", values_to = "value") %>%
  mutate(
    rowname = factor(rowname, levels = labels_sorted),
    name = factor(name, levels = c(
      "PDCD1", "CD247", "CTLA4", "LAG3", "HAVCR2", "TNFSF4", "TNFRSF4", "BTLA", "CD47", "SIRPA", "LGALS9"
    ))
  ) %>%
  ggplot(aes(x = name, y = rowname, fill = value)) +
  geom_tile() +
  scale_y_discrete(position = "right") +
  scale_fill_gradient(
    low = "white", high = "red"
  ) +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "top",
    text = element_text(size = 6),
    panel.border = element_rect(colour = "black", fill = NA, size = .5)
  )

pg.target
```

## SOX4 exp.

```{r , fig.height=12, fig.width=1}
x <- FetchData(srat, slot = "data", vars = "SOX4")
sox4_exp <- tibble(exp = x$SOX4, cell = srat$cellType3) %>% 
  group_by(cell) %>% 
  summarise(mexp = mean(exp)) %>% 
  mutate(cell = factor(cell, levels = labels_sorted),
         name = "SOX4")

# td-types
td_types <- readRDS( here::here("output", "00.Fig.Td-type", "td_types.rds"))
td_type_red <- ifelse(labels_sorted %in% td_types$level_3, "red", "black")

pg.sox4 <- sox4_exp %>%
  ggplot(aes(x = name, y = cell, fill = mexp)) +
  geom_tile() +
  scale_y_discrete(position = "right") +
  scale_fill_gradient(
    low = "white", high = "red"
  ) +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    axis.text.y = element_text(color = td_type_red),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "top",
    text = element_text(size = 6),
    panel.border = element_rect(colour = "black", fill = NA, size = .5)
  )

pg.sox4
```

## Align

tree

```{r , fig.height=12, fig.width=2}
pdf(file = here::here("output", DOCNAME, "Fig.cell_freq_hc_tree.pdf"), width = 2, height = 12)

dend <- as.dendrogram(hc)
# rev
dend <- rev(dend)
labels_sorted <- rev(hc$labels[hc$order])

# label color
label_colors <- metadata %>%
  dplyr::select(cellType, cellType3, -barcode) %>%
  distinct() %>%
  mutate(color = cell_color_maps[cellType]) %>%
  column_to_rownames("cellType3")

# Set the plot margin: bottom, left, top & right
par(
  mar = c(0, 0, 0, 4) + 0.1,
  xpd = NA
) # allow content to go into outer margin
dend <- dend %>%
  set("labels_cex", .4) %>% # label size
  set("labels_col", value = label_colors[labels_sorted, ]$color) %>% # label color
  set("branches_lwd", 2) %>%
  set("branches_k_color", k = 5, value = c(
    "#999999", "#E69F00", "#56B4E9", "#009E73",
    "#F0E442", "#0072B2", "#D55E00", "#CC79A7"
  )) %>% # branch color
  raise.dendrogram(1) # height
dend %>%
  plot(horiz = TRUE, axes = FALSE)
colored_bars(colors = label_colors[labels_sorted,]$color, horiz=TRUE,
             rowLabels = c("cell"), y_shift = .8, cex.rowLabels = 0.5)

invisible(dev.off())
```

scores

```{r, fig.width=7.2, fig.height=12}
p.roe + theme(axis.text.y = element_blank()) +
  pg.pathway + theme(axis.text.y = element_blank()) +
  pg.cytokine + theme(axis.text.y = element_blank()) +
  pg.cytokine_r + theme(axis.text.y = element_blank()) +
  pg.target + theme(axis.text.y = element_blank()) +
  pg.sox4 +
  plot_layout(nrow = 1, widths = c(1.5, 2.8, 2.5, 2.2, 2, 0.3))

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.cell_freq_scores.pdf"),
  width = 7.2, height = 12
)
```


## Session info
