---
title: "00.Fig.myeloid.DC.supp"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.myeloid.DC.supp"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
#library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(
  font_size = 6,# Overall font size
  rel_small = 6 / 6,# axis tick labels
  rel_tiny = 6 / 6,# caption
  rel_large = 6 / 6 # title
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# Seurat
library(Seurat)
library(monocle3)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load meta data

```{r load-meta, message=FALSE, warning=FALSE}
meta.data <- read_csv(here::here("output/04.rm_cells/seurat_dc.cellmeta.csv")) %>% 
  mutate(
    level_2 = case_when(
      level_2 == "tDC" ~ "mregDC",
      TRUE ~ level_2
    ),
    level_3 = case_when(
      level_3 == "tDC-C3" ~ "mregDC-C3",
      TRUE ~ level_3
    )
  )
sam_info <- read_csv(here::here("data/sample_info.3.csv")) %>%
  mutate(cell.filtered = as.numeric(cell.filtered))
```

Load rds

```{r load-rds, cache=FALSE}
srat = readRDS(here::here('output/04.rm_cells/seurat_dc.rds'))
srat
```

Modify srat

- tDC to mregDC
- levels

```{r}
srat@meta.data <- srat@meta.data %>% 
  rownames_to_column() %>% 
  mutate(
    level_2 = case_when(
      level_2 == "tDC" ~ "mregDC",
      TRUE ~ level_2
    ),
    level_3 = case_when(
      level_3 == "tDC-C3" ~ "mregDC-C3",
      TRUE ~ level_3
    )
  ) %>% 
  mutate(
    level_3 = factor(level_3, levels = c("cDC2-C1", "cDC2-C2", "mregDC-C3", "cDC1-C5", "cDC1-C6", "pDC-C7", "pDC-C8", "pDC-C4"))
  ) %>% 
  column_to_rownames()

Idents(srat) <- "level_3"
```

## Selected gene {.tabset}

### Chemokine receptors and Nonchemokine receptors

```{r fig.width=7.5, fig.height=4, warning=FALSE, message=FALSE}
markers <- list(
  chemokine_receptor = filter_features(srat, c("CCR2", "CCR5", "CXCR4", "CCR7", "XCR1", "CCR6", "CCR10", "CXCR3", "CXCR4", "CCR1", "CCR3")),
  # Nonchemokine receptors
  nonchemokine_receptor = filter_features(srat, c(
    "CMKLR1", "FPR1", "FPR2", "FPR3", "C5AR1", "PTAFR", "FPR2",
    "C1QBP", "PTGER2", "PTGER4", "AKT2", "LTB4R", "LTB4R2",
    "CYSLTR1", "GPR183", "S1PR1", "P2RY1", "RAPGEF3", "ACVR1B",
    "ACVR2A", "AGER", "IL18R1"
  )),
  tlr_and_adapter = filter_features(srat, c(
    "MYD88", "TICAM1", "TLR10", "TLR9",
    "TLR8", "TLR7", "TLR6", "TLR5",
    "TLR4", "TLR3", "TLR2", "TLR1"
  ))
)

# col order
col_order <- c("cDC2-C1", "cDC2-C2", "mregDC-C3", "cDC1-C5", "cDC1-C6", "pDC-C7", "pDC-C8", "pDC-C4")

# order of cells
df <- srat@meta.data %>%
  rownames_to_column("barcode") %>%
  droplevels() %>%
  dplyr::select(barcode, level_3, Origin2_n) %>%
  mutate(level_3 = factor(level_3, levels = col_order)) %>%
  arrange(level_3, Origin2_n)

# get mat
mat_raw <- t(FetchData(srat, vars = unlist(markers), slot = "data"))
mat_raw <- mat_raw[, df$barcode]

# top anno
top_ha <- HeatmapAnnotation(
  cell = df$level_3,
  col = list(cell = mye_level_3_color),
  show_annotation_name = c(cell = FALSE)
)
bottom_ha <- HeatmapAnnotation(
  tissue = df$Origin4,
  col = list(tissue = origin4_color_maps),
  show_annotation_name = c(tissue = FALSE)
)

# col
x <- ceiling(max(abs(mat_raw)))
col_fun <- circlize::colorRamp2(seq(0, x, length = 50), colorRampPalette(c("white", "red"))(50))

# plot
ht = Heatmap(mat_raw,
  name = "Exp.",
  col = col_fun,
  cluster_rows = FALSE, cluster_columns = FALSE,
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 6, fontface = "italic"),
  column_split = factor(
    df$level_3,
    levels = levels(df$level_3)
  ),
  row_split = factor(
    rep(names(markers), sapply(markers, length)),
    levels = names(markers)
  ),
  row_title_gp = gpar(fontsize = 6),
  column_title = NULL,
  column_gap = unit(0, "mm"),
  border = TRUE,
  top_annotation = top_ha,
  bottom_annotation = bottom_ha,
  heatmap_legend_param = list(
  title_position = "topcenter",
  direction = "vertical",
  labels_gp = gpar(fontsize = 6),
  title_gp = gpar(fontsize = 6),
  grid_height = unit(1.5, "mm"),
  grid_width = unit(1.5, "mm")
)
)

p <- plot_grid(grid.grabExpr(draw(ht)))
p

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.score_Chemokine_NonChemokine_receptor.pdf"),
  width = 7.5, height = 4
)
```

### DC functional score {.tabset}

```{r fig.width=7.3, fig.height=1.5}
func_markers <- list(
  # migration
  mig = filter_features(srat, c(
    "CCR7", "DOCK8", "EZH1", "EZH2", "L1CAM", "CMKLR1", "CD74", "HLA-DRA",
    "HLA-DQA1", "HLA-DRB5", "HLA-DPB1", "EPS8", "ASB2", "RAPGEF3", "RAP1A",
    "ITPR1", "MYH2", "RHOA", "DIAPH1", "CDC42", "ACTR2", "ACTR3", "CD37", "CD82", "CD2AP"
  )),
  # maturation
  mat = filter_features(srat, c(
    "CCR7", "IKKB", "CHUK", "HLA-DRA", "HLA-DRB1", "HLA-DQA1",
    "CD40", "CD80", "CD86", "CD83", "MYD88", "MAP3K7", "IFNA1",
    "IFNB1", "IFNG"
  )),
  # steady state maturation
  ste = filter_features(srat, c(
    "CCR7", "CCL19", "ITGB8", "LTB", "SOCS2", "RASAL1", "STAT4",
    "IL4I1", "CD40", "CCL22", "USP18", "IRF1", "GBP2", "IL15RA", "RELB"
  )),
  # Inflammatory DC maturation
  inf = filter_features(srat, c(
    "CD86", "CXCL10", "IRF7", "ISG15", "IFIT1", "SOCS1", "IL12B",
    "TNFIAP3", "CD40", "CCL22", "USP18", "IRF1", "GBP2", "IL15RA", "RELB"
  )),
  # KEGG atigen processing and presentation
  anti = filter_features(srat, c(
    "HLA-DOA", "HLA-DOB", "KLRC3", "KLRD1", "KLRC1", "KLRC2", "RFXAP", "RFX5",
    "IFNA5", "IFNA4", "IFNA2", "IFNA1", "LGMN", "PSME3", "CTSS", "HLA-C", "HLA-B",
    "HLA-DMB", "HLA-DMA", "HLA-A", "HSPA1L", "HSPA1B", "HSPA2", "KIR2DS5", "HLA-G",
    "KIR3DL1", "KIR3DL2", "HSPA1A", "RFXANK", "CREB1", "IFNA17", "HSPA5", "CD74",
    "HSPA4", "IFNA21", "CTSL", "IFNA6", "IFI30", "IFNA7", "IFNA8", "IFNA10", "IFNA13",
    "CIITA", "IFNA14", "IFNA16", "KIR2DL5A", "HSPA6", "HLA-DRB4", "CD4", "KIR2DL1",
    "HLA-DRB5", "LTA", "KIR3DL3", "HLA-DRB3", "TAP2", "KIR2DS4", "TAPBP", "CD8A",
    "CD8B", "TAP1", "KIR2DS3", "HSPA8", "KIR2DL4", "CANX", "KIR2DS1", "KIR2DL2",
    "KIR2DL3", "KLRC4", "NFYC", "HSP90AA1", "NFYA", "NFYB", "HLA-F", "CTSB",
    "HLA-E", "CALR", "HLA-DPA1", "HLA-DPB1", "HLA-DQA1", "PDIA3", "HLA-DQA2",
    "HLA-DQB1", "PSME1", "PSME2", "HSP90AB1", "B2M", "HLA-DRB1", "HLA-DRA"
  )),
  # KEGG leukocyte transendothelial migration
  leu = filter_features(srat, c(
    "CDH5", "JAM3", "AFDN", "TXK", "ITGAL", "ITGAM", "ITGA4", "SIPA1", "ROCK1", "MMP2",
    "PTK2", "MYL7", "NCF4", "MMP9", "NCF2", "CLDN20", "PECAM1", "VAV1", "VASP", "VCAM1",
    "ITGB1", "MYL12A", "THY1", "VAV2", "ITGB2", "CTNNA1", "JAM2", "PIK3R5", "CLDN18",
    "MAPK14", "MAPK12", "VCL", "CLDN17", "GNAI1", "MYL2", "GNAI2", "CTNND1", "MYL5",
    "PXN", "RAPGEF3", "MYL10", "ARHGAP5", "BCAR1", "GNAI3", "CLDN16", "CLDN23", "EZR",
    "CXCL12", "PLCG1", "PLCG2", "PTK2B", "RHOH", "OCLN", "CDC42", "ACTN2", "ACTN3",
    "PTPN11", "PRKCB", "PRKCA", "CLDN10", "CLDN8", "CLDN2", "CLDN6", "MYL9", "CLDN5",
    "CLDN1", "ICAM1", "RHOA", "CXCR4", "ESAM", "ACTN4", "PRKCG", "ACTN1", "MYLPF",
    "F11R", "NOX1", "CLDN19", "RAC2", "MAPK11", "MAPK13", "CLDN15", "CD99", "CLDN22",
    "ARHGAP35", "ACTG1", "RAC1", "ITK", "CTNNA2", "RASSF5", "CTNNB1", "CLDN11",
    "CLDN7", "MYL12B", "CLDN4", "CLDN3", "ACTB", "ROCK2", "RAP1A", "PIK3R3", "RAP1B",
    "CLDN14", "PIK3CA", "CLDN9", "MSN", "PIK3CB", "PIK3CD", "NCF1", "CTNNA3",
    "NOX3", "RAPGEF4", "CYBA", "PIK3CG", "VAV3", "CYBB", "PIK3R1", "PIK3R2"
  ))
)

# get score
srat <- AddModuleScore(srat, features = func_markers)
mat <- srat@meta.data %>%
  rownames_to_column("barcode") %>%
  dplyr::select(
    barcode,
    mig = Cluster1, mat = Cluster2, ste = Cluster3,
    inf = Cluster4, anti = Cluster5, leu = Cluster6
  ) %>%
  column_to_rownames("barcode") %>%
  t()
mat <- mat[, df$barcode]

# col
x <- ceiling(max(abs(mat)))
col_fun <- circlize::colorRamp2(seq(-x, x, length = 50), rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(50)))

# plot
ht = Heatmap(mat,
  name = "Exp.",
  col = col_fun,
  cluster_rows = FALSE, cluster_columns = FALSE,
  row_names_side = "left",
  row_names_gp = gpar(fontsize = 6),
  show_column_names = FALSE,
  column_split = factor(
    df$level_3,
    levels = levels(df$level_3)
  ),
  column_title = NULL,
  column_gap = unit(0, "mm"),
  border = TRUE,
  top_annotation = top_ha,
  bottom_annotation = bottom_ha,
  heatmap_legend_param = list(
  title_position = "topcenter",
  direction = "vertical",
  labels_gp = gpar(fontsize = 6),
  title_gp = gpar(fontsize = 6),
  grid_height = unit(1.5, "mm"),
  grid_width = unit(1.5, "mm")
)
)

p <- plot_grid(grid.grabExpr(draw(ht)))
p

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.score_DC_functional.pdf"),
  width = 7.3, height = 1.5
)
```

### P14 GO signatures

```{r fig.width=7, fig.height=1.5, message=FALSE, warning=FALSE}
func_markers <- list(
  # leukocyte chemotaxis
  leu_chemotaxis = filter_features(srat, GOfuncR::get_anno_genes("GO:0030595") %>% pull(gene)),
  # antigen processing and presentation
  antigen_pp = filter_features(srat, GOfuncR::get_anno_genes("GO:0019882") %>% pull(gene)),
  # leukocyte cell-cell adhesion
  leu_adhesion = filter_features(srat, GOfuncR::get_anno_genes("GO:0007159") %>% pull(gene)),
  # regulation of leukocyte mediated immunity
  leu_imm = filter_features(srat, GOfuncR::get_anno_genes("GO:0002703") %>% pull(gene)),
  # regulation of T cell activation
  reg_t_act = filter_features(srat, GOfuncR::get_anno_genes("GO:0050863") %>% pull(gene)),
  # T cell receptor signaling pathway
  t_recpt = filter_features(srat, GOfuncR::get_anno_genes("GO:0050852") %>% pull(gene))
)

# get score
srat <- AddModuleScore(srat, features = func_markers)
mat <- srat@meta.data %>%
  rownames_to_column("barcode") %>%
  dplyr::select(
    barcode,
    leu_chemotaxis = Cluster1, antigen_pp = Cluster2, leu_adhesion = Cluster3,
    leu_imm = Cluster4, reg_t_act = Cluster5, t_recpt = Cluster6
  ) %>%
  column_to_rownames("barcode") %>%
  t()
mat <- mat[, df$barcode]

# col
x <- ceiling(max(abs(mat)))
col_fun <- circlize::colorRamp2(seq(-x, x, length = 50), rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(50)))

# plot
ht = Heatmap(mat,
  name = "Exp.",
  col = col_fun,
  cluster_rows = FALSE, cluster_columns = FALSE,
  row_names_side = "left",
  row_names_gp = gpar(fontsize = 6),
  show_column_names = FALSE,
  column_split = factor(
    df$level_3,
    levels = levels(df$level_3)
  ),
  column_title = NULL,
  column_gap = unit(0, "mm"),
  border = TRUE,
  top_annotation = top_ha,
  bottom_annotation = bottom_ha,
  heatmap_legend_param = list(
  title_position = "topcenter",
  direction = "vertical",
  labels_gp = gpar(fontsize = 6),
  title_gp = gpar(fontsize = 6),
  grid_height = unit(1.5, "mm"),
  grid_width = unit(1.5, "mm")
)
)

p <- plot_grid(grid.grabExpr(draw(ht)))
p

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.score_Selected_GO.pdf"),
  width = 7.3, height = 1.5
)
```

### Chemokine

Markers: chemokine, list from: <https://en.wikipedia.org/wiki/Chemokine>

```{r Chemokine, fig.width=2, fig.height=2}
markers = list(
  homeostatic = filter_features(srat, c('CCL14', 'CCL19', 'CCL20', 'CCL21', 'CXCL12', 'CXCL13')),
  inflammatory = filter_features(srat, c('CXCL8', 'CCL2', 'CCL3', 'CCL4', 'CCL5', 'CCL11', 'CXCL10'))
  )

ht <- customDoHeatmap(srat, markers, z_score = T, idents = "level_3", cluster_columns = FALSE)
p <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.score_chemokine.pdf"),
  width = 2, height = 2
)
```

### Cytokines

- List from table 1 of: Turner, M.D., Nedjai, B., Hurst, T., and Pennington, D.J. (2014). Cytokines and chemokines: At the crossroads of cell signalling and inflammatory disease. Biochimica et Biophysica Acta (BBA) - Molecular Cell Research 1843, 2563–2582.

```{r Cytokines, fig.width=2, fig.height=4}
markers = list(
  # Adaptive immunity
  adaptive = filter_features(srat, c('IL2', 'IL7', 'IL15', 'IL21', 'CSF2', 'TSLP')),
  # Pro-inflammatory signalling
  proinflammatory = filter_features(srat, c('IL12A', 'IL12B', 'IL23A', 'IL27', 'EBI3', 'IL10', 'IL19', 'IL20', 'IL22', 'IL24')),
  # Anti-inflammatory signalling
  antiinflammatory = filter_features(srat, c('IL1A', 'IL1B', 'IL1RN', 'IL18', 'IL33', 'IL36A', 'IL36G', 'IL36RN', 'IL6', 'IL31', 'CNTF', 'CTF1', 'LIF', 'SPP1', 'OSM', 'TNF', 'BAFF', 'TNFSF13', 'IL17A', 'IFNB1', 'IFNK', 'IFNG'))
  )

ht <- customDoHeatmap(srat, markers, z_score = T, idents = "level_3", cluster_columns = FALSE)
p <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.score_Cytokines.pdf"),
  width = 2, height = 4
)
```



















