---
title: "00.Fig.overview.240615"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.overview.240615"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# Seurat
library(Seurat)

# vis
# library(scRNAtoolVis)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load rds

```{r load-rds, cache=FALSE}
srat <- readRDS(here::here("output/04.rm_cells/seurat.rds"))
srat
```

## UMAP {.tabset}

### cell type

```{r fig.width=2, fig.height=2}
p.umap.cellType <- DimPlot(srat,
  reduction = "umap", group.by = "cellType",
  label = TRUE, label.size = geom.text.size, raster = TRUE,
  shuffle = TRUE
) +
  scale_color_manual(values = cell_color_maps) +
  labs(title = NULL) +
  my_theme +
  theme(
    legend.position = "none",
    plot.margin = margin(rep(0, 4), unit = "cm")
  )
p.umap.cellType

# saveRDS(p.umap.cellType, file = here::here("output", DOCNAME, "Fig.umap.cellType.rds"))
ggsave2(p.umap.cellType,
  filename = here::here("output", DOCNAME, "Fig.umap.cellType.pdf"),
  height = 2, width = 2, device = cairo_pdf
)
```

### Origin2

```{r fig.width=2.6, fig.height=2}
p.umap.origin2 <- DimPlot(srat,
  reduction = "umap", group.by = "Origin2_n",
  raster = TRUE, shuffle = TRUE
) +
  scale_color_manual(values = origin2_color_maps) +
  labs(title = NULL) +
  my_theme
p.umap.origin2

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.umap.Origin2.pdf"),
  height = 2, width = 2.6, device = cairo_pdf
)
```

Split

```{r fig.width=7.5, fig.height=2}
p.umap.splitByOrigin2 <- DimPlot(srat,
  reduction = "umap",
  split.by = "Origin2_n", ncol = 7,
  raster = TRUE
) +
  scale_color_manual(values = cell_color_maps) +
  my_theme +
  NoAxes() +
  theme(
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.height = unit(.4, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
    plot.margin = margin(rep(0, 4), unit = "cm"),
    strip.background = element_rect(fill = "#f3f2f1")
  )
p.umap.splitByOrigin2

ggsave2(p.umap.splitByOrigin2,
  filename = here::here("output", DOCNAME, "Fig.umap.splitByOrigin2.pdf"),
  height = 2, width = 7.5, device = cairo_pdf
)
```

### Patient

```{r fig.width=2.5, fig.height=2}
p.umap.patient <- DimPlot(srat,
  reduction = "umap", group.by = "Patient", raster = TRUE, pt.size = .05, shuffle = TRUE
) +
  scale_color_manual(values = patient_color_maps) +
  labs(title = NULL) +
  my_theme
p.umap.patient

ggsave2(p.umap.patient,
  filename = here::here("output", DOCNAME, "Fig.umap.Patient.pdf"),
  height = 2, width = 2.5, device = cairo_pdf
)
```

## Sample composition

Source

```{r sample-compositions, fig.width=6, fig.height=2}
p.sample_compo <- PlotSampleFraction(seurat_obj = srat, "cellType", cell_color_maps)
p.sample_compo

# saveRDS(p, file = here::here("output", DOCNAME, "Fig.sampleComposition.rds"))
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.sampleComposition.pdf"),
  height = 3, width = 7.8, device = cairo_pdf
)
```

## UMAP of epi, fib, endo (epi as cluster)

```{r fig.width=7.5, fig.height=2.5}
srat.epi <- readRDS(here::here("output/04.rm_cells", "seurat_epithelia.rds"))
srat.endo <- readRDS(here::here("output/04.rm_cells", "seurat_endothelia.rds"))
srat.fib <- readRDS(here::here("output/04.rm_cells", "seurat_fibroblasts.rds"))

p.umap.endo <- DimPlot(srat.endo,
  reduction = "umap", group.by = "level_2",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE, shuffle = TRUE
) +
  scale_color_manual(values = endo_level_2_color) +
  labs(title = NULL) +
  my_theme +
  NoLegend()

p.umap.fib <- DimPlot(srat.fib,
  reduction = "umap", group.by = "level_3",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE, shuffle = TRUE
) +
  scale_color_manual(values = fib_level_3_color) +
  labs(title = NULL) +
  my_theme +
  NoLegend()

p.umap.epi <- DimPlot(srat.epi,
  reduction = "umap", group.by = "seurat_clusters",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE, shuffle = TRUE
) +
  scale_color_manual(values = cluster_color_maps) +
  labs(title = NULL) +
  my_theme +
  NoLegend()

p.umap.epi + p.umap.fib + p.umap.endo +
  plot_layout(ncol = 3)
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.main.umap_stromal.pdf"),
  height = 2.5, width = 7.5, device = cairo_pdf
)
```


```{r fig.width=7.5, fig.height=3}
p.umap.endo <- DimPlot(srat.endo,
  reduction = "umap", group.by = "level_2",
  label.size = geom.text.size,
  raster = TRUE, shuffle = TRUE
) +
  scale_color_manual(values = endo_level_2_color) +
  labs(title = NULL) +
  my_theme +
  guides(color = guide_legend(ncol = 4, byrow = TRUE)) +
  theme(legend.margin = margin(rep(0, 4), unit = "cm"),
        legend.spacing.y = unit(0, "cm"),
        legend.key.height = unit(.4, "cm"),
        legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
        legend.position = "bottom")

p.umap.fib <- DimPlot(srat.fib,
  reduction = "umap", group.by = "level_3",
  label.size = geom.text.size,
  raster = TRUE, shuffle = TRUE
) +
  scale_color_manual(values = fib_level_3_color) +
  labs(title = NULL) +
  my_theme +
  guides(color = guide_legend(ncol = 4, byrow = TRUE)) +
  theme(legend.margin = margin(rep(0, 4), unit = "cm"),
        legend.spacing.y = unit(0, "cm"),
        legend.key.height = unit(.4, "cm"),
        legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
        legend.position = "bottom")

p.umap.epi <- DimPlot(srat.epi,
  reduction = "umap", group.by = "seurat_clusters",
  label.size = geom.text.size,
  raster = TRUE, shuffle = TRUE
) +
  scale_color_manual(values = cluster_color_maps) +
  labs(title = NULL) +
  my_theme +
  guides(color = guide_legend(ncol = 8, byrow = TRUE)) +
  theme(legend.margin = margin(rep(0, 4), unit = "cm"),
        legend.spacing.y = unit(0, "cm"),
        legend.key.height = unit(.4, "cm"),
        legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
        legend.position = "bottom")

p.umap.epi + p.umap.fib + p.umap.endo +
  plot_layout(ncol = 3)
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.main.umap_stromal.bottom_legend.pdf"),
  height = 3, width = 7.5, device = cairo_pdf
)
```

## Markers

DotPlot of markers

```{r plot-cell-markers-supp-dot, fig.width=5, fig.height=1.8}
p.marker.dot <- DotPlot(srat, features = c(
  # T
  "CD3D", "CD3E", "CD2",
  # B
  "CD19", "MS4A1", "CD79A", "IGHM",
  # mye
  "LYZ", "FCER1G", "TYROBP", "AIF1",
  # epi
  "CNFN", "CRNN", "KRT6A", "KRT13",
  # fib
  "DCN", "COL1A1", "COL1A2", "LUM",
  # endo
  "CDH5", "VWF", "PECAM1", "CLDN5",
  # platelets
  "PPBP", "PF4"
), dot.scale = 4) +
  labs(x = NULL, y = NULL) +
  scale_y_discrete(limits = rev) +
  my_theme +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1)
  )
p.marker.dot

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.dotplot.markers.pdf"),
  height = 1.8, width = 5, device = cairo_pdf
)
```

## Composition plot


### celltype by Origin2_n

```{r fig.width=2, fig.height=1.5}
srat@meta.data %>% 
  group_by(Origin2_n, cellType) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin2_n, y = Percent, fill = cellType)) + 
  geom_bar(stat = 'identity') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.compo.cellTypeByOrigin2.pdf"),
  width = 2, height = 1.5, device = cairo_pdf
)
```



### Frac. changes by Origin2

```{r overall-fraction-change-Origin2, fig.width=10, fig.height=2.5, cache=FALSE}
# T
df.t <- readRDS(here::here("output", "freq.frac_change.t", "tb.fc.level_1.rds"))
p.fc_cd8 <- PlotFracChange(df.t %>% filter(level_1 == "CD8"), x = "Origin2_n") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.fc_cd4 <- PlotFracChange(df.t %>% filter(level_1 == "CD4"), x = "Origin2_n") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.fc_treg <- PlotFracChange(df.t %>% filter(level_1 == "Treg"), x = "Origin2_n") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# B
df.b <- readRDS(here::here("output", "freq.frac_change.b", "tb.fc.rds"))
p.fc_b <- PlotFracChange(df.b, x = "Origin2_n") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# mac
df.mye <- readRDS(here::here("output", "freq.frac_change.mye", "tb.fc.level_1.rds"))
p.fc_mono <- PlotFracChange(df.mye %>% filter(level_1 == "Mono"), x = "Origin2_n") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.fc_mac <- PlotFracChange(df.mye %>% filter(level_1 == "MΦ"), x = "Origin2_n") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.fc_dc <- PlotFracChange(df.mye %>% filter(level_1 == "DC"), x = "Origin2_n") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# fib
df.fib <- readRDS(here::here("output", "freq.frac_change.fib", "tb.fc.rds"))
p.fc_fib <- PlotFracChange(df.fib, x = "Origin2_n") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# endo
df.endo <- readRDS(here::here("output", "freq.frac_change.endo", "tb.fc.rds"))
p.fc_endo <- PlotFracChange(df.endo, x = "Origin2_n") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p.fc_cd8 + p.fc_cd4 + p.fc_treg +
  p.fc_b +
  p.fc_mono + p.fc_mac + p.fc_dc +
  p.fc_fib + p.fc_endo +
  plot_layout(nrow = 1, guides = "collect")

# save
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.frac_change.byOrigin2.pdf"),
  height = 2.5, width = 10, device = cairo_pdf
)
```





### N3-N5

肿瘤组织内异质性

```{r n3-n5, fig.height=3, fig.width=2.5}
p.cellByN3N4N5 <- srat@meta.data %>%
  filter(Tissue == "Tumor") %>%
  group_by(Patient, Origin, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100) %>%
  ggplot(aes(x = Origin, y = Percent, fill = cellType)) +
  geom_col() +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(. ~ Patient) +
  my_theme +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.height = unit(.4, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
    plot.margin = margin(rep(0, 4), unit = "cm"),
    strip.background = element_rect(fill = "#f3f2f1")
  )
p.cellByN3N4N5
```

### Metastatic

转移阳性的病人与阴性病人

```{r Metastatic, fig.height=3, fig.width=2.5}
p.cellByMetastatic <- srat@meta.data %>%
  filter(Origin2_n != "pLN") %>%
  group_by(Origin2_n, Metastatic, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100) %>%
  ggplot(aes(x = Metastatic, y = Percent, fill = cellType)) +
  geom_col() +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(. ~ Origin2_n, ncol = 3) +
  my_theme +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.height = unit(.4, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
    plot.margin = margin(rep(0, 4), unit = "cm"),
    strip.background = element_rect(fill = "#f3f2f1")
  )
p.cellByMetastatic
```

## Align plots

### Main

Row 2: celltype umap, sample composition

```{r fig.width=8, fig.height=2}
plot_grid(p.umap.cellType, p.sample_compo, ncol = 2, rel_widths = c(2, 6))

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.main.r2.pdf"),
  height = 2, width = 8, device = cairo_pdf
)
```

Row 3: umap of Origin2, split

```{r fig.width=7.5, fig.height=1.1}
p.umap.splitByOrigin2
```

Row 4: umap of T, Mye, B

Row 5: umap of epi, fib, endo

### Supp

Row 1: umap of patients, dotplot of markers

```{r fig.width=7.5, fig.height=2}
plot_grid(p.umap.patient, p.marker.dot, rel_widths = c(1, 2))

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.supp.r1.pdf"),
  height = 2, width = 7.5, device = cairo_pdf
)
```

Row 2

```{r fig.width=5, fig.height=3}
p.cellByN3N4N5 + p.cellByMetastatic +
  plot_layout(ncol = 2)

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.supp.r2.pdf"),
  height = 3, width = 5, device = cairo_pdf
)
```

## Session info

