---
title: "00.Fig.endothelia"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.endothelia"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
theme_set(theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6, # title
  font_family = "Arial"
))
my_theme <- theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6, # title
  font_family = "Arial"
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)
library(circlize)

# color
library(ggsci)

# Seurat
library(Seurat)

# monocle
library(monocle3)

# enrichment
library(clusterProfiler)
library(org.Hs.eg.db)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

## Load data

Load srat

```{r load-rds, cache=FALSE}
srat <- readRDS(here::here("output/04.rm_cells/seurat_endothelia.rds"))
Idents(srat) <- "level_2"
srat
```

Load metadata

```{r load-meta, cache=FALSE, message=FALSE, warning=FALSE}
metadata <- read_csv(here::here("output/04.rm_cells/seurat_endothelia.cellmeta.csv")) %>%
  mutate(cellType3 = level_2)
sam_info <- read_csv(here::here("data/sample_info.3.csv"))
```

Load c2c

```{r load-c2c, cache=FALSE}
df_c2c <- readRDS(here::here("data", "c2c.endo.rds"))
head(df_c2c)
```

## UMAP {.tabset}

### Level 2

```{r fig.width=3, fig.height=2.2, cache=FALSE}
p.umap.cellType <- DimPlot(srat,
  reduction = "umap", group.by = "level_2",
  raster = TRUE
) +
  scale_color_manual(values = endo_level_2_color) +
  labs(title = NULL) +
  my_theme
p.umap.cellType

ggsave2(p.umap.cellType,
  filename = here::here("output", DOCNAME, "Fig.umap.cellType.pdf"),
  height = 2.5, width = 3, device = cairo_pdf
)
```

With label

```{r fig.width=2.5, fig.height=2.5, cache=FALSE}
p.umap.celltype <- DimPlot(srat,
  reduction = "umap", group.by = "level_2",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE
) +
  scale_color_manual(values = endo_level_2_color) +
  labs(title = NULL) +
  my_theme +
  theme(legend.position = "none")
p.umap.celltype

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.umap.cellType.labelled.pdf"),
  height = 2.5, width = 2.5, device = cairo_pdf
)
```

## Classification

### markers

```{r fig.width=3, fig.height=3.5, message=FALSE, warning=FALSE}
markers <- list(
  arteries = filter_features(srat, c("GJA5", "DKK2", "FBLN5")),
  pcv = filter_features(srat, c("SELE", "SELP", "ACKR1", "CPE")),
  immature = filter_features(srat, c("RBP7", "GSN", "ID2", "ID3", "ENG", "PLVAP", "VWA1", "MMP2", "KDR")),
  tip = filter_features(srat, c(
    "CXCR4", "ANGPTL2", "ESM1", "APLN" # , "PGF", "PDGFB", 'ESM1', 'PGF'
  )),
  lymphatics = filter_features(srat, c(
    "PDPN", "LYVE1" # , "CCL21", "TFF3", "MMRN1"
  ))
)

mat <- AverageExpression(srat,
  assays = "RNA", features = unlist(markers),
  slot = "data", verbose = F
)$RNA
mat <- log1p(as.matrix(mat))
# scale
mat_scale <- pheatmap:::scale_rows(mat)

# color
x <- ceiling(max(abs(mat_scale)))
col_fun <- circlize::colorRamp2(c(-x, 0, x), c("#476fa9", "#ffffff", "#ca3226"))
legend_name <- "Exp."

ht <- ComplexHeatmap::Heatmap(mat_scale,
  name = legend_name,
  col = col_fun,
  cluster_columns = FALSE, cluster_rows = FALSE,
  show_row_names = TRUE, show_column_names = TRUE,
  row_names_side = "right",
  row_names_gp = gpar(fontsize = 6, fontface = "italic"),
  column_names_gp = gpar(fontsize = 6),
  column_names_rot = 60,
  row_split = factor(
    rep(names(markers), sapply(markers, length)),
    levels = names(markers)
  ),
  row_title_gp = gpar(fontsize = 6),
  column_split = factor(c(
    rep(c("arteries"), 1),
    rep(c("pcv"), 5),
    rep(c("immature"), 1),
    rep(c("tip"), 1),
    rep(c("lymph"), 2)
  ),
  levels = c("arteries", "pcv", "immature", "tip", "lymph")
  ),
  column_title = NULL,
  left_annotation = rowAnnotation(
    empty = anno_empty(
      border = FALSE,
      width = unit(1, "mm")
    ),
    foo = anno_block(
      gp = gpar(
        fill = "gray",
        col = "gray"
      ),
      width = unit(0.5, "mm")
    )
  ),
  heatmap_legend_param = list(
    title_position = "topcenter",
    direction = "vertical",
    labels_gp = gpar(fontsize = 6),
    title_gp = gpar(fontsize = 8),
    grid_height = unit(3, "mm"),
    grid_width = unit(3, "mm")
  )
)

# draw(ht)
p.markers.canonical <- plot_grid(grid.grabExpr(draw(ht)))
p.markers.canonical

# save
saveRDS(p.markers.canonical, file = here::here("output", DOCNAME, "Fig.markers.canonical.rds"))
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.markers.canonical.pdf"),
  height = 3.5, width = 3, device = cairo_pdf
)
```

Small

```{r fig.width=1.8, fig.height=2, message=FALSE, warning=FALSE}
ht <- ComplexHeatmap::Heatmap(mat_scale,
  name = legend_name,
  col = col_fun,
  cluster_columns = FALSE, cluster_rows = FALSE,
  show_row_names = TRUE, show_column_names = TRUE,
  row_names_side = "right",
  row_names_gp = gpar(fontsize = 4, fontface = "italic"),
  column_names_gp = gpar(fontsize = 4),
  column_names_rot = 60,
  row_split = factor(
    rep(names(markers), sapply(markers, length)),
    levels = names(markers)
  ),
  row_title_gp = gpar(fontsize = 4),
  column_split = factor(c(
    rep(c("arteries"), 1),
    rep(c("pcv"), 5),
    rep(c("immature"), 1),
    rep(c("tip"), 1),
    rep(c("lymph"), 2)
  ),
  levels = c("arteries", "pcv", "immature", "tip", "lymph")
  ),
  column_title = NULL,
  left_annotation = rowAnnotation(
    empty = anno_empty(
      border = FALSE,
      width = unit(1, "mm")
    ),
    foo = anno_block(
      gp = gpar(
        fill = "gray",
        col = "gray"
      ),
      width = unit(0.5, "mm")
    )
  ),
  heatmap_legend_param = list(
    title_position = "topcenter",
    direction = "vertical",
    labels_gp = gpar(fontsize = 4),
    title_gp = gpar(fontsize = 6),
    grid_height = unit(3, "mm"),
    grid_width = unit(3, "mm")
  )
)

# draw(ht)
p.markers.canonical <- plot_grid(grid.grabExpr(draw(ht)))
p.markers.canonical

# save
saveRDS(p.markers.canonical, file = here::here("output", DOCNAME, "Fig.markers.canonical.small.rds"))
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.markers.canonical.small.pdf"),
  height = 2, width = 1.8, device = cairo_pdf
)
```


## Origin2 enrichment

```{r fig.width=1.8, fig.height=2, warning=FALSE}
cluster_cellNum <- metadata %>%
  group_by(cellType3) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  group_by(cellType3, Origin2_n) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp <- setdiff(unique(metadata$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x <- tibble(cellType3 = unique(cluster_cellNum$cellType3), Origin2_n = t, o = 0)
    cluster_byOrigin2 <- bind_rows(cluster_byOrigin2, x)
  }
}

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin2_n") %>%
  left_join(cluster_cellNum, by = "cellType3") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType3, Origin2_n, r) %>%
  spread(Origin2_n, r, fill = 0) %>%
  column_to_rownames("cellType3")

# order
tmp.df <- r_oe %>%
  rownames_to_column("cellType3") %>%
  arrange(nLN, pLN, Normal, Adjacent, Tumor)

# data for plot
r_oe[r_oe > 10] <- 10
d4p <- r_oe %>%
  rownames_to_column("cellType3") %>%
  pivot_longer(cols = !cellType3, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "prePBMC", "postPBMC",
      "nLN", "pLN",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType3 = factor(cellType3, levels = tmp.df$cellType3)
  )

# % lineage prop
df.lineage_prop <- metadata %>%
  group_by(cellType3, Origin2_n) %>%
  summarise(cellType3.cell_num = n()) %>%
  mutate(cellType3.cell_num = as.numeric(cellType3.cell_num)) %>%
  ungroup() %>%
  mutate(tot = nrow(metadata)) %>%
  mutate(lineage.prop = cellType3.cell_num * 100 / tot)

# plot
p.roe <- d4p %>%
  filter(!str_detect(Origin2_n, "PBMC")) %>%
  ggplot(aes(x = Origin2_n, y = cellType3, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.lineage_prop,
    aes(x = Origin2_n, y = cellType3, size = lineage.prop),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "% lineage") +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
p.roe

# save
saveRDS(p.roe, file = here::here("output", DOCNAME, "Fig.Roe.byOrigin2_n.rds"))
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.Roe.byOrigin2_n.pdf"),
  height = 2, width = 1.8, device = cairo_pdf
)
```

## Selected genes

### Leukocyte Adhesion Molecules, and Chemokines

- E-selectin is only expressed in activated endothelia cells

- In activated endothelium, E-selectin expression is largely confined to postcapillary venules (Aird, 2012)

- In “resting” endothelium, P-selectin is stored intracellularly in preexisting Weibel–Palade bodies and is expressed preferentially in postcapillary venules (Aird William C., 2007)

- Binding of endothelial E-selectin and P-selectin to leukocyte carbohydrate-based ligands promotes rolling of leukocytes to the endothelium

- Binding of endothelial cell adhesion molecules, including vascular cell adhesion molecule
(VCAM)-1 and intercellular adhesion molecule (ICAM)-1 to leukocyte integrins mediates firm adhesion.

- T cells for tumor immunotherapy was enhanced when proinflammatory agents that up-regulated ICAM and VCAM (Dudley, 2012)

```{r selectin-no-PECAM1, fig.width=3, fig.height=3}
p <- StackedVlnPlot(srat, features = c("SELE", "SELP", "ICAM1", "ICAM2", "VCAM1"))
p

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.exp_selected_genes.selectin.pdf"),
  height = 3, width = 3, device = cairo_pdf
)
```

## cmp. subtypes

### pcv-C1C2C3_vs_pcv-C5

```{r fig.width=2.5, fig.height=2}
ck <- read.table(here::here('output', "06.endothelia.cmp_pcv", 'deg.cmp_tn_pcv.KEGG.txt'), sep = "\t", header = T)

d4p <- ck %>% 
  filter(Description %in% c("Cytokine-cytokine receptor interaction",
                            "MAPK signaling pathway")) %>% 
  rowwise %>% 
  mutate(ratio = eval(rlang::parse_expr(GeneRatio)))

ggplot(data = d4p, aes(x = Cluster, y = Description)) +
  geom_point(aes(size = ratio, color = p.adjust)) +
  scale_color_continuous(high = "#e06663", low = "#327eba") +
  labs(x = NULL, y = NULL) +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

ggsave2(
  filename = here::here("output", DOCNAME, "deg.cmp_pcv-C1C2C3_vs_pcv-C5.KEGG.pdf"),
  width = 2.5, height = 2, device = cairo_pdf
)
```

### LEC-C2_vs_LEC-C1

```{r fig.width=2.5, fig.height=2}
deg <- read_csv(here::here('output', "06.endothelia.cmp_lymphatic", 'deg.LEC-C2_vs_LEC-C1.csv'))
compareList = list(
  LEC_C2 = deg %>%filter(p_val_adj < 0.01, avg_logFC > 0.25) %>% pull(gene) %>%
    bitr(fromType="SYMBOL", toType=c("ENTREZID"), OrgDb="org.Hs.eg.db") %>% pull(ENTREZID),
  LEC_C1 = deg %>% filter(p_val_adj < 0.01, avg_logFC < -0.25) %>% pull(gene) %>%
    bitr(fromType="SYMBOL", toType=c("ENTREZID"), OrgDb="org.Hs.eg.db") %>% pull(ENTREZID)
)

cg.BP = compareCluster(compareList, fun = "enrichGO", OrgDb = org.Hs.eg.db, ont = "BP",
                       pAdjustMethod = "BH", pvalueCutoff = 0.01, qvalueCutoff = 0.05, readable = T)

d4p <- cg.BP
d4p@compareClusterResult <- d4p@compareClusterResult %>% 
  filter(Description %in% c("aging", "translational initiation"))


dotplot(d4p) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 40)) + 
  #scale_size(range = c(0, 5)) +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

ggsave2(
  filename = here::here("output", DOCNAME, "deg.cmp_LEC-C2_vs_LEC-C1.GO-BP.pdf"),
  width = 2.5, height = 2, device = cairo_pdf
)
```


## Session info
