---
title: "00.Fig.endothelia.supp"
---

```{r knitr, include = FALSE}
DOCNAME <- "Fig.endothelia.supp"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
my_theme <- theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6, # title
  font_family = "Arial"
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)
library(circlize)

# color
library(ggsci)

# srat
library(Seurat)

# monocle
library(monocle3)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

## Load data

Load srat

```{r load-rds, cache=FALSE}
srat <- readRDS(here::here("output/04.rm_cells/seurat_endothelia.rds"))
Idents(srat) <- "level_2"
srat
```

Load metadata

```{r load-meta, cache=FALSE, message=FALSE, warning=FALSE}
metadata <- read_csv(here::here("output/04.rm_cells/seurat_endothelia.cellmeta.csv")) %>%
    mutate(cellType3 = level_2)
sam_info <- read_csv(here::here("data/sample_info.3.csv"))
```

Load c2c

```{r load-c2c, cache=FALSE}
df_c2c = readRDS(here::here("data", "c2c.endo.rds"))
head(df_c2c)
```

## Classification {.tabset}

### FeaturePlot

```{r FeaturePlot-markers, fig.width=5, fig.height=2.3, cache=FALSE}
p.markers = customFeaturePlot(srat, pt.size = 0.1, features = c(
  # blood
  "FLT1",
  # lymphatic
  "PDPN",
  # vein
  "ACKR1",
  # artery
  "GJA5", "VEGFC",
  # tip cells
  "ESM1", "PGF", "PDGFB",
  # pcv
  "SELE", "SELP"
  # 
)) %>%
  wrap_plots(nrow = 2, ncol = 5) & theme(plot.margin = margin(0))
p.markers
```

### SingleR

```{r SinlgeR-Goveia, cache=FALSE, fig.height=3, fig.width=4}
cmp_data_dir = '../../sc_data_to_cmp/Goveia_2020'

matched = readRDS(file.path(cmp_data_dir, 'matched.res_0.8.rds'))
rownames(matched) = df_c2c$level_2
p.singler = summary_ggplot(matched, xlab = "This study", ylab = 'Goveia_2020',
                           geom.text.size = geom.text.size) +
  my_theme +
  theme(axis.text.x=element_text(angle=60, hjust=1))
p.singler
```

### Top 50 DEGs

Get top 50 markers

```{r get-top-markers, message=FALSE, warning=FALSE, fig.width=3, fig.height=4}
# get markers
markers = read_tsv(here::here('output/04.rm_cells/seurat_endothelia.markers.txt'))
# top 50
top50 = markers %>%
  group_by(cluster) %>%
  top_n(50, wt = avg_logFC)

# rename
top50 = top50 %>%
  left_join(df_c2c, by = "cluster")
# de-duplicate
top50 = top50[!duplicated(top50$gene),]

# to list
top50_lst = lapply(levels(top50$level_2), function(x) {
  top50 %>%
    filter(level_2 == x) %>%
    pull(gene)
})
names(top50_lst) = levels(top50$level_2)

# get exp
mat = AverageExpression(srat, assays = "RNA", features = unlist(top50_lst),
                        slot = "data", verbose = F)$RNA
mat = log1p(as.matrix(mat))

# scale
mat_scale = pheatmap:::scale_rows(mat)

# color
x = ceiling(max(abs(mat_scale)))
col_fun = circlize::colorRamp2(c(-x, 0, x), c("blue", "white", "red"))

# marker
marker_2_lab = c("PLVAP", "SPARC", # 
                 "APLNR", "SOX4",
                 "ESM1", "PGF",
                 "ACKR1", "SELP", "SELE",
                 "CA2", "RGCC",
                 "PROX1", "PDPN",
                 "APOD", "TIMP1",
                 "CXCL8", "IL6",
                 "FBLN5", "GJA5",
                 "CD74", "CLU"
                 )
# location
marker_2_lab = sapply(marker_2_lab, function(x){which(rownames(mat) == x)}, simplify = T)
ha = rowAnnotation(foo = anno_mark(at = marker_2_lab,
                                   labels = names(marker_2_lab),
                                   labels_gp = gpar(fontsize = 6)))

ht = Heatmap(mat_scale, name = "Exp.",
        col = col_fun,
        right_annotation = ha,
        cluster_columns = F,
        cluster_rows = F,
        show_row_names = FALSE,
        column_names_gp = gpar(fontsize = 6),
        #column_names_rot = 60,
        heatmap_legend_param = list(#direction = "horizontal",
                                    labels_gp = gpar(fontsize = 6),
                                    title_gp = gpar(fontsize = 6),
                                    #title_position = "leftcenter",
                                    grid_height = unit(2, "mm"),
                                    grid_width = unit(2, "mm"))
        )
#draw(ht, heatmap_legend_side = "bottom")
p.markers.top50 = plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p.markers.top50
```

## Composition

### Frac. changes by Origin4

```{r overall-fraction-change-Origin4, fig.width=2, fig.height=2, cache=FALSE}
tb.fc <- readRDS(here::here("output", "freq.frac_change.endo", "tb.fc.rds"))

p.fractionChange <- PlotFracChange(tb.fc, x = "Origin4") +
  my_theme +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.fractionChange

# save
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.frac_change.byOrigin4.pdf"),
  height = 2.5, width = 2, device = cairo_pdf
)
```

### Origin4 enrichment

```{r fig.width=2, fig.height=2, warning=FALSE}
cluster_cellNum <- metadata %>%
  group_by(cellType3) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin4) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  group_by(cellType3, Origin4) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp <- setdiff(unique(metadata$Origin4), cluster_byOrigin2$Origin4)
if (length(tmp) > 0) {
  for (t in tmp) {
    x <- tibble(cellType3 = unique(cluster_cellNum$cellType3), Origin4 = t, o = 0)
    cluster_byOrigin2 <- bind_rows(cluster_byOrigin2, x)
  }
}

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin4") %>%
  left_join(cluster_cellNum, by = "cellType3") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType3, Origin4, r) %>%
  spread(Origin4, r, fill = 0) %>%
  column_to_rownames("cellType3")

# order
tmp.df <- r_oe %>%
  rownames_to_column("cellType3") %>%
  arrange(nLN, m_nLN, m_pLN, Normal, Adjacent, Tumor)

# data for plot
r_oe[r_oe > 10] <- 10
d4p <- r_oe %>%
  rownames_to_column("cellType3") %>%
  pivot_longer(cols = !cellType3, names_to = "Origin4") %>%
  mutate(
    Origin4 = factor(Origin4, levels = c(
      "prePBMC", "postPBMC",
      "nLN", "m_nLN", "m_pLN",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType3 = factor(cellType3, levels = tmp.df$cellType3)
  )

# % lineage prop
df.lineage_prop <- metadata %>%
  group_by(cellType3, Origin4) %>%
  summarise(cellType3.cell_num = n()) %>%
  mutate(cellType3.cell_num = as.numeric(cellType3.cell_num)) %>%
  ungroup() %>%
  mutate(tot = nrow(metadata)) %>%
  mutate(lineage.prop = cellType3.cell_num * 100 / tot)

# plot
p.roe <- d4p %>%
  filter(!str_detect(Origin4, "PBMC")) %>%
  ggplot(aes(x = Origin4, y = cellType3, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.lineage_prop,
    aes(x = Origin4, y = cellType3, size = lineage.prop),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "% lineage") +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
p.roe

# save
saveRDS(p.roe, file = here::here("output", DOCNAME, "Fig.Roe.byOrigin4.rds"))
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.Roe.byOrigin4.pdf"),
  height = 2, width = 2, device = cairo_pdf
)
```

## AUCell/SCENIC

Hallmark

```{r fig.width=3, fig.height=3}
cell_type <- "endothelia"
anno_level <- "level_2"

# load
de_tf.glm_patient <- read_tsv(here::here("output", "DE_MSigDB_H", paste0(cell_type, ".de_MSigDB_H.glm_patient.tsv")))

# cell types
d4p <- de_tf.glm_patient %>%
  left_join(df_c2c, by = "cluster") %>%
  dplyr::rename(cellType3 = level_2)

# t_mat
t_mat <- d4p %>%
  pivot_wider(id_cols = "gene", names_from = "cellType3", values_from = "t") %>%
  column_to_rownames("gene")
t_mat <- t_mat[, levels(d4p$cellType3)]

# select rows to show
selected_row <- d4p %>%
  filter(p_adj < 0.05) %>% # keep only those with p_adj < 0.05
  group_by(cluster) %>%
  top_n(3, wt = t) %>%
  pull(gene) %>%
  unique()

p.glm.H <- plot_heatmap_glm(as.matrix(t_mat[selected_row, ]),
  cluster_columns = FALSE,
  show_row_dend = FALSE,
  show_column_dend = FALSE
)
p.glm.H

# save
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.DE_Hallmark.pdf"),
  height = 3, width = 3.3, device = cairo_pdf
)
```

TFs

```{r fig.width=2.5, fig.height=3.5}
cell_type <- "endothelia"
anno_level <- "level_2"

# load
de_tf.glm_patient <- read_csv(here::here("output", paste0("DE_TF.", cell_type), "de_tf.glm_patient.csv"))

# cell types
d4p <- de_tf.glm_patient %>%
  left_join(df_c2c, by = "cluster") %>%
  dplyr::rename(cellType3 = level_2)

# t_mat
t_mat <- d4p %>%
  pivot_wider(id_cols = "gene", names_from = "cellType3", values_from = "t") %>%
  column_to_rownames("gene")
t_mat <- t_mat[, levels(d4p$cellType3)]

# select rows to show
selected_row <- d4p %>%
  filter(p_adj < 0.05) %>% # keep only those with p_adj < 0.05
  group_by(cluster) %>%
  top_n(3, wt = t) %>%
  pull(gene) %>%
  unique()

p.glm.TF <- plot_heatmap_glm(as.matrix(t_mat[selected_row, ]),
  cluster_columns = FALSE,
  show_row_dend = FALSE
)
p.glm.TF

# save
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.DE_TFs.pdf"),
  height = 4, width = 2.5, device = cairo_pdf
)
```

## Selected genes

### Angiogenesis

- Paper: Hida, K., Maishi, N., Annan, D.A., and Hida, Y. (2018). Contribution of Tumor Endothelial Cells in Cancer Progression. Int J Mol Sci 19.

```{r Angiogenesis, fig.width=2.5, fig.height=3}
markers <- list(
  Activator = c("FLT1", "KDR", "FLT4", "FGF2", "ANGPT1", "ANGPT2", "TEK", "HGF", "EGF", "PDGFA", "PDGFB", "PGF"),
  Inhibitor = c("THBS1", "DLL4", "VASH1", "RCAN1", "COL18A1", "COL4A1", "COL4A2")
)

ht <- customDoHeatmap(srat, markers, z_score = T, idents = "level_2", cluster_columns = FALSE)
p <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p
```

### MHC-II

- Paper: Goveia, J., Rohlenova, K., Taverna, F., Treps, L., Conradi, L.-C., Pircher, A., Geldhof, V., Rooij, L.P.M.H. de, Kalucka, J., Sokol, L., et al. (2020). An Integrated Gene Expression Landscape Profiling Approach to Identify Lung Tumor Endothelial Cell Heterogeneity and Angiogenic Candidates. Cancer Cell 37, 21-36.e13.

```{r Goveia-MHC-II, fig.width=2.5, fig.height=2.5, cache=FALSE}
markers <- c('HLA-DMA', 'HLA-DMB', 'HLA-DOA', 'HLA-DOB', 'HLA-DPA1', 'HLA-DPB1', 'HLA-DQA1', 'HLA-DQA2', 'HLA-DQB1', 'HLA-DQB2', 'HLA-DRA', 'HLA-DRB1', 'HLA-DRB5', 'CD40', 'CD80', 'CD86')

ht <- customDoHeatmap(srat, markers, z_score = T, idents = "level_2", cluster_columns = FALSE)
p <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p
```

### Endothelia Cell Activation {.tabset}

Ref: [Mechanisms of endothelia Activation](https://link.springer.com/referenceworkentry/10.1007%2F978-0-387-84828-0_183)

### NF-κB

NF-κB as a Central Regulator of endothelia Cell Activation

```{r NF-κB, fig.width=5, fig.height=4}
CellChat::StackedVlnPlot(srat, features = c("NFKB1", "RELA", "RELB", "REL", "NFKB2"))
```

IκB

```{r IκB, fig.width=5, fig.height=3}
CellChat::StackedVlnPlot(srat, features = c("CHUK", "IKBKB", "IKBKG"))
```

### Leukocyte Adhesion Molecules, and Chemokines

- E-selectin is only expressed in activated endothelia cells

- In activated endothelium, E-selectin expression is largely confined to postcapillary venules (Aird, 2012)

- In “resting” endothelium, P-selectin is stored intracellularly in preexisting Weibel–Palade bodies and is expressed preferentially in postcapillary venules (Aird William C., 2007)

- Binding of endothelial E-selectin and P-selectin to leukocyte carbohydrate-based ligands promotes rolling of leukocytes to the endothelium

- Binding of endothelial cell adhesion molecules, including vascular cell adhesion molecule
(VCAM)-1 and intercellular adhesion molecule (ICAM)-1 to leukocyte integrins mediates firm adhesion.

- T cells for tumor immunotherapy was enhanced when proinflammatory agents that up-regulated ICAM and VCAM (Dudley, 2012)

```{r selectin, fig.width=5, fig.height=4}
CellChat::StackedVlnPlot(srat, features = c("SELE", "SELP", "ICAM1", "ICAM2", "VCAM1", "PECAM1"))
```

Chemokines (IL-6, IL-8, MCP-1)

```{r Chemokines, fig.width=4, fig.height=3}
CellChat::StackedVlnPlot(srat, features = c("IL6", "CXCL8", "CCL2"))
```


## Session info
