---
title: "00.Fig.myeloid"
---

```{r knitr, include = FALSE}
DOCNAME <- "00.Fig.myeloid"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = FALSE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(scales)
library(EnvStats)
library(ggplotify)
library(ggcorrplot)
library(ggpubr)
library(ggrepel)
library(ggrastr)

# patch
library(patchwork)
library(cowplot)
theme_set(theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
))
my_theme <- theme_cowplot(
  font_size = 6, # Overall font size
  rel_small = 6 / 6, # axis tick labels
  rel_tiny = 6 / 6, # caption
  rel_large = 6 / 6 # title
)
geom.text.size <- 6 / (14 / 5)

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# pvclust
library(pvclust)
library(dendextend)

# Seurat
library(Seurat)

# monocle
library(monocle)
```

Load code

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load rds

```{r load-rds, cache=FALSE}
srat <- readRDS(here::here("output/04.rm_cells/seurat_myeloid.rds"))
Idents(srat) <- "level_3"
srat
```

Load cell meta

```{r load-cell-meta, cache=FALSE, message=FALSE, warning=FALSE}
cell_meta <- read_csv(here::here("output/04.rm_cells/seurat_myeloid.cellmeta.csv")) %>%
  dplyr::select(barcode, seurat_clusters, level_1, level_2, level_3)
head(cell_meta)
```

Cell annotations

```{r anno, cache=FALSE}
df.c2c <- readRDS(here::here("data", "c2c.mye.rds"))
head(df.c2c)
```

## UMAP {.tabset}

### Level 3

```{r fig.width=3.5, fig.height=2.5, cache=FALSE}
p.umap.celltype <- DimPlot(srat, reduction = "umap", group.by = "level_3", raster = TRUE) +
  scale_color_manual(values = mye_level_3_color) +
  labs(title = NULL) +
  my_theme +
  # guides(color = guide_legend(ncol = 4, byrow = TRUE)) +
  theme(
    legend.margin = margin(rep(0, 4), unit = "cm"),
    legend.spacing.y = unit(0, "cm"),
    legend.key.height = unit(.4, "cm"),
    legend.box.margin = margin(rep(-.2, 4), unit = "cm"),
    legend.position = "right"
  )
p.umap.celltype

# saveRDS(p.umap.level_3, file = here::here("output", DOCNAME, "Fig.umap.cellType.rds"))
ggsave2(p.umap.celltype,
  filename = here::here("output", DOCNAME, "Fig.umap.cellType.pdf"),
  height = 2.5, width = 3.5, device = cairo_pdf
)
```

With label

```{r fig.width=3, fig.height=3, cache=FALSE}
p.umap.celltype.label <- DimPlot(srat,
  reduction = "umap", group.by = "level_3",
  label = TRUE, label.size = geom.text.size,
  raster = TRUE
) +
  scale_color_manual(values = mye_level_3_color) +
  labs(title = NULL) +
  my_theme +
  theme(legend.position = "none")
p.umap.celltype.label

ggsave2(
  filename = here::here("output", DOCNAME, "Fig.umap.cellType.labelled.pdf"),
  height = 2, width = 2, device = cairo_pdf
)
```

## Classification markers

```{r fig.width=5.5, fig.height=4, message=FALSE, warning=FALSE}
markers <- list(
  mono = filter_features(srat, c("FCN1", "VCAN", "S100A8")),
  mac = filter_features(srat, c("CD163", "C1QA")),
  cDC = filter_features(srat, c("CD1A", "CD1C", "CLEC10A")),
  tDC = filter_features(srat, c("LAMP3", "FSCN1", "IDO1")),
  pDC = filter_features(srat, c("ICOSLG", "LILRA4", "CLEC4C")),
  mast = filter_features(srat, c("MS4A2", "TPSAB1", "TPSB2")),
  CD34 = "CD34"
)

mat <- AverageExpression(srat,
  assays = "RNA", features = unlist(markers),
  slot = "data", verbose = F
)$RNA
mat <- log1p(as.matrix(mat))
# scale
mat_scale <- pheatmap:::scale_rows(mat)

# change order
mat_scale <- mat_scale[,c("cMono-C1", "cMono-C2", "cMono-C3", "cMono-C4", "cMono-C6", "ncMono-C5",
                          "MΦ-C1", "MΦ-C2", "MΦ-C3", "MΦ-C4", "MΦ-C5", "MΦ-C6", "MΦ-C7",
                          "cDC2-C1", "cDC2-C2", "cDC1-C5", "cDC1-C6", "tDC-C3",
                          "pDC-C4", "pDC-C7", "pDC-C8", "Mast-C1", "Mast-C2", "Mast-C3", "CD34+ cells")]

# color
x <- ceiling(max(abs(mat_scale)))
col_fun <- circlize::colorRamp2(c(-x, 0, x), c("#476fa9", "#ffffff", "#ca3226"))
legend_name <- "Exp."

ht <- ComplexHeatmap::Heatmap(mat_scale,
  name = legend_name,
  col = col_fun,
  cluster_columns = FALSE, cluster_rows = FALSE,
  show_row_names = TRUE, show_column_names = TRUE,
  row_names_side = "right",
  row_names_gp = gpar(fontsize = 6, fontface = "italic"),
  column_names_gp = gpar(fontsize = 6),
  column_names_rot = 60,
  row_split = factor(
    rep(names(markers), sapply(markers, length)),
    levels = names(markers)
  ),
  row_title_gp = gpar(fontsize = 6),
  column_split = factor(c(
    rep(c("Mono"), 6),
    rep(c("Mac"), 7),
    rep(c("DC"), 8),
    rep(c("Mast"), 3),
    "CD34"
  ),
  levels = c("Mono", "Mac", "DC", "Mast", "CD34")
  ),
  column_title = NULL,
  left_annotation = rowAnnotation(
    empty = anno_empty(
      border = FALSE,
      width = unit(1, "mm")
    ),
    foo = anno_block(
      gp = gpar(
        fill = "gray",
        col = "gray"
      ),
      width = unit(0.5, "mm")
    )
  ),
  heatmap_legend_param = list(
    title_position = "topcenter",
    direction = "vertical",
    labels_gp = gpar(fontsize = 6),
    title_gp = gpar(fontsize = 8),
    grid_height = unit(3, "mm"),
    grid_width = unit(3, "mm")
  )
)

# draw(ht)
p.markers.classification <- plot_grid(grid.grabExpr(draw(ht)))
p.markers.classification

# save
saveRDS(p.markers.classification, file = here::here("output", DOCNAME, "Fig.markers.heatmap.rds"))

ggsave2(p.markers.classification,
  filename = here::here("output", DOCNAME, "Fig.markers.heatmap.pdf"),
  height = 4, width = 5.5, device = cairo_pdf
)
```

Small

```{r fig.width=3, fig.height=1.8, message=FALSE, warning=FALSE}
ht <- ComplexHeatmap::Heatmap(mat_scale,
  name = legend_name,
  col = col_fun,
  cluster_columns = FALSE, cluster_rows = FALSE,
  show_row_names = TRUE, show_column_names = TRUE,
  row_names_side = "right",
  row_names_gp = gpar(fontsize = 4, fontface = "italic"),
  column_names_gp = gpar(fontsize = 4),
  column_names_rot = 60,
  row_split = factor(
    rep(names(markers), sapply(markers, length)),
    levels = names(markers)
  ),
  row_title_gp = gpar(fontsize = 4),
  column_split = factor(c(
    rep(c("Mono"), 6),
    rep(c("Mac"), 7),
    rep(c("DC"), 8),
    rep(c("Mast"), 3),
    "CD34"
  ),
  levels = c("Mono", "Mac", "DC", "Mast", "CD34")
  ),
  column_title = NULL,
  left_annotation = rowAnnotation(
    empty = anno_empty(
      border = FALSE,
      width = unit(1, "mm")
    ),
    foo = anno_block(
      gp = gpar(
        fill = "gray",
        col = "gray"
      ),
      width = unit(0.5, "mm")
    )
  ),
  heatmap_legend_param = list(
    title_position = "topcenter",
    direction = "vertical",
    labels_gp = gpar(fontsize = 4),
    title_gp = gpar(fontsize = 6),
    grid_height = unit(3, "mm"),
    grid_width = unit(3, "mm")
  )
)

# draw(ht)
p.markers.classification <- plot_grid(grid.grabExpr(draw(ht)))
p.markers.classification

# save
saveRDS(p.markers.classification, file = here::here("output", DOCNAME, "Fig.markers.heatmap.small.rds"))

ggsave2(p.markers.classification,
  filename = here::here("output", DOCNAME, "Fig.markers.heatmap.small.pdf"),
  height = 1.8, width = 3, device = cairo_pdf
)
```

Monocyte-derived Dendritic cells (Mo-DC)

```{r fig.width=5, fig.height=6}
DotPlot(srat, features = c("MS4A2", # FCYRI
                                      "CD14", 
                                      "CD1A", "CD1C", 
                                      "SIRPA", # CD172a 
                                      "MRC1" # CD206
                               )) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


## Roe Origin4

```{r fig.width=2, fig.height=3.6, warning=FALSE}
p.roe.mye = readRDS(here::here("output", "freq.enrichment_cellType3", "p.roe.mye.rds"))
p.roe.mye

ggsave2(p.roe.mye, filename = here::here("output", DOCNAME, "Fig.roe_mye.Origin4.pdf"),
        height = 3.6, width = 2)
```


## Roe Origin2_n (DC and Mac)

```{r load-meta, cache=FALSE, message=FALSE, warning=FALSE}
metadata <- read_csv(here::here("output/04.rm_cells/seurat.cellmeta.csv")) %>% 
  
sam_info <- read_csv(here::here("data/sample_info.3.csv"))
```

```{r fig.width=2, fig.height=2.8, warning=FALSE}
cluster_cellNum <- metadata %>%
  filter(cellType2 %in% c("MΦ", "DC")) %>%
  group_by(cellType3) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  filter(cellType2 %in% c("MΦ", "DC")) %>%
  group_by(cellType3, Origin2_n) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp <- setdiff(unique(metadata$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x <- tibble(cellType3 = unique(cluster_cellNum$cellType3), Origin2_n = t, o = 0)
    cluster_byOrigin2 <- bind_rows(cluster_byOrigin2, x)
  }
}

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin2_n") %>%
  left_join(cluster_cellNum, by = "cellType3") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType3, Origin2_n, r) %>%
  spread(Origin2_n, r, fill = 0) %>%
  column_to_rownames("cellType3")

# order
tmp.df <- r_oe %>%
  rownames_to_column("cellType3") %>%
  arrange(prePBMC, postPBMC, nLN, pLN, Normal, Adjacent, Tumor)

# data for plot
r_oe[r_oe > 10] <- 10
d4p <- r_oe %>%
  rownames_to_column("cellType3") %>%
  pivot_longer(cols = !cellType3, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "prePBMC", "postPBMC",
      "nLN", "pLN",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType3 = factor(cellType3, levels = tmp.df$cellType3)
  )

# % lineage prop
a <- metadata %>%
  filter(cellType2 %in% c("MΦ", "DC")) %>%
  group_by(cellType3, cellType, Origin2_n) %>%
  summarise(cellType3.cell_num = n()) %>%
  mutate(cellType3.cell_num = as.numeric(cellType3.cell_num))
b <- metadata %>%
  filter(cellType2 %in% c("MΦ", "DC")) %>%
  group_by(cellType) %>%
  summarise(cellType.cell_num = n()) %>%
  mutate(cellType.cell_num = as.numeric(cellType.cell_num))
df.lineage_prop <- a %>%
  left_join(b, by = "cellType") %>%
  mutate(lineage.prop = cellType3.cell_num * 100 / cellType.cell_num)

# plot
p.roe.mye <- ggplot(d4p, aes(x = Origin2_n, y = cellType3, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.lineage_prop,
    aes(x = Origin2_n, y = cellType3, size = lineage.prop),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "% lineage") +
  theme_cowplot(font_size = 6) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
p.roe.mye

ggsave2(p.roe.mye, filename = here::here("output", DOCNAME, "Fig.Roe_mye_mac_DC.pdf"),
        height = 2.8, width = 2)
```



## Composition

### cellType compos.

```{r fig.width=5, fig.height=3}
p1 <- srat@meta.data %>%
  group_by(level_3, Origin4) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100) %>%
  ggplot(aes(x = level_3, y = Percent, fill = Origin4)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = origin4_color_maps) +
  labs(fill = NULL) +
  coord_flip() +
  guides(fill = guide_legend(ncol = 1, byrow = TRUE)) +
  scale_y_continuous(expand = c(0, 0)) +
  my_theme
p2 <- srat@meta.data %>%
  group_by(level_3, Patient) %>%
  summarise(n = n()) %>%
  mutate(Percent = n / sum(n) * 100) %>%
  ggplot(aes(x = level_3, y = Percent, fill = Patient)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = patient_color_maps) +
  labs(fill = NULL) +
  coord_flip() +
  guides(fill = guide_legend(ncol = 1, byrow = TRUE)) +
  scale_y_continuous(expand = c(0, 0)) +
  my_theme +
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )
p3 <- srat@meta.data %>%
  group_by(level_3) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = level_3, y = log2(n), fill = "blue")) +
  geom_bar(stat = "identity", color = "#00008f") +
  scale_fill_manual(values = "#8f8fb1") +
  labs(fill = NULL) +
  coord_flip() +
  scale_y_continuous(expand = c(0, 0)) +
  my_theme +
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  )

p.stack <- plot_grid(p1, p2, p3, ncol = 3, rel_widths = c(2.1, 1.5, 1))
p.stack

# save
ggsave2(
  filename = here::here("output", DOCNAME, "Fig.cellTypeComposition.pdf"),
  height = 2, width = 5, device = cairo_pdf
)
```

## Selected genes

### Chemokine

Markers: chemokine, list from: <https://en.wikipedia.org/wiki/Chemokine>

```{r Chemokine, fig.width=3.5, fig.height=2}
markers = list(
  homeostatic = filter_features(srat, c('CCL14', 'CCL19', 'CCL20', 'CCL21', 'CXCL12', 'CXCL13')),
  inflammatory = filter_features(srat, c('CXCL8', 'CCL2', 'CCL3', 'CCL4', 'CCL5', 'CCL11', 'CXCL10'))
  )

ht <- customDoHeatmap(srat, markers, z_score = T, idents = "level_3", cluster_columns = FALSE)
p <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p
```

### Cytokines

- List from table 1 of: Turner, M.D., Nedjai, B., Hurst, T., and Pennington, D.J. (2014). Cytokines and chemokines: At the crossroads of cell signalling and inflammatory disease. Biochimica et Biophysica Acta (BBA) - Molecular Cell Research 1843, 2563–2582.

```{r Cytokines, fig.width=3.5, fig.height=4}
markers = list(
  # Adaptive immunity
  adaptive = filter_features(srat, c('IL2', 'IL4', 'IL7', 'IL15', 'IL21', 'CSF2', 'IL13', 'TSLP')),
  # Pro-inflammatory signalling
  proinflammatory = filter_features(srat, c('IL12A', 'IL12B', 'IL23A', 'IL27', 'EBI3', 'IL10', 'IL19', 'IL20', 'IL22', 'IL24', 'IL26')),
  # Anti-inflammatory signalling
  antiinflammatory = filter_features(srat, c('IL1A', 'IL1B', 'IL1RN', 'IL18', 'IL33', 'IL36A', 'IL36G', 'IL36RN', 'IL6', 'IL31', 'CNTF', 'CTF1', 'LIF', 'SPP1', 'OSM', 'TNF', 'BAFF', 'TNFSF13', 'IL17A', 'IFNB1', 'IFNK', 'IFNG'))
  )

ht <- customDoHeatmap(srat, markers, z_score = T, idents = "level_3", cluster_columns = FALSE)
p <- plot_grid(grid.grabExpr(draw(ht, heatmap_legend_side = "right")))
p
```










## Session info
