---
title: "01.clean_2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r knitr, include = FALSE}
DOCNAME = "01.clean_2"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = TRUE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# ggplot
library(cowplot)
library(ggpubr)
library(ggcorrplot)
library(ggrepel)
theme_set(theme_cowplot())

# heatmap
library(pheatmap)

# color
library(ggsci)
library(RColorBrewer)

# single cell
library(Matrix)
library(Seurat)
library(SingleCellExperiment)
library(scater)
library(scran)

# doublets
library(scds)

# soupx
library(SoupX)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/plot.R"))
```

# 01.clean {.tabset}

这个脚本用于清洗数据
* 过滤低质量细胞
* 过滤双细胞
* 使用 SoupX 矫正 count
* 使用矫正后的 count 构建 Seurat 对象，修改细胞名字，并且保存到磁盘

```{r set-environment, cache=FALSE}
cellranger_dir = '/niuyw-usb-disk/Projects/scESCA/cellranger/cellranger_count_outs'

# calculate confounders
.confounder_cal <- function(seurat_obj) {
  seurat_obj[["percent.mt"]] <- PercentageFeatureSet(seurat_obj, pattern = "^MT-")
  seurat_obj[["percent.heat"]] <- PercentageFeatureSet(seurat_obj, features = filter_features(seurat_obj, heat_genes.hs))
  seurat_obj[["percent.dissociation"]] <- PercentageFeatureSet(seurat_obj, features = filter_features(seurat_obj, dissocitaion_genes.hs))
  RPS.genes <- grep(pattern = "^RPS", x = rownames(seurat_obj), value = TRUE)
  RPL.genes <- grep(pattern = "^RPL", x = rownames(seurat_obj), value = TRUE)
  ribo.genes = c(RPS.genes, RPL.genes)
  seurat_obj[["percent.ribo"]] <- PercentageFeatureSet(seurat_obj, features = ribo.genes)
  return (seurat_obj)
}

# build seurat obj
.build_seurat <- function(filtered_count_dir, sam) {
  # seurat
  seurat.data = Read10X(data.dir = filtered_count_dir)
  seurat.raw = CreateSeuratObject(counts = seurat.data)
  seurat.raw = .confounder_cal(seurat_obj = seurat.raw)
  seurat.raw$Source = str_replace(sam, '_', '.')
  return(seurat.raw)
}

# filter cells
.filter_cells <- function(seurat_obj){
  # filter low-quality cells
  reasons = lowQuality_filter(seurat_obj, gene_fixed_low = 500, gene_fixed_high = 3000)
  seurat = seurat_obj[,!reasons$discard]
  
  # filter doublets
  seurat = .doublet_scds(seurat)
  high_scds = seurat$high_scds
  print(table(high_scds))
  seurat = subset(seurat, subset = high_scds == FALSE)
  return(seurat)
}

# soupx
.soupx <- function(raw_count_dir, seurat_obj) {
  seurat = fast_cluster(seurat_obj, res = 3)
  tod = Read10X(data.dir = raw_count_dir)
  toc = GetAssayData(seurat, assay = 'RNA', slot = 'counts')
  sc = SoupChannel(tod, toc, keepDroplets = TRUE)
  
  DR <- FetchData(seurat,
                  vars = c("seurat_clusters", "UMAP_1", "UMAP_2")) %>%
    dplyr::select(RD1 = UMAP_1, RD2 = UMAP_2, Cluster = seurat_clusters)
  
  # profiling the soup
  sc = estimateSoup(sc)
  # add dimenstion-reduction
  sc = setDR(sc, DR)
  # add cluster info
  sc = setClusters(sc, DR$Cluster)
  return(list(sc = sc, DR = DR))
}
```

## Use IG genes for SoupX {.tabset}

```{r clean-each-1, results = "hide"}
samples = c('S0619_P1', 'S0619_LN1', 'S0619_LN2', 'S0619_LN3', 'S0619_LN4', 'S0619_LN5', 'S0619_LN6', 'S0619_N3', 'S0619_N4', 'S0619_N5', 'S0730_P1', 'S0730_P2', 'S0730_LN1', 'S0730_LN2', 'S0730_LN3', 'S0730_LN4', 'S0730_N3', 'S0730_N4', 'S0730_N5', 'S0819_P1', 'S0819_LN1', 'S0819_LN2', 'S0819_LN3', 'S0819_LN4', 'S0819_LN5', 'S0819_N1', 'S0819_N2', 'S0819_N3', 'S0819_N5', 'S0920_P1', 'S0920_P2', 'S0920_LN1', 'S0920_LN2', 'S0920_LN3', 'S0920_LN4', 'S0920_N3', 'S0920_N4', 'S0920_N5', 'S1125_P2', 'S1125_LN1', 'S1125_LN2', 'S1125_LN3', 'S1125_LN4', 'S1125_LN5', 'S1125_LN6', 'S1125_N2', 'S1125_N3', 'S1125_N4', 'S1125_N5', 'S1204_P1', 'S1204_P2', 'S1204_LN1', 'S1204_LN2', 'S1204_LN3', 'S1204_LN5', 'S1204_LN6', 'S1204_LN7', 'S1204_N1', 'S1204_N2', 'S1204_N3', 'S1204_N4')

src_list <- lapply(samples, function(sam){
    src <- c(
        "### {{sam}} {.unnumbered}",
        "```{r {{sam}}-filter}",
        "sam = '{{sam}}'",
        "data_dir = file.path(cellranger_dir, sam)",
        "raw_count_dir = file.path(data_dir, 'raw_feature_bc_matrix')",
        "filtered_count_dir = file.path(data_dir, 'filtered_feature_bc_matrix')",
        "# prepare for soupx",
        "seurat.raw = .build_seurat(filtered_count_dir, sam)",
        "seurat = .filter_cells(seurat.raw)",
        "seurat",
        "```",
        "",
        "#### No SoupX",
        "```{r {{sam}}-no-1}",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-no-2, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "",
        "#### SoupX",
        "",
        "```{r {{sam}}-soupx}",
        "tmp = .soupx(raw_count_dir, seurat)",
        "sc = tmp$sc",
        "DR = tmp$DR",
        "plotMarkerDistribution(sc)",
        "```",
        "",
        "```{r {{sam}}-soupx-plotMarkerMap}",
        "useToEst = estimateNonExpressingCells(sc, nonExpressedGeneList = list(IG = soupx_genes$ig))",
        "plotMarkerMap(sc, geneSet = soupx_genes$ig, DR = DR, useToEst = useToEst)",
        "```",
        "",
        "```{r {{sam}}-soupx-cal}",
        "sc = calculateContaminationFraction(sc, list(IG = soupx_genes$ig), useToEst = useToEst)",
        "mean(sc$metaData$rho)",
        "```",
        "",
        "```{r {{sam}}-seurat-DimPlot}",
        "data_dir = file.path('/niuyw-usb-disk/Projects/scESCA/cellranger/soupx', sam)",
        "seurat = readRDS(file.path(data_dir, 'seurat.rds'))",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-plot-markers, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Use epithelia genes for SoupX {.tabset}

```{r clean-each-2, results = "hide"}
samples = c('S0619_N2', 'S0730_N1', 'S0730_N2', 'S0920_N1', 'S0920_N2', 'S1125_N1')

src_list <- lapply(samples, function(sam){
    src <- c(
        "### {{sam}} {.unnumbered}",
        "```{r {{sam}}-filter}",
        "sam = '{{sam}}'",
        "data_dir = file.path(cellranger_dir, sam)",
        "raw_count_dir = file.path(data_dir, 'raw_feature_bc_matrix')",
        "filtered_count_dir = file.path(data_dir, 'filtered_feature_bc_matrix')",
        "# prepare for soupx",
        "seurat.raw = .build_seurat(filtered_count_dir, sam)",
        "seurat = .filter_cells(seurat.raw)",
        "seurat",
        "```",
        "",
        "#### No SoupX",
        "```{r {{sam}}-no-1}",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-no-2, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "",
        "#### SoupX",
        "",
        "```{r {{sam}}-soupx}",
        "tmp = .soupx(raw_count_dir, seurat)",
        "sc = tmp$sc",
        "DR = tmp$DR",
        "plotMarkerDistribution(sc)",
        "```",
        "",
        "```{r {{sam}}-soupx-plotMarkerMap}",
        "useToEst = estimateNonExpressingCells(sc, nonExpressedGeneList = list(IG = soupx_genes$epithelia))",
        "plotMarkerMap(sc, geneSet = soupx_genes$epithelia, DR = DR, useToEst = useToEst)",
        "```",
        "",
        "```{r {{sam}}-soupx-cal}",
        "sc = calculateContaminationFraction(sc, list(IG = soupx_genes$epithelia), useToEst = useToEst)",
        "mean(sc$metaData$rho)",
        "```",
        "",
        "```{r {{sam}}-seurat-DimPlot}",
        "data_dir = file.path('/niuyw-usb-disk/Projects/scESCA/cellranger/soupx', sam)",
        "seurat = readRDS(file.path(data_dir, 'seurat.rds'))",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-plot-markers, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Use platelet genes for SoupX {.tabset}

```{r clean-each-3, results = "hide"}
samples = c('S0619_P2', 'S0819_P2')

src_list <- lapply(samples, function(sam){
    src <- c(
        "### {{sam}} {.unnumbered}",
        "```{r {{sam}}-filter}",
        "sam = '{{sam}}'",
        "data_dir = file.path(cellranger_dir, sam)",
        "raw_count_dir = file.path(data_dir, 'raw_feature_bc_matrix')",
        "filtered_count_dir = file.path(data_dir, 'filtered_feature_bc_matrix')",
        "# prepare for soupx",
        "seurat.raw = .build_seurat(filtered_count_dir, sam)",
        "seurat = .filter_cells(seurat.raw)",
        "seurat",
        "```",
        "",
        "#### No SoupX",
        "```{r {{sam}}-no-1}",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-no-2, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "",
        "#### SoupX",
        "",
        "```{r {{sam}}-soupx}",
        "tmp = .soupx(raw_count_dir, seurat)",
        "sc = tmp$sc",
        "DR = tmp$DR",
        "plotMarkerDistribution(sc)",
        "```",
        "",
        "```{r {{sam}}-soupx-plotMarkerMap}",
        "useToEst = estimateNonExpressingCells(sc, nonExpressedGeneList = list(IG = soupx_genes$platelet))",
        "plotMarkerMap(sc, geneSet = soupx_genes$platelet, DR = DR, useToEst = useToEst)",
        "```",
        "",
        "```{r {{sam}}-soupx-cal}",
        "sc = calculateContaminationFraction(sc, list(IG = soupx_genes$platelet), useToEst = useToEst)",
        "mean(sc$metaData$rho)",
        "```",
        "",
        "```{r {{sam}}-seurat-DimPlot}",
        "data_dir = file.path('/niuyw-usb-disk/Projects/scESCA/cellranger/soupx', sam)",
        "seurat = readRDS(file.path(data_dir, 'seurat.rds'))",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-plot-markers, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Use mast cell genes for SoupX {.tabset}

```{r clean-each-4, results = "hide"}
samples = c('S0619_N1', 'S1204_N5')

src_list <- lapply(samples, function(sam){
    src <- c(
        "### {{sam}} {.unnumbered}",
        "```{r {{sam}}-filter}",
        "sam = '{{sam}}'",
        "data_dir = file.path(cellranger_dir, sam)",
        "raw_count_dir = file.path(data_dir, 'raw_feature_bc_matrix')",
        "filtered_count_dir = file.path(data_dir, 'filtered_feature_bc_matrix')",
        "# prepare for soupx",
        "seurat.raw = .build_seurat(filtered_count_dir, sam)",
        "seurat = .filter_cells(seurat.raw)",
        "seurat",
        "```",
        "",
        "#### No SoupX",
        "```{r {{sam}}-no-1}",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-no-2, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "",
        "#### SoupX",
        "",
        "```{r {{sam}}-soupx}",
        "tmp = .soupx(raw_count_dir, seurat)",
        "sc = tmp$sc",
        "DR = tmp$DR",
        "plotMarkerDistribution(sc)",
        "```",
        "",
        "```{r {{sam}}-soupx-plotMarkerMap}",
        "useToEst = estimateNonExpressingCells(sc, nonExpressedGeneList = list(IG = soupx_genes$mast))",
        "plotMarkerMap(sc, geneSet = soupx_genes$mast, DR = DR, useToEst = useToEst)",
        "```",
        "",
        "```{r {{sam}}-soupx-cal}",
        "sc = calculateContaminationFraction(sc, list(IG = soupx_genes$mast), useToEst = useToEst)",
        "mean(sc$metaData$rho)",
        "```",
        "",
        "```{r {{sam}}-seurat-DimPlot}",
        "data_dir = file.path('/niuyw-usb-disk/Projects/scESCA/cellranger/soupx', sam)",
        "seurat = readRDS(file.path(data_dir, 'seurat.rds'))",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-plot-markers, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`


## Use fibroblasts cell genes for SoupX {.tabset}

```{r clean-each-5, results = "hide"}
samples = c('S0819_N4')

src_list <- lapply(samples, function(sam){
    src <- c(
        "### {{sam}} {.unnumbered}",
        "```{r {{sam}}-filter}",
        "sam = '{{sam}}'",
        "data_dir = file.path(cellranger_dir, sam)",
        "raw_count_dir = file.path(data_dir, 'raw_feature_bc_matrix')",
        "filtered_count_dir = file.path(data_dir, 'filtered_feature_bc_matrix')",
        "# prepare for soupx",
        "seurat.raw = .build_seurat(filtered_count_dir, sam)",
        "seurat = .filter_cells(seurat.raw)",
        "seurat",
        "```",
        "",
        "#### No SoupX",
        "```{r {{sam}}-no-1}",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-no-2, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "",
        "#### SoupX",
        "",
        "```{r {{sam}}-soupx}",
        "tmp = .soupx(raw_count_dir, seurat)",
        "sc = tmp$sc",
        "DR = tmp$DR",
        "plotMarkerDistribution(sc)",
        "```",
        "",
        "```{r {{sam}}-soupx-plotMarkerMap}",
        "useToEst = estimateNonExpressingCells(sc, nonExpressedGeneList = list(IG = soupx_genes$fibroblasts))",
        "plotMarkerMap(sc, geneSet = soupx_genes$fibroblasts, DR = DR, useToEst = useToEst)",
        "```",
        "",
        "```{r {{sam}}-soupx-cal}",
        "sc = calculateContaminationFraction(sc, list(IG = soupx_genes$fibroblasts), useToEst = useToEst)",
        "mean(sc$metaData$rho)",
        "```",
        "",
        "```{r {{sam}}-seurat-DimPlot}",
        "data_dir = file.path('/niuyw-usb-disk/Projects/scESCA/cellranger/soupx', sam)",
        "seurat = readRDS(file.path(data_dir, 'seurat.rds'))",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-plot-markers, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Use macrophage cell genes for SoupX {.tabset}

```{r clean-each-6, results = "hide"}
samples = c('S1125_P1')

src_list <- lapply(samples, function(sam){
    src <- c(
        "### {{sam}} {.unnumbered}",
        "```{r {{sam}}-filter}",
        "sam = '{{sam}}'",
        "data_dir = file.path(cellranger_dir, sam)",
        "raw_count_dir = file.path(data_dir, 'raw_feature_bc_matrix')",
        "filtered_count_dir = file.path(data_dir, 'filtered_feature_bc_matrix')",
        "# prepare for soupx",
        "seurat.raw = .build_seurat(filtered_count_dir, sam)",
        "seurat = .filter_cells(seurat.raw)",
        "seurat",
        "```",
        "",
        "#### No SoupX",
        "```{r {{sam}}-no-1}",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-no-2, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "",
        "#### SoupX",
        "",
        "```{r {{sam}}-soupx}",
        "tmp = .soupx(raw_count_dir, seurat)",
        "sc = tmp$sc",
        "DR = tmp$DR",
        "plotMarkerDistribution(sc)",
        "```",
        "",
        "```{r {{sam}}-soupx-plotMarkerMap}",
        "useToEst = estimateNonExpressingCells(sc, nonExpressedGeneList = list(IG = soupx_genes$mp))",
        "plotMarkerMap(sc, geneSet = soupx_genes$mp, DR = DR, useToEst = useToEst)",
        "```",
        "",
        "```{r {{sam}}-soupx-cal}",
        "sc = calculateContaminationFraction(sc, list(IG = soupx_genes$mp), useToEst = useToEst)",
        "mean(sc$metaData$rho)",
        "```",
        "",
        "```{r {{sam}}-seurat-DimPlot}",
        "data_dir = file.path('/niuyw-usb-disk/Projects/scESCA/cellranger/soupx', sam)",
        "seurat = readRDS(file.path(data_dir, 'seurat.rds'))",
        "seurat = fast_cluster(seurat)",
        "DimPlot(seurat, reduction = 'umap', label = T, pt.size = 0.1)",
        "```",
        "```{r {{sam}}-plot-markers, fig.width=12, fig.height=6}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 3, group.by = 'seurat_clusters') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        "")
    #print(src)
    knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

