---
title: "exp_similarity.cross_tissue.Origin2"
---

```{r knitr, include = FALSE}
DOCNAME = "exp_similarity.cross_tissue.Origin2"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

根据低维 PC 空间中细胞的临近，来判断不同组织中细胞与其他组织的相似性。

参考下面文章中的 Figure2

* Zhang, Q., He, Y., Luo, N., Patel, S. J., Han, Y., Gao, R., Modak, M., Carotta, S., Haslinger, C., Kind, D., Peet, G. W., Zhong, G., Lu, S., Zhu, W., Mao, Y., Xiao, M., Bergmann, M., Hu, X., Kerkar, S. P., … Zhang, Z. (2019). Landscape and Dynamics of Single Immune Cells in Hepatocellular Carcinoma. Cell, 179(4), 829-845.e20. https://doi.org/10.1016/j.cell.2019.10.003


## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(ggplotify)
library(ggcorrplot)
library(ggpubr)

# patch
library(patchwork)
library(cowplot)
theme_set(theme_cowplot(font_size = 10,
                        rel_small = 8/10,
                        rel_tiny = 6/10,
                        rel_large = 10/10,
                        font_family = "Arial"))
# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# BiocNeighbors
library(BiocNeighbors)

# Seurat
library(Seurat)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/glm.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

## prePBMC {.tabset}

Load

```{r}
t = "prePBMC"
p_mat = readRDS(here::here("output", DOCNAME, paste0(t, ".p_mat.rds")))
p_mat = p_mat[!str_detect(rownames(p_mat), "Epi_"),]
tissue_origin = readRDS(here::here("output", DOCNAME, paste0(t, ".tissue_origin.rds")))
tissue_origin = tissue_origin[!str_detect(tissue_origin$NN_type, "Epi_"),]
```

### Overall

```{r fig.width=5, fig.height=4}
# Generate barplot of pecent rate
plot_sum <- t(as.data.frame(prop.table(table(tissue_origin$NN_tissue))*100))
plot_sum <- as.data.frame(t(plot_sum))
colnames(plot_sum) <- c("Tissue", "Percent")
plot_sum$Percent <- as.numeric(as.character(plot_sum$Percent))
plot_sum$Tissue <- factor(plot_sum$Tissue, levels = c("prePBMC", "postPBMC", "nLN", "m_nLN", "m_pLN",
                                                "Normal", "Adjacent", "Tumor"))
p <- ggplot(data=plot_sum, aes(x=Tissue, y=Percent)) +
    geom_bar(stat="identity", fill="steelblue") +
    scale_y_continuous(expand = c(0, 0))
p
```

### p_mat

```{r fig.width=4, fig.height=15}
pheatmap::pheatmap(p_mat, cluster_rows = FALSE, cluster_cols = FALSE,
                   display_numbers = TRUE,
                   color=colorRampPalette(c("red", "white", "blue"))(100),
                   fontsize = 7)
```

### By cellType3

```{r fig.width=16, fig.height=5}
d4p = tissue_origin %>%
  group_by(cellType3, NN_tissue) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
# sort
tmp_df = d4p %>%
  filter(NN_tissue == "Tumor") %>%
  arrange(-Percent)
d4p$cellType3 = factor(d4p$cellType3, levels = c(tmp_df$cellType3, setdiff(d4p$cellType3, tmp_df$cellType3)))

d4p %>%
  ggplot(aes(x = cellType3, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# save
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "Percent") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.percent.txt")))
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "n") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.number.txt")))
```

### Facet by cellType3

```{r fig.width=12, fig.height=20}
d4p %>%
  ggplot(aes(x = NN_tissue, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(.~cellType3) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## postPBMC {.tabset}

Load

```{r}
t = "postPBMC"
p_mat = readRDS(here::here("output", DOCNAME, paste0(t, ".p_mat.rds")))
p_mat = p_mat[!str_detect(rownames(p_mat), "Epi_"),]
tissue_origin = readRDS(here::here("output", DOCNAME, paste0(t, ".tissue_origin.rds")))
tissue_origin = tissue_origin[!str_detect(tissue_origin$NN_type, "Epi_"),]
```

### Overall

```{r fig.width=5, fig.height=4}
# Generate barplot of pecent rate
plot_sum <- t(as.data.frame(prop.table(table(tissue_origin$NN_tissue))*100))
plot_sum <- as.data.frame(t(plot_sum))
colnames(plot_sum) <- c("Tissue", "Percent")
plot_sum$Percent <- as.numeric(as.character(plot_sum$Percent))
plot_sum$Tissue <- factor(plot_sum$Tissue, levels = c("prePBMC", "postPBMC", "nLN", "m_nLN", "m_pLN",
                                                "Normal", "Adjacent", "Tumor"))
p <- ggplot(data=plot_sum, aes(x=Tissue, y=Percent)) +
    geom_bar(stat="identity", fill="steelblue") +
    scale_y_continuous(expand = c(0, 0))
p
```

### p_mat

```{r fig.width=4, fig.height=15}
pheatmap::pheatmap(p_mat, cluster_rows = FALSE, cluster_cols = FALSE,
                   display_numbers = TRUE,
                   color=colorRampPalette(c("red", "white", "blue"))(100),
                   fontsize = 7)
```

### By cellType3

```{r fig.width=16, fig.height=5}
d4p = tissue_origin %>%
  group_by(cellType3, NN_tissue) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
# sort
tmp_df = d4p %>%
  filter(NN_tissue == "Tumor") %>%
  arrange(-Percent)
d4p$cellType3 = factor(d4p$cellType3, levels = c(tmp_df$cellType3, setdiff(d4p$cellType3, tmp_df$cellType3)))

d4p %>%
  ggplot(aes(x = cellType3, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# save
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "Percent") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.percent.txt")))
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "n") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.number.txt")))
```

### Facet by cellType3

```{r fig.width=12, fig.height=20}
d4p %>%
  ggplot(aes(x = NN_tissue, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(.~cellType3) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## nLN {.tabset}

Load

```{r}
t = "nLN"
p_mat = readRDS(here::here("output", DOCNAME, paste0(t, ".p_mat.rds")))
p_mat = p_mat[!str_detect(rownames(p_mat), "Epi_"),]
tissue_origin = readRDS(here::here("output", DOCNAME, paste0(t, ".tissue_origin.rds")))
tissue_origin = tissue_origin[!str_detect(tissue_origin$NN_type, "Epi_"),]
```

### Overall

```{r fig.width=5, fig.height=4}
# Generate barplot of pecent rate
plot_sum <- t(as.data.frame(prop.table(table(tissue_origin$NN_tissue))*100))
plot_sum <- as.data.frame(t(plot_sum))
colnames(plot_sum) <- c("Tissue", "Percent")
plot_sum$Percent <- as.numeric(as.character(plot_sum$Percent))
plot_sum$Tissue <- factor(plot_sum$Tissue, levels = c("prePBMC", "postPBMC", "nLN", "m_nLN", "m_pLN",
                                                "Normal", "Adjacent", "Tumor"))
p <- ggplot(data=plot_sum, aes(x=Tissue, y=Percent)) +
    geom_bar(stat="identity", fill="steelblue") +
    scale_y_continuous(expand = c(0, 0))
p
```

### p_mat

```{r fig.width=4, fig.height=15}
pheatmap::pheatmap(p_mat, cluster_rows = FALSE, cluster_cols = FALSE,
                   display_numbers = TRUE,
                   color=colorRampPalette(c("red", "white", "blue"))(100),
                   fontsize = 7)
```

### By cellType3

```{r fig.width=16, fig.height=5}
d4p = tissue_origin %>%
  group_by(cellType3, NN_tissue) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
# sort
tmp_df = d4p %>%
  filter(NN_tissue == "Tumor") %>%
  arrange(-Percent)
d4p$cellType3 = factor(d4p$cellType3, levels = c(tmp_df$cellType3, setdiff(d4p$cellType3, tmp_df$cellType3)))

d4p %>%
  ggplot(aes(x = cellType3, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# save
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "Percent") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.percent.txt")))
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "n") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.number.txt")))
```

### Facet by cellType3

```{r fig.width=12, fig.height=20}
d4p %>%
  ggplot(aes(x = NN_tissue, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(.~cellType3) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


## pLN {.tabset}

Load

```{r}
t = "pLN"
p_mat = readRDS(here::here("output", DOCNAME, paste0(t, ".p_mat.rds")))
p_mat = p_mat[!str_detect(rownames(p_mat), "Epi_"),]
tissue_origin = readRDS(here::here("output", DOCNAME, paste0(t, ".tissue_origin.rds")))
tissue_origin = tissue_origin[!str_detect(tissue_origin$NN_type, "Epi_"),]
```

### Overall

```{r fig.width=5, fig.height=4}
# Generate barplot of pecent rate
plot_sum <- t(as.data.frame(prop.table(table(tissue_origin$NN_tissue))*100))
plot_sum <- as.data.frame(t(plot_sum))
colnames(plot_sum) <- c("Tissue", "Percent")
plot_sum$Percent <- as.numeric(as.character(plot_sum$Percent))
plot_sum$Tissue <- factor(plot_sum$Tissue, levels = c("prePBMC", "postPBMC", "nLN", "m_nLN", "m_pLN",
                                                "Normal", "Adjacent", "Tumor"))
p <- ggplot(data=plot_sum, aes(x=Tissue, y=Percent)) +
    geom_bar(stat="identity", fill="steelblue") +
    scale_y_continuous(expand = c(0, 0))
p
```

### p_mat

```{r fig.width=4, fig.height=15}
pheatmap::pheatmap(p_mat, cluster_rows = FALSE, cluster_cols = FALSE,
                   display_numbers = TRUE,
                   color=colorRampPalette(c("red", "white", "blue"))(100),
                   fontsize = 7)
```

### By cellType3

```{r fig.width=16, fig.height=5}
d4p = tissue_origin %>%
  group_by(cellType3, NN_tissue) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
# sort
tmp_df = d4p %>%
  filter(NN_tissue == "Tumor") %>%
  arrange(-Percent)
d4p$cellType3 = factor(d4p$cellType3, levels = c(tmp_df$cellType3, setdiff(d4p$cellType3, tmp_df$cellType3)))

d4p %>%
  ggplot(aes(x = cellType3, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# save
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "Percent") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.percent.txt")))
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "n") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.number.txt")))
```

### Facet by cellType3

```{r fig.width=12, fig.height=20}
d4p %>%
  ggplot(aes(x = NN_tissue, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(.~cellType3) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## Normal {.tabset}

Load

```{r}
t = "Normal"
p_mat = readRDS(here::here("output", DOCNAME, paste0(t, ".p_mat.rds")))
p_mat = p_mat[!str_detect(rownames(p_mat), "Epi_"),]
tissue_origin = readRDS(here::here("output", DOCNAME, paste0(t, ".tissue_origin.rds")))
tissue_origin = tissue_origin[!str_detect(tissue_origin$NN_type, "Epi_"),]
```

### Overall

```{r fig.width=5, fig.height=4}
# Generate barplot of pecent rate
plot_sum <- t(as.data.frame(prop.table(table(tissue_origin$NN_tissue))*100))
plot_sum <- as.data.frame(t(plot_sum))
colnames(plot_sum) <- c("Tissue", "Percent")
plot_sum$Percent <- as.numeric(as.character(plot_sum$Percent))
plot_sum$Tissue <- factor(plot_sum$Tissue, levels = c("prePBMC", "postPBMC", "nLN", "m_nLN", "m_pLN",
                                                "Normal", "Adjacent", "Tumor"))
p <- ggplot(data=plot_sum, aes(x=Tissue, y=Percent)) +
    geom_bar(stat="identity", fill="steelblue") +
    scale_y_continuous(expand = c(0, 0))
p
```

### p_mat

```{r fig.width=4, fig.height=15}
pheatmap::pheatmap(p_mat, cluster_rows = FALSE, cluster_cols = FALSE,
                   display_numbers = TRUE,
                   color=colorRampPalette(c("red", "white", "blue"))(100),
                   fontsize = 7)
```

### By cellType3

```{r fig.width=16, fig.height=5}
d4p = tissue_origin %>%
  group_by(cellType3, NN_tissue) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
# sort
tmp_df = d4p %>%
  filter(NN_tissue == "Tumor") %>%
  arrange(-Percent)
d4p$cellType3 = factor(d4p$cellType3, levels = c(tmp_df$cellType3, setdiff(d4p$cellType3, tmp_df$cellType3)))

d4p %>%
  ggplot(aes(x = cellType3, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# save
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "Percent") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.percent.txt")))
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "n") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.number.txt")))
```

### Facet by cellType3

```{r fig.width=12, fig.height=20}
d4p %>%
  ggplot(aes(x = NN_tissue, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(.~cellType3) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## Adjacent {.tabset}

Load

```{r}
t = "Adjacent"
p_mat = readRDS(here::here("output", DOCNAME, paste0(t, ".p_mat.rds")))
p_mat = p_mat[!str_detect(rownames(p_mat), "Epi_"),]
tissue_origin = readRDS(here::here("output", DOCNAME, paste0(t, ".tissue_origin.rds")))
tissue_origin = tissue_origin[!str_detect(tissue_origin$NN_type, "Epi_"),]
```

### Overall

```{r fig.width=5, fig.height=4}
# Generate barplot of pecent rate
plot_sum <- t(as.data.frame(prop.table(table(tissue_origin$NN_tissue))*100))
plot_sum <- as.data.frame(t(plot_sum))
colnames(plot_sum) <- c("Tissue", "Percent")
plot_sum$Percent <- as.numeric(as.character(plot_sum$Percent))
plot_sum$Tissue <- factor(plot_sum$Tissue, levels = c("prePBMC", "postPBMC", "nLN", "m_nLN", "m_pLN",
                                                "Normal", "Adjacent", "Tumor"))
p <- ggplot(data=plot_sum, aes(x=Tissue, y=Percent)) +
    geom_bar(stat="identity", fill="steelblue") +
    scale_y_continuous(expand = c(0, 0))
p
```

### p_mat

```{r fig.width=4, fig.height=15}
pheatmap::pheatmap(p_mat, cluster_rows = FALSE, cluster_cols = FALSE,
                   display_numbers = TRUE,
                   color=colorRampPalette(c("red", "white", "blue"))(100),
                   fontsize = 7)
```

### By cellType3

```{r fig.width=16, fig.height=5}
d4p = tissue_origin %>%
  group_by(cellType3, NN_tissue) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
# sort
tmp_df = d4p %>%
  filter(NN_tissue == "Tumor") %>%
  arrange(-Percent)
d4p$cellType3 = factor(d4p$cellType3, levels = c(tmp_df$cellType3, setdiff(d4p$cellType3, tmp_df$cellType3)))

d4p %>%
  ggplot(aes(x = cellType3, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# save
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "Percent") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.percent.txt")))
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "n") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.number.txt")))
```

### Facet by cellType3

```{r fig.width=12, fig.height=20}
d4p %>%
  ggplot(aes(x = NN_tissue, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(.~cellType3) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## Tumor {.tabset}

Load

```{r}
t = "Tumor"
p_mat = readRDS(here::here("output", DOCNAME, paste0(t, ".p_mat.rds")))
p_mat = p_mat[!str_detect(rownames(p_mat), "Epi_"),]
tissue_origin = readRDS(here::here("output", DOCNAME, paste0(t, ".tissue_origin.rds")))
tissue_origin = tissue_origin[!str_detect(tissue_origin$NN_type, "Epi_"),]
```

### Overall

```{r fig.width=5, fig.height=4}
# Generate barplot of pecent rate
plot_sum <- t(as.data.frame(prop.table(table(tissue_origin$NN_tissue))*100))
plot_sum <- as.data.frame(t(plot_sum))
colnames(plot_sum) <- c("Tissue", "Percent")
plot_sum$Percent <- as.numeric(as.character(plot_sum$Percent))
plot_sum$Tissue <- factor(plot_sum$Tissue, levels = c("prePBMC", "postPBMC", "nLN", "m_nLN", "m_pLN",
                                                "Normal", "Adjacent", "Tumor"))
p <- ggplot(data=plot_sum, aes(x=Tissue, y=Percent)) +
    geom_bar(stat="identity", fill="steelblue") +
    scale_y_continuous(expand = c(0, 0))
p
```

### p_mat

```{r fig.width=4, fig.height=15}
pheatmap::pheatmap(p_mat, cluster_rows = FALSE, cluster_cols = FALSE,
                   display_numbers = TRUE,
                   color=colorRampPalette(c("red", "white", "blue"))(100),
                   fontsize = 7)
```

### By cellType3

```{r fig.width=16, fig.height=5}
d4p = tissue_origin %>%
  group_by(cellType3, NN_tissue) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
# sort
tmp_df = d4p %>%
  filter(NN_tissue == "Tumor") %>%
  arrange(-Percent)
d4p$cellType3 = factor(d4p$cellType3, levels = c(tmp_df$cellType3, setdiff(d4p$cellType3, tmp_df$cellType3)))

d4p %>%
  ggplot(aes(x = cellType3, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# save
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "Percent") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.percent.txt")))
d4p %>% pivot_wider(id_cols = "cellType3", names_from = "NN_tissue", values_from = "n") %>%
  write_tsv(here::here("output", DOCNAME, paste0(t, ".exp_similarity.cross_tissue.number.txt")))
```

### Facet by cellType3

```{r fig.width=12, fig.height=20}
d4p %>%
  ggplot(aes(x = NN_tissue, y = Percent, fill = NN_tissue)) +
  geom_col() +
  scale_fill_manual(values = origin4_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(.~cellType3) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## Session info
