---
title: "03.sub_cluster.tcells"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r knitr, include = FALSE}
DOCNAME = "03.sub_cluster.tcells"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = TRUE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(cowplot)
library(pheatmap)
library(ggpubr)
library(ggcorrplot)
theme_set(theme_cowplot())
library(ggsci)
library(viridis)

# Seurat
library(Seurat)

# clustree
library(clustree)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/plot.R"))
```

Load rds

```{r load-rds, cache=FALSE}
seurat = readRDS(here::here('output/02.cell_type/seurat_tcells.rds'))
seurat
```

## sub-clustering

```{r sub-cluster, eval=FALSE}
# not run, slow
seurat = get_cluster(seurat)
seurat = seurat$seurat_obj
```

## Dimensionlity reduction {.tabset}

### UMAP {.tabset}

```{r umap-res, results = "hide"}
resolutions = seq(0, 2, 0.2)
src_list <- lapply(resolutions, function(res) {
    src <- c(
        "#### Res {{res}} {.unnumbered}",
        "```{r res-umap-{{res}}, fig.width=8, fig.height=8}",
        "DimPlot(seurat, reduction = 'umap', group.by = 'RNA_snn_res.{{res}}', label = TRUE, pt.size = 0.5)",
        "```",
        ""
    )
    knitr::knit_expand(text = src)
})
out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

### tSNE {.tabset}

```{r tsne-res, results = "hide"}
resolutions = seq(0, 2, 0.2)
src_list <- lapply(resolutions, function(res) {
    src <- c(
        "#### Res {{res}} {.unnumbered}",
        "```{r res-tsne-{{res}}, fig.width=8, fig.height=8}",
        "DimPlot(seurat, reduction = 'tsne', group.by = 'RNA_snn_res.{{res}}', label = TRUE, pt.size = 0.5)",
        "```",
        ""
    )
    knitr::knit_expand(text = src)
})
out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Expression of lineage markers {.tabset}

```{r markers-dot, results = "hide"}
resolutions = seq(0, 2, 0.2)
src_list <- lapply(resolutions, function(res) {
    src <- c(
        "### Res {{res}} {.unnumbered}",
        "```{r res-markers-dot-{{res}}, fig.width=12, fig.height=10}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 4, group.by = 'RNA_snn_res.{{res}}') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        ""
    )
    knitr::knit_expand(text = src)
})
out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Classification markers {.tabset}

```{r classify-t-nk, fig.width=12, fig.height=24}
FeaturePlot(seurat, features = unique(c(markers.tcells, markers.NK)), ncol = 3)
```

## Cluster compositions

Cluster compositions and cell number

```{r cluster-check, fig.height=12, fig.width=12, cache=FALSE}
res <- 2
Idents(seurat) <- paste0("RNA_snn_res.", res)
seurat$seurat_clusters <- Idents(seurat)
cluster_stack(seurat)
```

## Percent of VDJ

```{r percent-vdj, fig.width=8, fig.height=12}
percent_vdj(seurat)
```

## Specific cell number of each cluster {.tabset}

### Tissue

```{r cell-num-source-tissue}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Tissue) %>%
    summarise(n = n()),
  filter = 'top'
)
```

### Patient

```{r cell-num-source-patient}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Patient) %>%
    summarise(n = n()),
  filter = 'top'
)
```

### Origin

```{r cell-num-source-origin}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Origin) %>%
    summarise(n = n()),
  filter = 'top'
)
```

### Source

```{r cell-num-source-source}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Source) %>%
    summarise(n = n()),
  filter = 'top'
)
```

## Clustering trees {.tabset}

Clustering trees show the relationship between clusterings at adjacent resolutions. Each cluster is represented as a node in a graph and the edges show the overlap between clusters.

### Standard

Coloured by clustering resolution.

```{r clustree, fig.height=12, fig.width=12}
clustree(seurat, prefix = 'RNA_snn_res.')
```

### Stability

Coloured by the SC3 stability metric.

```{r clustree-stability, fig.height=12, fig.width=12}
clustree(seurat, node_colour = "sc3_stability")
```

### Genes {.tabset}

Coloured by the expression of known marker genes.

```{r known-genes}
known_genes <- c(
    # immune
    "PTPRC",
    # T
    "CD3D", "CD3G", "CD3E", "CD247",
    # CD4, CD8
    "CD4", "CD8A", "CD8B", 
    # Treg
    "FOXP3", "IL2RA",
    # MAIT
    "TRAV1-2",
    # gamma-delta T
    "TRDV1", "TRDV2", "TRGV9",
    # NKT/NK
    "GNLY", "GZMK", "KLRF1", "NKG7", "KLRB1"
    
)
is_present <- known_genes %in% rownames(seurat)
```

> The following genes aren't present in this dataset and will be skipped:
> `r known_genes[!is_present]`

```{r clustree-genes, results = "hide"}
src_list <- lapply(known_genes[is_present], function(gene) {
    src <- c("#### {{gene}} {.unnumbered}",
             "```{r clustree-{{gene}}, fig.height=12, fig.width=12}",
             "clustree(seurat, node_colour = '{{gene}}',",
                      "node_colour_aggr = 'mean', exprs = 'data') +",
             "scale_colour_viridis_c(option = 'plasma', begin = 0.3)",
             "```",
             "")
    knitr::knit_expand(text = src)
})
out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Session info

```{r session-info, cache=FALSE}
sessionInfo()
```
