---
title: "04.sub_cluster.fibroblasts.2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r knitr, include = FALSE}
DOCNAME = "04.sub_cluster.fibroblasts.2"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = TRUE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(cowplot)
library(pheatmap)
library(ComplexHeatmap)
library(ggpubr)
library(ggcorrplot)
theme_set(theme_cowplot())
library(ggsci)
library(viridis)

# Seurat
library(Seurat)

# clustree
library(clustree)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/plot.R"))
```

Load rds

```{r load-rds, cache=FALSE}
seurat = readRDS(here::here('output/04.rm_cells/seurat_fibroblasts.2.rds'))
seurat
```

## sub-clustering

```{r sub-cluster, eval=FALSE}
# not run, slow
seurat = get_cluster(seurat)
seurat = seurat$seurat_obj
```

## Dimensionlity reduction {.tabset}

### UMAP {.tabset}

```{r umap-res, results = "hide"}
resolutions = seq(0, 2, 0.2)
src_list <- lapply(resolutions, function(res) {
    src <- c(
        "#### Res {{res}} {.unnumbered}",
        "```{r res-umap-{{res}}, fig.width=8, fig.height=8}",
        "DimPlot(seurat, reduction = 'umap', group.by = 'RNA_snn_res.{{res}}', label = TRUE, pt.size = 0.5)",
        "```",
        ""
    )
    knitr::knit_expand(text = src)
})
out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

### tSNE {.tabset}

```{r tsne-res, results = "hide"}
resolutions = seq(0, 2, 0.2)
src_list <- lapply(resolutions, function(res) {
    src <- c(
        "#### Res {{res}} {.unnumbered}",
        "```{r res-tsne-{{res}}, fig.width=8, fig.height=8}",
        "DimPlot(seurat, reduction = 'tsne', group.by = 'RNA_snn_res.{{res}}', label = TRUE, pt.size = 0.5)",
        "```",
        ""
    )
    knitr::knit_expand(text = src)
})
out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Expression of lineage markers {.tabset}

```{r markers-dot, results = "hide"}
resolutions = seq(0, 2, 0.2)
src_list <- lapply(resolutions, function(res) {
    src <- c(
        "### Res {{res}} {.unnumbered}",
        "```{r res-markers-dot-{{res}}, fig.width=12, fig.height=10}",
        "DotPlot(seurat, features = rev(unlist(markers.to.plot)), dot.scale = 4, group.by = 'RNA_snn_res.{{res}}') + RotatedAxis() + theme(axis.text.x = element_text(size = rel(0.7)), axis.text.y = element_text(size = rel(0.9)), panel.grid = element_line(color = 'grey'), legend.position = 'bottom', legend.box = 'vertical')",
        "```",
        ""
    )
    knitr::knit_expand(text = src)
})
out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## iCAF, myCAF and apCAF

Ref:
* Elyada, E., Bolisetty, M., Laise, P., Flynn, W.F., Courtois, E.T., Burkhart, R.A., Teinor, J.A., Belleau, P., Biffi, G., Lucito, M.S., et al. (2019). Cross-Species Single-Cell Analysis of Pancreatic Ductal Adenocarcinoma Reveals Antigen-Presenting Cancer-Associated Fibroblasts. Cancer Discov 9, 1102–1123.

- myCAFs were primarily adjacent to cancer cells.
- iCAFs were located in the desmoplastic areas of the tumor, farther away from the cancer cells.
- apCAF expresses MHC class II (MHCII)–related genes and induces T-cell receptor (TCR) ligation in CD4+ T cells in an antigen-dependent manner

```{r type-of-fibroblasts, fig.height=6, fig.width=6}
markers = list(
  iCAF = filter_features(seurat, c('IL6', 'PDGFRA', 'CXCL12', 'CFD', 'DPT', 'AGTR1', 'HAS1', 'CXCL1', 'CXCL2', 'CCL2', 'CXCL8', 'LMNA', 'PLA2G2A', 'CLU', 'EMP1')),
  myCAF = filter_features(seurat, c('ACTA2', 'TAGLN', 'MMP11', 'HOPX', 'POSTN', 'TPM1', 'MYL9', 'TPM2')),
  apCAF = filter_features(seurat, c('CD74', 'HLA-DRA', 'HLA-DPA1', 'HLA-DQA1', 'SYTL1')),
  costimulatory = filter_features(seurat, c('KLRK1', 'ICOS', 'TNFRSF18', 'CD27', 'TNFRSF9', 'CD226', 'TNFRSF4', 'CD40', 'CD80', 'CD86'))
)

markers_aveByCluster = AverageExpression(seurat, assays = "RNA", features = unlist(markers), slot = "data", verbose = FALSE)
mat = pheatmap:::scale_rows(as.matrix(markers_aveByCluster$RNA))

x = ceiling(max(abs(mat)))
col_fun = circlize::colorRamp2(c(-x, 0, x), c("#3658d3", "#ffffff", "#900504"))

Heatmap(mat, name = 'Exp',
        col = col_fun,
        cluster_columns = F, cluster_rows = F,
        row_split = factor(rep(
          c('iCAF', 'myCAF', 'apCAF', 'costimulatory'), 
          sapply(markers, length)),
          levels = c('myCAF', 'iCAF', 'apCAF', 'costimulatory'))
        )
```

## Cluster compositions

Cluster compositions and cell number

```{r cluster-check, fig.height=12, fig.width=12, cache=FALSE}
res <- 2
Idents(seurat) <- paste0("RNA_snn_res.", res)
seurat$seurat_clusters <- Idents(seurat)
cluster_stack(seurat)
```

## Percent of VDJ

```{r percent-vdj, fig.width=8, fig.height=10}
percent_vdj(seurat)
```

## Specific cell number of each cluster {.tabset}

### Tissue

```{r cell-num-source-tissue}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Tissue) %>%
    summarise(n = n()),
  filter = 'top'
)
```

### Patient

```{r cell-num-source-patient}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Patient) %>%
    summarise(n = n()),
  filter = 'top'
)
```

### Origin

```{r cell-num-source-origin}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Origin) %>%
    summarise(n = n()),
  filter = 'top'
)
```

### Source

```{r cell-num-source-source}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Source) %>%
    summarise(n = n()),
  filter = 'top'
)
```

## Clustering trees {.tabset}

Clustering trees show the relationship between clusterings at adjacent resolutions. Each cluster is represented as a node in a graph and the edges show the overlap between clusters.

### Standard

Coloured by clustering resolution.

```{r clustree, fig.height=12, fig.width=12}
clustree(seurat, prefix = 'RNA_snn_res.')
```

### Stability

Coloured by the SC3 stability metric.

```{r clustree-stability, fig.height=12, fig.width=12}
clustree(seurat, node_colour = "sc3_stability")
```

### Genes {.tabset}

Coloured by the expression of known marker genes.

```{r known-genes}
known_genes <- c(
    # fibroblasts
    "COL1A1", "COL1A2", "COL3A1", "LUM", "DCN", "C1R",
    # iCAF
    "IL6", "PDGFRA", "CXCL12", "CFD", "DPT", "AGTR1", "HAS1", "CXCL1", "CXCL2", "CCL2", "CXCL8", "LMNA", "PLA2G2A", "CLU", "EMP1",
    # myCAF
    "ACTA2", "TAGLN", "MMP11", "HOPX", "POSTN", "TPM1", "MYL9", "TPM2",
    # apCAF
    "CD74", "HLA-DRA", "HLA-DPA1", "HLA-DQA1"
)
is_present <- known_genes %in% rownames(seurat)
```

> The following genes aren't present in this dataset and will be skipped:
> `r known_genes[!is_present]`

```{r clustree-genes, results = "hide"}
src_list <- lapply(known_genes[is_present], function(gene) {
    src <- c("#### {{gene}} {.unnumbered}",
             "```{r clustree-{{gene}}, fig.height=12, fig.width=12}",
             "clustree(seurat, node_colour = '{{gene}}',",
                      "node_colour_aggr = 'mean', exprs = 'data') +",
             "scale_colour_viridis_c(option = 'plasma', begin = 0.3)",
             "```",
             "")
    knitr::knit_expand(text = src)
})
out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

