---
title: "05.cell_type"
---

```{r knitr, include = FALSE}
DOCNAME = "05.cell_type"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)
dir.create(here::here("docs", "assets", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = TRUE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# Seurat
library(Seurat)

# Plotting
library(scales)
library(ggrastr)
library(ggcorrplot)
library(pheatmap)
library(ggpubr)
library(ggplotify)
library(ggrepel)

# pvclust
library(pvclust)
library(dendextend)

# fonts
library(extrafont)
#font_import()
#font_import(paths = "/T02Data/niuyw/.local/share/fonts")
loadfonts(device = "pdf", quiet = TRUE)

# patch
library(cowplot)
theme_set(theme_cowplot(font_size = 12,
                        rel_small = 10/12,
                        rel_tiny = 8/12,
                        rel_large = 12/12,
                        font_family = "Arial"))
library(patchwork)

# color
library(ggsci)
```

```{r source, cache=FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load rds directly (since the process is too slow)

```{r load-rds, cache=FALSE}
seurat = readRDS(here::here('output/04.rm_cells/seurat.rds'))
Idents(seurat) <- 'cellType'
seurat
```

## UMAP/tSNE/PCA {.tabset}

### UMAP {.tabset}

#### Cluster

```{r umap, fig.width=7, fig.height=3, cache=FALSE}
p.umap.cellType = DimPlot(seurat, reduction = 'umap', group.by = 'cellType',
                          label = TRUE, label.size = 3) + 
  scale_color_manual(values = cell_color_maps) +
  theme_void() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1), 
        legend.position = "none",
        plot.margin = margin(0))
#p.umap.cellType
ggsave2(p.umap.cellType, filename = here::here("output", DOCNAME, "p.umap.cellType.png"),
       height = 2.5, width = 2.5, type = "cairo")

p.umap.vdj = DimPlot(seurat, reduction = 'umap', group.by = 'VDJ') + 
  scale_color_manual(values = vdj_color_maps) +
  theme_void() +
  theme(legend.key.size = unit(0.15, "inch"))
ggsave2(p.umap.vdj, filename = here::here("output", DOCNAME, "p.umap.vdj.png"),
       height = 2, width = 2.5, type = "cairo")

p.umap.cellType + p.umap.vdj
```

#### Origin2_n

```{r plot-umap-Origin2_n, cache=FALSE}
p.umap.Origin2_n = DimPlot_rast(seurat, reduction = "umap", group.by = 'Origin2_n') +
  scale_color_manual(values = origin2_color_maps) +
  theme_void() +
  theme(legend.key.size = unit(0.15, "inch"))
ggsave2(p.umap.Origin2_n, filename = here::here("output", DOCNAME, "p.umap.Origin2_n.png"),
       height = 2, width = 2.5, type = "cairo")
p.umap.Origin2_n
```

```{r plot-umap-Origin2_n-2, fig.width=12, fig.height=10}
DimPlot(seurat, reduction = "umap", split.by = 'Origin2_n', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Patient

```{r plot-umap-patient, cache=FALSE}
p.umap.Patient = DimPlot_rast(seurat, reduction = "umap", group.by = 'Patient') + 
  scale_color_manual(values = patient_color_maps) +
  theme_void() +
  theme(legend.key.size = unit(0.15, "inch"))

p.umap.Patient
```

```{r plot-umap-patient-2, fig.width=12, fig.height=8}
DimPlot(seurat, reduction = "umap", split.by = 'Patient', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Tissue

```{r plot-umap-tissue, cache=FALSE}
p.umap.Tissue = DimPlot_rast(seurat, reduction = "umap", group.by = 'Tissue') + 
  scale_color_manual(values = tissue_color_maps) +
  labs(title = "Tissue")

p.umap.Tissue
```

```{r plot-umap-tissue-2, fig.width=12, fig.height=8}
DimPlot(seurat, reduction = "umap", split.by = 'Tissue', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Metastasis_n

```{r plot-umap-Metastasis}
DimPlot(seurat, reduction = "umap", group.by = 'Metastasis_n') + 
  scale_color_manual(values = metastasis_color_maps)
```

```{r plot-umap-Metastasis-2, fig.height=4, fig.width=12}
DimPlot(seurat, reduction = "umap", split.by = 'Metastasis_n') + 
  scale_color_manual(values = cell_color_maps)
```

#### Origin2

```{r plot-umap-origin2}
DimPlot(seurat, reduction = "umap", group.by = 'Origin2') +
  scale_color_manual(values = origin2_color_maps)
```

```{r plot-umap-origin2-2, fig.width=12, fig.height=10}
DimPlot(seurat, reduction = "umap", split.by = 'Origin2', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Origin3

```{r plot-umap-origin3}
DimPlot(seurat, reduction = "umap", group.by = 'Origin3') +
  scale_color_manual(values = origin3_color_maps)
```

```{r plot-umap-origin3-2, fig.width=12, fig.height=10}
DimPlot(seurat, reduction = "umap", split.by = 'Origin3', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Drainage

```{r plot-umap-Drainage}
DimPlot(seurat, reduction = "umap", group.by = 'Drainage') + 
  scale_color_manual(values = drainage_color_maps)
```

```{r plot-umap-Drainage-2, fig.height=4, fig.width=12}
DimPlot(seurat, reduction = "umap", split.by = 'Drainage') + 
  scale_color_manual(values = cell_color_maps)
```

#### PBMC

```{r plot-umap-PBMC}
if (sum(seurat$Tissue == 'PBMC') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'PBMC'), reduction = "umap") + 
    scale_color_manual(values = cell_color_maps)
}
```

```{r plot-umap-PBMC-2, fig.width=9, fig.height=4}
if (sum(seurat$Tissue == 'PBMC') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'PBMC'), reduction = "umap", split.by = 'Origin') + 
    scale_color_manual(values = cell_color_maps)
}
```

### tSNE {.tabset}

#### Cluster

```{r plot-tsne-cluster}
DimPlot(seurat, reduction = "tsne", label = T) +
  scale_color_manual(values = cell_color_maps)
```

#### VDJ

```{r plot-tsne-vdj}
DimPlot(seurat, reduction = "tsne", label = F, group.by = 'VDJ') +
  scale_color_manual(values = vdj_color_maps)
```

#### Tissue

```{r plot-tsne-tissue}
DimPlot(seurat, reduction = "tsne", group.by = 'Tissue') + 
  scale_color_manual(values = tissue_color_maps)
```

```{r plot-tsne-tissue-2, fig.width=12, fig.height=8}
DimPlot(seurat, reduction = "tsne", split.by = 'Tissue', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Origin2_n

```{r plot-tsne-Origin2_n}
DimPlot(seurat, reduction = "tsne", group.by = 'Origin2_n') +
  scale_color_manual(values = origin2_color_maps)
```

```{r plot-tsne-Origin2_n-2, fig.width=12, fig.height=10}
DimPlot(seurat, reduction = "tsne", split.by = 'Origin2_n', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Metastasis_n

```{r plot-tsne-Metastasis}
DimPlot(seurat, reduction = "tsne", group.by = 'Metastasis_n') + 
  scale_color_manual(values = metastasis_color_maps)
```

```{r plot-tsne-Metastasis-2, fig.height=4, fig.width=12}
DimPlot(seurat, reduction = "tsne", split.by = 'Metastasis_n') + 
  scale_color_manual(values = cell_color_maps)
```

#### Origin2

```{r plot-tsne-origin2}
DimPlot(seurat, reduction = "tsne", group.by = 'Origin2') +
  scale_color_manual(values = origin2_color_maps)
```

```{r plot-tsne-origin2-2, fig.width=12, fig.height=10}
DimPlot(seurat, reduction = "tsne", split.by = 'Origin2', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Origin3

```{r plot-tsne-origin3}
DimPlot(seurat, reduction = "tsne", group.by = 'Origin3') +
  scale_color_manual(values = origin3_color_maps)
```

```{r plot-tsne-origin3-2, fig.width=12, fig.height=10}
DimPlot(seurat, reduction = "tsne", split.by = 'Origin3', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Drainage

```{r plot-tsne-Drainage}
DimPlot(seurat, reduction = "tsne", group.by = 'Drainage') + 
  scale_color_manual(values = drainage_color_maps)
```

```{r plot-tsne-Drainage-2, fig.height=4, fig.width=12}
DimPlot(seurat, reduction = "tsne", split.by = 'Drainage') + 
  scale_color_manual(values = cell_color_maps)
```

#### Patient

```{r plot-tsne-patient}
DimPlot(seurat, reduction = "tsne", group.by = 'Patient') + 
  scale_color_manual(values = patient_color_maps)
```

```{r plot-tsne-patient-2, fig.width=12, fig.height=8}
DimPlot(seurat, reduction = "tsne", split.by = 'Patient', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### PBMC

```{r plot-tsne-PBMC}
if (sum(seurat$Tissue == 'PBMC') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'PBMC'), reduction = "tsne") + 
    scale_color_manual(values = cell_color_maps)
}
```

```{r plot-tsne-PBMC-2, fig.width=9, fig.height=4}
if (sum(seurat$Tissue == 'PBMC') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'PBMC'), reduction = "tsne", split.by = 'Origin') + 
    scale_color_manual(values = cell_color_maps)
}
```

#### LN

```{r plot-tsne-LN}
if (sum(seurat$Tissue == 'LN') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'LN'), reduction = "tsne") + 
    scale_color_manual(values = cell_color_maps)
}
```

```{r plot-tsne-LN-2, fig.width=12, fig.height=12}
if (sum(seurat$Tissue == 'LN') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'LN'), reduction = "tsne", split.by = 'Origin', ncol = 3) + 
    scale_color_manual(values = cell_color_maps)
}
```

```{r plot-tsne-LN-3, fig.width=10, fig.height=5}
if (sum(seurat$Tissue == 'LN') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'LN'), reduction = "tsne", split.by = 'Metastasis_n') + 
    scale_color_manual(values = cell_color_maps)
}
```

#### Tumor

```{r plot-tsne-Tumor}
if (sum(seurat$Tissue == 'Tumor') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'Tumor'), reduction = "tsne", pt.size = 0.1) + 
    scale_color_manual(values = cell_color_maps)
}
```

```{r plot-tsne-Tumor-2, fig.width=12, fig.height=4}
if (sum(seurat$Tissue == 'Tumor') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'Tumor'), reduction = "tsne", split.by = 'Origin', pt.size = 0.1) + 
    scale_color_manual(values = cell_color_maps)
}
```

#### Origin

```{r plot-tsne-origin}
DimPlot(seurat, reduction = "tsne", group.by = 'Origin') +
  scale_color_manual(values = origin_color_maps)
```

```{r plot-tsne-origin-2, fig.width=12, fig.height=16}
DimPlot(seurat, reduction = "tsne", split.by = 'Origin', ncol = 4) +
  scale_color_manual(values = cell_color_maps)
```

### PCA {.tabset}

#### Cluster

```{r plot-pca-cluster}
DimPlot(seurat, reduction = "pca", label = T) +
  scale_color_manual(values = cell_color_maps)
```

#### VDJ

```{r plot-pca-vdj}
DimPlot(seurat, reduction = "pca", label = F, group.by = 'VDJ') +
  scale_color_manual(values = vdj_color_maps)
```

#### Tissue

```{r plot-pca-tissue}
DimPlot(seurat, reduction = "pca", group.by = 'Tissue') + 
  scale_color_manual(values = tissue_color_maps)
```

```{r plot-pca-tissue-2, fig.width=12, fig.height=8}
DimPlot(seurat, reduction = "pca", split.by = 'Tissue', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Origin2_n

```{r plot-pca-origin2_n}
DimPlot(seurat, reduction = "pca", group.by = 'Origin2_n') +
  scale_color_manual(values = origin2_color_maps)
```

```{r plot-pca-origin2_n-2, fig.width=12, fig.height=10}
DimPlot(seurat, reduction = "pca", split.by = 'Origin2_n', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Metastasis_n

```{r plot-pca-Metastasis}
DimPlot(seurat, reduction = "pca", group.by = 'Metastasis_n') + 
  scale_color_manual(values = metastasis_color_maps)
```

```{r plot-pca-Metastasis-2, fig.height=4, fig.width=12}
DimPlot(seurat, reduction = "pca", split.by = 'Metastasis_n') + 
  scale_color_manual(values = cell_color_maps)
```

#### Origin2

```{r plot-pca-origin2}
DimPlot(seurat, reduction = "pca", group.by = 'Origin2') +
  scale_color_manual(values = origin2_color_maps)
```

```{r plot-pca-origin2-2, fig.width=12, fig.height=10}
DimPlot(seurat, reduction = "pca", split.by = 'Origin2', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Origin3

```{r plot-pca-origin3}
DimPlot(seurat, reduction = "pca", group.by = 'Origin3') +
  scale_color_manual(values = origin3_color_maps)
```

```{r plot-pca-origin3-2, fig.width=12, fig.height=10}
DimPlot(seurat, reduction = "pca", split.by = 'Origin3', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### Drainage

```{r plot-pca-Drainage}
DimPlot(seurat, reduction = "pca", group.by = 'Drainage') + 
  scale_color_manual(values = drainage_color_maps)
```

```{r plot-pca-Drainage-2, fig.height=4, fig.width=12}
DimPlot(seurat, reduction = "pca", split.by = 'Drainage') + 
  scale_color_manual(values = cell_color_maps)
```

#### Patient

```{r plot-pca-patient}
DimPlot(seurat, reduction = "pca", group.by = 'Patient') + 
  scale_color_manual(values = patient_color_maps)
```

```{r plot-pca-patient-2, fig.width=12, fig.height=8}
DimPlot(seurat, reduction = "pca", split.by = 'Patient', ncol = 3) +
  scale_color_manual(values = cell_color_maps)
```

#### PBMC

```{r plot-pca-PBMC}
if (sum(seurat$Tissue == 'PBMC') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'PBMC'), reduction = "pca") + 
    scale_color_manual(values = cell_color_maps)
}
```

```{r plot-pca-PBMC-2, fig.width=9, fig.height=4}
if (sum(seurat$Tissue == 'PBMC') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'PBMC'), reduction = "pca", split.by = 'Origin') + 
    scale_color_manual(values = cell_color_maps)
}
```

#### LN

```{r plot-pca-LN}
if (sum(seurat$Tissue == 'LN') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'LN'), reduction = "pca") + 
    scale_color_manual(values = cell_color_maps)
}
```

```{r plot-pca-LN-2, fig.width=12, fig.height=12}
if (sum(seurat$Tissue == 'LN') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'LN'), reduction = "pca", split.by = 'Origin', ncol = 3) + 
    scale_color_manual(values = cell_color_maps)
}
```

```{r plot-pca-LN-3, fig.width=10, fig.height=5}
if (sum(seurat$Tissue == 'LN') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'LN'), reduction = "pca", split.by = 'Metastasis_n') + 
    scale_color_manual(values = cell_color_maps)
}
```

#### Tumor

```{r plot-pca-Tumor}
if (sum(seurat$Tissue == 'Tumor') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'Tumor'), reduction = "pca", pt.size = 0.1) + 
    scale_color_manual(values = cell_color_maps)
}
```

```{r plot-pca-Tumor-2, fig.width=12, fig.height=4}
if (sum(seurat$Tissue == 'Tumor') > 2) {
  DimPlot(subset(seurat, subset = Tissue == 'Tumor'), reduction = "pca", split.by = 'Origin', pt.size = 0.1) + 
    scale_color_manual(values = cell_color_maps)
}
```

#### Origin

```{r plot-pca-origin}
DimPlot(seurat, reduction = "pca", group.by = 'Origin') +
  scale_color_manual(values = origin_color_maps)
```

```{r plot-pca-origin-2, fig.width=12, fig.height=16}
DimPlot(seurat, reduction = "pca", split.by = 'Origin', ncol = 4) +
  scale_color_manual(values = cell_color_maps)
```

## Markers {.tabset}

Find markers again

```{r find-cell-markers, cache=FALSE, eval=FALSE}
cell.markers = FindAllMarkers(seurat, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.5, verbose = T)
write.table(cell.markers, file = here::here('output', DOCNAME, 'cell.markers.txt'), sep = '\t', quote = F, row.names = F)
```

```{r read-cell-markers, cache=FALSE, eval=FALSE}
cell.markers = read.table(here::here('output', DOCNAME, 'cell.markers.txt'), header = T, sep = "\t", stringsAsFactors = F)
DT::datatable(cell.markers %>% 
                group_by(cluster) %>% 
                top_n(n = 2, wt = avg_logFC))
```

Plot markers (main)

```{r plot-cell-markers-main, fig.width=7.5, fig.height=1.2, cache=FALSE}
p.marker.main = customFeaturePlot(seurat, features = c(
  # T
  "CD3D",
  # B
  "CD79A",
  # mye
  "LYZ",
  # mast
  "TPSAB1",
  # epi
  "KRT6A",
  # fib
  "COL1A2",
  # endo
  "CDH5",
  # platelets
  "PPBP"
  # 
)) %>%
  wrap_plots(nrow = 1) & theme(plot.margin = margin(0))
p.marker.main
```

Plot markers (supp.)

```{r plot-cell-markers-supp, fig.width=7.5, fig.height=5, cache=FALSE}
p.marker.supp = customFeaturePlot(seurat, features = c(
  # T
  'CD3D', 'CD3G', 'CD3E', 'CD2',
  # B
  'CD19', 'MS4A1', 'CD79A', 'IGHM',
  # mye
  "LYZ", "FCER1G", "TYROBP", "AIF1",
  # mast
  "TPSAB1", "TPSB2", "MS4A2", 'KIT',
  # epi
  "CNFN", "KRT6A", "CRNN", "KRT13",
  # fib
  'DCN', 'COL1A1', 'COL1A2', 'LUM',
  # endo
  "PECAM1", "CLDN5", "CDH5", "VWF",
  # platelets
  "PPBP", "PF4", "ITGA2B", "ITGB3"
  # 
)) %>%
  wrap_plots(nrow = 4, byrow = FALSE) & theme(plot.margin = margin(0))
p.marker.supp
```

### T cells

```{r umap-tcells-markers, fig.width=8, fig.height=8}
FeaturePlot1 <- function(seurat_obj, reduction, features) {
  p <- FeaturePlot(seurat_obj, reduction = reduction, features = features, combine = FALSE)
  p <- lapply(X = p, FUN = function(x) (x + NoLegend()))
  patchwork::wrap_plots(p)
}

FeaturePlot1(seurat, 'umap', c('CD3D', 'CD3G', 'CD3E', 'CD2'))
```

### B cells

```{r umap-bcells-markers, fig.width=8, fig.height=8}
FeaturePlot1(seurat, 'umap', c('CD19', 'MS4A1', 'CD79A', 'IGHM'))
```

### Myeloid

```{r umap-myeloid-markers, fig.width=8, fig.height=8}
FeaturePlot1(seurat, 'umap', c("LYZ", "FCER1G", "TYROBP", "AIF1"))
```

### Mast cells

```{r umap-Mast-markers, fig.width=8, fig.height=8}
FeaturePlot1(seurat, 'umap', c("TPSAB1", "TPSB2", "MS4A2", 'KIT'))
```

### Epithelia

```{r umap-epithelia-markers, fig.width=8, fig.height=8}
epithelia_markers = c("KRT13", "SPRR3", "CNFN", "KRT6A", "CRNN", "CSTA", "SOX2", "TP63", "EPCAM")
FeaturePlot1(seurat, 'umap', epithelia_markers)
```

### Fibroblasts

```{r umap-fibroblasts-markers, fig.width=8, fig.height=8}
FeaturePlot1(seurat, 'umap', c('DCN', 'COL1A1', 'COL1A2', 'LUM'))
```

### Endothelia

```{r umap-endothelia-markers, fig.width=10, fig.height=6}
FeaturePlot1(seurat, 'umap', markers.endothelia)
```

### Pericyte

```{r umap-Pericyte-markers, fig.width=10, fig.height=6}
FeaturePlot1(seurat, 'umap', markers.pericyte)
```

### Mesothelia

```{r umap-Mesothelia-markers, fig.width=6, fig.height=6}
FeaturePlot1(seurat, 'umap', markers.mesothelia)
```

### Platelets

```{r umap-Platelets-markers, fig.width=8, fig.height=8}
FeaturePlot1(seurat, 'umap', markers.platelets)
```

## Cluster quality

```{r cluster-qc, fig.width=7, fig.height=5, cache=FALSE}
p.quality.nCount_RNA = seurat@meta.data %>%
  group_by(cellType, nCount_RNA) %>%
  summarise() %>%
  ggplot(aes(cellType, nCount_RNA)) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot(aes(fill = 'blue'), outlier.alpha = 0.1, color = '#00008f') +
  scale_fill_manual(values = '#8f8fb1') +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  theme(axis.title.y = element_blank(),
        legend.position = "none") +
  coord_flip()

p.quality.nFeature_RNA = seurat@meta.data %>%
  group_by(cellType, nFeature_RNA) %>%
  summarise() %>%
  ggplot(aes(cellType, nFeature_RNA)) +
  stat_boxplot(geom = "errorbar", width = 0.4) +
  geom_boxplot(aes(fill = 'blue'), outlier.alpha = 0.1, color = '#00008f') +
  scale_fill_manual(values = '#8f8fb1') +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  theme(axis.title.y = element_blank(),
        legend.position = "none") +
  coord_flip()

features.to.plot = c('percent.mt', 'percent.ribo',
                     'percent.dissociation', 'percent.heat')
p_list = map(features.to.plot, feature_box, seurat_obj = seurat, groupby = 'cellType')


p_list = list(p.quality.nCount_RNA, p.quality.nFeature_RNA,
              p_list[[1]], p_list[[2]], p_list[[3]], p_list[[4]])

p.quality = p_list %>%
  wrap_plots(ncol = 3)
p.quality
```

## Percent of VDJ

```{r percent-vdj, fig.height=4, fig.width=6}
percent_vdj(seurat, groupby = 'cellType')
```

## cellType compositions

```{r cell-proportion, fig.height=2.5, fig.width=7.5, cache=FALSE}
p.composition = cluster_stack(seurat, group_by = "cellType")
p.composition
```

Tissue/Origin3/Origin

```{r cell-proportion-2, fig.height=3.5, fig.width=10, cache=FALSE}
cluster_stack2(seurat, group_by = "cellType")
```

## Sample composition

Source

```{r sample-compositions, fig.width=10, fig.height=3}
PlotSampleFraction(seurat_obj = seurat, 'cellType', cell_color_maps)
```

Origin2_n by cellType

```{r sample-compositions-Origin2_n, fig.width=7, fig.height=3}
p.cell_by_origin2 = seurat@meta.data %>%
  filter(cellType != "Platelets") %>%
  group_by(Origin2_n, cellType) %>%
  summarise(n = n()) %>%
  mutate(prop = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = prop, fill = Origin2_n)) + 
  #geom_col(color = 'white', width = .8) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = origin2_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  guides(fill = guide_legend(ncol = 1)) +
  labs(x = NULL, y = "Fraction (%)", fill = NULL) +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 60, hjust = 1))
p.cell_by_origin2
```

Origin2_n by cellType by Patient

```{r sample-compositions-Origin2_n-by-Patient, fig.width=7.5, fig.height=4}
seurat@meta.data %>%
  filter(cellType != "Platelets") %>%
  group_by(Patient, Origin2_n, cellType) %>%
  summarise(n = n()) %>%
  mutate(prop = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = prop, fill = Origin2_n)) + 
  #geom_col(color = 'white', width = .8) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = origin2_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(.~Patient) +
  guides(fill = guide_legend(ncol = 1)) +
  labs(x = NULL, y = "Fraction (%)", fill = NULL) +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 60, hjust = 1))
```

Another way to show

```{r sample-compositions-Origin2_n-by-Patient-2, fig.width=7.5, fig.height=3}
p.cell_by_origin2 = seurat@meta.data %>%
  filter(cellType != "Platelets") %>%
  group_by(Patient, Origin2_n, cellType) %>%
  summarise(n = n()) %>%
  mutate(prop = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin2_n, y = prop, fill = cellType)) + 
  geom_col(color = 'white', width = .8) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_grid(.~Patient) +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = NULL, y = "Fraction", fill = NULL) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 60, hjust = 1))
p.cell_by_origin2
```


## Enrichment by {.tabset}

### Origin2_n

```{r Origin2_n-enrichment, fig.width=4, fig.height=3, cache=FALSE}
cluster_cellNum = seurat@meta.data %>% 
  group_by(cellType) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum = seurat@misc$sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 = seurat@meta.data %>%
  group_by(cellType, Origin2_n) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp = setdiff(levels(cluster_byOrigin2$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x = tibble(cellType = levels(cluster_cellNum$cellType), Origin2_n = t, o = 0)
    cluster_byOrigin2 = bind_rows(cluster_byOrigin2, x)
  }
}

r_oe = cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = 'Origin2_n') %>%
  left_join(cluster_cellNum, by = 'cellType') %>%
  mutate(e = (origin.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(cellType, Origin2_n, r) %>% 
  spread(Origin2_n, r, fill = 0) %>% 
  column_to_rownames("cellType")

# use ggplot for better control
p.enrich.origin2 = plot_origin2_enrich(r_oe)
p.enrich.origin2
```

### Tissue

```{r tissue-enrichment, fig.width=4, fig.height=4, cache=FALSE}
cluster_cellNum = seurat@meta.data %>% 
  group_by(cellType) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

tissue_CellNum = seurat@misc$sam_info %>%
  group_by(Tissue) %>%
  summarise(tissue.cell_num = sum(cell.filtered)) %>%
  mutate(tissue.cell_num = as.numeric(tissue.cell_num)) %>%
  mutate(total = sum(tissue.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byTissue = seurat@meta.data %>%
  group_by(cellType, Tissue) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add tissues not existing
tmp = setdiff(levels(cluster_byTissue$Tissue), cluster_byTissue$Tissue)
if (length(tmp) > 0) {
  for (t in tmp) {
    x = tibble(cellType = levels(cluster_byTissue$cellType), Tissue = t, o = 0)
    cluster_byTissue = bind_rows(cluster_byTissue, x)
  }
}

r_oe = cluster_byTissue %>%
  left_join(tissue_CellNum, by = 'Tissue') %>%
  left_join(cluster_cellNum, by = 'cellType') %>%
  mutate(e = (tissue.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(cellType, Tissue, r) %>% 
  spread(Tissue, r, fill = 0) %>% 
  column_to_rownames("cellType")

# use ggplot for better control
gg = plot_tissue_enrich(r_oe)
gg
```

### Origin3

```{r origin3-enrichment, fig.width=5, fig.height=4, cache=FALSE}
cluster_cellNum = seurat@meta.data %>% 
  group_by(cellType) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin_CellNum = seurat@misc$sam_info %>%
  group_by(Origin3) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin3 = seurat@meta.data %>% 
  group_by(cellType, Origin3) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp = setdiff(levels(cluster_byOrigin3$Origin3), cluster_byOrigin3$Origin3)
if (length(tmp) > 0) {
  for (t in tmp) {
    x = tibble(cellType = levels(cluster_cellNum$cellType), Origin3 = t, o = 0)
    cluster_byOrigin3 = bind_rows(cluster_byOrigin3, x)
  }
}

r_oe = cluster_byOrigin3 %>%
  left_join(origin_CellNum, by = 'Origin3') %>%
  left_join(cluster_cellNum, by = 'cellType') %>%
  mutate(e = (origin.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(cellType, Origin3, r) %>% 
  spread(Origin3, r, fill = 0) %>% 
  column_to_rownames("cellType")

# use ggplot for better control
gg = plot_origin3_enrich(r_oe)
gg
```

### Metastasis P/N patient: Tumor

```{r Metastasis-p-n-patient-tumor-enrichment, fig.width=2.5, fig.height=4, cache=FALSE}
patient_info = tibble(Patient = c("S0619", "S0730", "S0819", "S0920", "S1125", "S1204"),
                      Metastatic = c('P', 'N', 'N', 'N', 'P', 'P'))

cluster_cellNum = seurat@meta.data %>% 
  group_by(cellType) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

patient_CellNum = seurat@misc$sam_info %>%
  filter(Tissue == 'Tumor') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(Metastatic) %>%
  summarise(patient.cell_num = sum(cell.filtered)) %>%
  mutate(patient.cell_num = as.numeric(patient.cell_num)) %>%
  mutate(total = sum(patient.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byPatient = seurat@meta.data %>% 
  filter(Tissue == 'Tumor') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(cellType, Metastatic) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp = setdiff(c("N", "P"), cluster_byPatient$Metastatic)
if (length(tmp) > 0) {
  for (t in tmp) {
    x = tibble(cellType = levels(cluster_cellNum$cellType), Metastatic = t, o = 0)
    cluster_byPatient = bind_rows(cluster_byPatient, x)
  }
}

r_oe = cluster_byPatient %>%
  left_join(patient_CellNum, by = 'Metastatic') %>%
  left_join(cluster_cellNum, by = 'cellType') %>%
  mutate(e = (patient.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(cellType, Metastatic, r) %>% 
  spread(Metastatic, r, fill = 0) %>% 
  column_to_rownames("cellType")

# use ggplot for better control
gg = plot_metastasis_enrich(r_oe)
gg
```

### Metastasis P/N patient: Adjacent

```{r Metastasis-p-n-patient-Adjacent-enrichment, fig.width=2.5, fig.height=4, cache=FALSE}
patient_info = tibble(Patient = c("S0619", "S0730", "S0819", "S0920", "S1125", "S1204"),
                      Metastatic = c('P', 'N', 'N', 'N', 'P', 'P'))

cluster_cellNum = seurat@meta.data %>% 
  group_by(cellType) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

patient_CellNum = seurat@misc$sam_info %>%
  filter(Tissue == 'Adjacent') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(Metastatic) %>%
  summarise(patient.cell_num = sum(cell.filtered)) %>%
  mutate(patient.cell_num = as.numeric(patient.cell_num)) %>%
  mutate(total = sum(patient.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byPatient = seurat@meta.data %>% 
  filter(Tissue == 'Adjacent') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(cellType, Metastatic) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp = setdiff(c("N", "P"), cluster_byPatient$Metastatic)
if (length(tmp) > 0) {
  for (t in tmp) {
    x = tibble(cellType = levels(cluster_cellNum$cellType), Metastatic = t, o = 0)
    cluster_byPatient = bind_rows(cluster_byPatient, x)
  }
}

r_oe = cluster_byPatient %>%
  left_join(patient_CellNum, by = 'Metastatic') %>%
  left_join(cluster_cellNum, by = 'cellType') %>%
  mutate(e = (patient.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(cellType, Metastatic, r) %>% 
  spread(Metastatic, r, fill = 0) %>% 
  column_to_rownames("cellType")

# use ggplot for better control
gg = plot_metastasis_enrich(r_oe)
gg
```

### Metastasis P/N patient: Normal

```{r Metastasis-p-n-patient-Normal-enrichment, fig.width=2.5, fig.height=4, cache=FALSE}
patient_info = tibble(Patient = c("S0619", "S0730", "S0819", "S0920", "S1125", "S1204"),
                      Metastatic = c('P', 'N', 'N', 'N', 'P', 'P'))

cluster_cellNum = seurat@meta.data %>% 
  group_by(cellType) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

patient_CellNum = seurat@misc$sam_info %>%
  filter(Tissue == 'Normal') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(Metastatic) %>%
  summarise(patient.cell_num = sum(cell.filtered)) %>%
  mutate(patient.cell_num = as.numeric(patient.cell_num)) %>%
  mutate(total = sum(patient.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byPatient = seurat@meta.data %>% 
  filter(Tissue == 'Normal') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(cellType, Metastatic) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp = setdiff(c("N", "P"), cluster_byPatient$Metastatic)
if (length(tmp) > 0) {
  for (t in tmp) {
    x = tibble(cellType = levels(cluster_cellNum$cellType), Metastatic = t, o = 0)
    cluster_byPatient = bind_rows(cluster_byPatient, x)
  }
}

r_oe = cluster_byPatient %>%
  left_join(patient_CellNum, by = 'Metastatic') %>%
  left_join(cluster_cellNum, by = 'cellType') %>%
  mutate(e = (patient.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(cellType, Metastatic, r) %>% 
  spread(Metastatic, r, fill = 0) %>% 
  column_to_rownames("cellType")

# use ggplot for better control
gg = plot_metastasis_enrich(r_oe)
gg
```

### Metastasis P/N patient: PBMC

```{r Metastasis-p-n-patient-PBMC-enrichment, fig.width=2.5, fig.height=4, cache=FALSE}
patient_info = tibble(Patient = c("S0619", "S0730", "S0819", "S0920", "S1125", "S1204"),
                      Metastatic = c('P', 'N', 'N', 'N', 'P', 'P'))

cluster_cellNum = seurat@meta.data %>% 
  group_by(cellType) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

patient_CellNum = seurat@misc$sam_info %>%
  filter(Tissue == 'PBMC') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(Metastatic) %>%
  summarise(patient.cell_num = sum(cell.filtered)) %>%
  mutate(patient.cell_num = as.numeric(patient.cell_num)) %>%
  mutate(total = sum(patient.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byPatient = seurat@meta.data %>% 
  filter(Tissue == 'PBMC') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(cellType, Metastatic) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp = setdiff(c("N", "P"), cluster_byPatient$Metastatic)
if (length(tmp) > 0) {
  for (t in tmp) {
    x = tibble(cellType = levels(cluster_cellNum$cellType), Metastatic = t, o = 0)
    cluster_byPatient = bind_rows(cluster_byPatient, x)
  }
}

r_oe = cluster_byPatient %>%
  left_join(patient_CellNum, by = 'Metastatic') %>%
  left_join(cluster_cellNum, by = 'cellType') %>%
  mutate(e = (patient.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(cellType, Metastatic, r) %>% 
  spread(Metastatic, r, fill = 0) %>% 
  column_to_rownames("cellType")

# use ggplot for better control
gg = plot_metastasis_enrich(r_oe)
gg
```

### Metastasis P/N patient: LN

```{r Metastasis-p-n-patient-LN-enrichment, fig.width=2.5, fig.height=4, cache=FALSE}
patient_info = tibble(Patient = c("S0619", "S0730", "S0819", "S0920", "S1125", "S1204"),
                      Metastatic = c('P', 'N', 'N', 'N', 'P', 'P'))

cluster_cellNum = seurat@meta.data %>% 
  group_by(cellType) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

patient_CellNum = seurat@misc$sam_info %>%
  filter(Tissue == 'LN') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(Metastatic) %>%
  summarise(patient.cell_num = sum(cell.filtered)) %>%
  mutate(patient.cell_num = as.numeric(patient.cell_num)) %>%
  mutate(total = sum(patient.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byPatient = seurat@meta.data %>% 
  filter(Tissue == 'LN') %>%
  left_join(patient_info, by = 'Patient') %>%
  group_by(cellType, Metastatic) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

r_oe = cluster_byPatient %>%
  left_join(patient_CellNum, by = 'Metastatic') %>%
  left_join(cluster_cellNum, by = 'cellType') %>%
  mutate(e = (patient.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(cellType, Metastatic, r) %>% 
  spread(Metastatic, r, fill = 0) %>% 
  column_to_rownames("cellType")

# use ggplot for better control
gg = plot_metastasis_enrich(r_oe)
gg
```

### Origin

```{r origin-enrichment, fig.width=8, fig.height=4, cache=FALSE}
cluster_cellNum = seurat@meta.data %>% 
  group_by(cellType) %>% 
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin_CellNum = seurat@misc$sam_info %>%
  group_by(Origin) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin = seurat@meta.data %>% 
  group_by(cellType, Origin) %>% 
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp = setdiff(levels(cluster_byOrigin$Origin), cluster_byOrigin$Origin)
if (length(tmp) > 0) {
  for (t in tmp) {
    x = tibble(cellType = levels(cluster_cellNum$cellType), Origin = t, o = 0)
    cluster_byOrigin = bind_rows(cluster_byOrigin, x)
  }
}

r_oe = cluster_byOrigin %>%
  left_join(origin_CellNum, by = 'Origin') %>%
  left_join(cluster_cellNum, by = 'cellType') %>%
  mutate(e = (origin.cell_num * cluster.cell_num)/total) %>%
  mutate(r = o/e) %>%
  dplyr::select(cellType, Origin, r) %>% 
  spread(Origin, r, fill = 0) %>% 
  column_to_rownames("cellType")

# use ggplot for better control
gg = plot_origin_enrich(r_oe)
gg
```

## Cell number {.tabset}

```{r cell-num, fig.width=10, fig.height=15}
p_cluster = seurat@meta.data %>%
  group_by(cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes("", Percent, fill = cellType)) +
  geom_bar(width = 1, size = 1, color = "white", stat = "identity") +
  coord_polar("y") +
  geom_text_repel(aes(label = paste0(round(Percent, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, title = 'cellType') +
  scale_fill_manual(values = cell_color_maps)  +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))

p_tissue = seurat@meta.data %>%
  group_by(Tissue) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes("", Percent, fill = Tissue)) +
  geom_bar(width = 1, size = 1, color = "white", stat = "identity") +
  coord_polar("y") +
  geom_text(aes(label = paste0(round(Percent, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, title = 'Tissue') +
  scale_fill_manual(values = tissue_color_maps)  +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))

p_origin = seurat@meta.data %>%
  group_by(Origin) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes("", Percent, fill = Origin)) +
  geom_bar(width = 1, size = 1, color = "white", stat = "identity") +
  coord_polar("y") +
  geom_text_repel(aes(label = paste0(round(Percent, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, title = 'Origin') +
  scale_fill_manual(values = origin_color_maps)  +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))

p_origin2 = seurat@meta.data %>%
  group_by(Origin2_n) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes("", Percent, fill = Origin2_n)) +
  geom_bar(width = 1, size = 1, color = "white", stat = "identity") +
  coord_polar("y") +
  geom_text_repel(aes(label = paste0(round(Percent, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, title = 'Origin2_n') +
  scale_fill_manual(values = origin2_color_maps)  +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))

p_origin3 = seurat@meta.data %>%
  group_by(Origin3) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes("", Percent, fill = Origin3)) +
  geom_bar(width = 1, size = 1, color = "white", stat = "identity") +
  coord_polar("y") +
  geom_text_repel(aes(label = paste0(round(Percent, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, title = 'Origin3') +
  scale_fill_manual(values = origin3_color_maps)  +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))

p_metastasis = seurat@meta.data %>%
  group_by(Metastasis_n) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes("", Percent, fill = Metastasis_n)) +
  geom_bar(width = 1, size = 1, color = "white", stat = "identity") +
  coord_polar("y") +
  geom_text_repel(aes(label = paste0(round(Percent, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, title = 'Metastasis_n') +
  scale_fill_manual(values = metastasis_color_maps)  +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))

p_drainage = seurat@meta.data %>%
  group_by(Drainage) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes("", Percent, fill = Drainage)) +
  geom_bar(width = 1, size = 1, color = "white", stat = "identity") +
  coord_polar("y") +
  geom_text_repel(aes(label = paste0(round(Percent, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, title = 'Drainage') +
  scale_fill_manual(values = drainage_color_maps)  +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))

p_patient = seurat@meta.data %>%
  group_by(Patient) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes("", Percent, fill = Patient)) +
  geom_bar(width = 1, size = 1, color = "white", stat = "identity") +
  coord_polar("y") +
  geom_text_repel(aes(label = paste0(round(Percent, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, title = 'Patient') +
  scale_fill_manual(values = patient_color_maps)  +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))

p_cluster + p_tissue +
  p_metastasis + p_origin2 +
  p_drainage + p_origin3 +
  p_patient + p_origin +
  plot_layout(ncol = 2)
```

### cellType

```{r cell-num-cellType, cache=FALSE}
DT::datatable(
  seurat@meta.data %>% 
    group_by(cellType) %>% 
    summarise(n = n()) %>% 
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Tissue

```{r cell-num-Tissue, cache=FALSE}
DT::datatable(
  seurat@meta.data %>%
    group_by(Tissue) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Origin2_n

```{r cell-num-Origin2_n, cache=FALSE}
DT::datatable(
  seurat@meta.data %>%
    group_by(Origin2_n) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Origin3

```{r cell-num-Origin3, cache=FALSE}
DT::datatable(
  seurat@meta.data %>%
    group_by(Origin3) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Patient

```{r cell-num-Patient, cache=FALSE}
DT::datatable(
  seurat@meta.data %>%
    group_by(Patient) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Origin

```{r cell-num-Origin, cache=FALSE}
DT::datatable(
  seurat@meta.data %>%
    group_by(Origin) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Source

```{r cell-num-Source, cache=FALSE}
DT::datatable(
  seurat@meta.data %>%
    group_by(Source) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

## ... by cluster {.tabset}

```{r Tissue-by-cluster, fig.width=8, fig.height=12}
p_tissue = seurat@meta.data %>% 
  group_by(Tissue, cellType) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Tissue, y = Percent, fill = cellType)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p_patient = seurat@meta.data %>% 
  group_by(Patient, cellType) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Patient, y = Percent, fill = cellType)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p_origin = seurat@meta.data %>% 
  group_by(Origin, cellType) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin, y = Percent, fill = cellType)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p_origin2 = seurat@meta.data %>% 
  group_by(Origin2_n, cellType) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin2_n, y = Percent, fill = cellType)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p_origin3 = seurat@meta.data %>% 
  group_by(Origin3, cellType) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin3, y = Percent, fill = cellType)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

#(p1 + p2 + p3 + p4)/p5 +
#  plot_layout(guides = 'collect')
p1 = p_origin2 + p_patient
p2 = p_tissue | p_origin3
p3 = p_origin
p1/p2/p3 +
  plot_layout(guides = 'collect')
```

### Tissue

```{r cell-num-tissue-by-cellType}
DT::datatable(
  seurat@meta.data %>%
    group_by(Tissue, cellType) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

```{r frac-Tissue, fig.width=8, fig.height=10}
Tissues = c("PBMC", 'LN', 'Normal', "Adjacent", "Tumor")
sapply(Tissues, PlotFracTissue, seurat_obj = seurat, groupby = 'cellType', color_maps = cell_color_maps, simplify = F) %>%
  wrap_plots() +
  plot_layout(ncol = 2, guides = 'collect')
```

### Origin2_n

```{r cell-num-Origin2_n-by-cluster}
DT::datatable(
  seurat@meta.data %>%
    group_by(Origin2_n, cellType) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

```{r frac-Origin2_n, fig.width=12, fig.height=12}
Origin2s_n = intersect(levels(seurat$Origin2_n), unique(seurat$Origin2_n))
sapply(Origin2s_n, PlotFrac, seurat_obj = seurat, plot_what = "Origin2_n", groupby = "cellType", color_maps = cell_color_maps, simplify = F) %>%
  wrap_plots() +
  plot_layout(ncol = 3, guides = 'collect')
```

### Origin3

```{r cell-num-Origin3-by-cluster}
DT::datatable(
  seurat@meta.data %>%
    group_by(Origin3, cellType) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

```{r frac-Origin3, fig.width=12, fig.height=12}
Origins3 = intersect(levels(seurat$Origin3), unique(seurat$Origin3))
sapply(Origins3, PlotFrac, seurat_obj = seurat, plot_what = "Origin3", groupby = "cellType", color_maps = cell_color_maps, simplify = F) %>%
  wrap_plots() +
  plot_layout(ncol = 3, guides = 'collect')
```

### Origin3 of LN_N

```{r cell-num-Origin3-by-cluster-LN_N}
DT::datatable(
  seurat@meta.data %>%
    filter(Metastasis_n == "N") %>%
    group_by(Origin3, cellType) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

```{r frac-Origin3-LN_N, fig.width=10, fig.height=4, cache=FALSE}
srat_subset = subset(seurat, subset = Metastasis_n == "N")
Origins3 = intersect(levels(srat_subset$Origin3), unique(srat_subset$Origin3))
sapply(Origins3, PlotFrac, seurat_obj = srat_subset, plot_what = "Origin3", groupby = "cellType", color_maps = cell_color_maps, simplify = F) %>%
  wrap_plots() +
  plot_layout(ncol = 3, guides = 'collect')
```

### Origin3 of LN_P

```{r cell-num-Origin3-by-cluster-LN_P}
DT::datatable(
  seurat@meta.data %>%
    filter(Metastasis_n == "P") %>%
    group_by(Origin3, cellType) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

```{r frac-Origin3-LN_P, fig.width=10, fig.height=4, cache=FALSE}
srat_subset = subset(seurat, subset = Metastasis_n == "P")
Origins3 = intersect(levels(srat_subset$Origin3), unique(srat_subset$Origin3))
sapply(Origins3, PlotFrac, seurat_obj = srat_subset, plot_what = "Origin3", groupby = "cellType", color_maps = cell_color_maps, simplify = F) %>%
  wrap_plots() +
  plot_layout(ncol = 3, guides = 'collect')
```

### Origin

```{r cell-num-Origin-by-cluster}
DT::datatable(
  seurat@meta.data %>%
    group_by(Origin, cellType) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

```{r frac-Origin, fig.width=12, fig.height=20}
Origins = intersect(levels(seurat$Origin), unique(seurat$Origin))
sapply(Origins, PlotFrac, seurat_obj = seurat, plot_what = "Origin", groupby = "cellType", color_maps = cell_color_maps, simplify = F) %>%
  wrap_plots() +
  plot_layout(ncol = 3, guides = 'collect')
```

### Patient

```{r cell-num-Patient-by-cluster}
DT::datatable(
  seurat@meta.data %>%
    group_by(Patient, cellType) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

```{r frac-patient, fig.width=12, fig.height=8}
Patients = c("S0730", "S0819", "S0920", "S0619", "S1125", "S1204")
sapply(Patients, PlotFrac, seurat_obj = seurat, plot_what = "Patient", groupby = "cellType", color_maps = cell_color_maps, simplify = F) %>%
  wrap_plots() +
  plot_layout(ncol = 3, guides = 'collect')
```

### Source

```{r cell-num-Source-by-cluster}
DT::datatable(
  seurat@meta.data %>%
    group_by(Source, cellType) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

```{r frac-source, fig.height=18, fig.width=12}
sapply(Patients, PlotFracSource, seurat_obj = seurat, groupby = "cellType", color_maps = cell_color_maps, simplify = F) %>%
  wrap_plots(ncol = 2) +
  plot_layout(guides = 'collect')
```

## cellType by ... {.tabset}

```{r cell-num-source, fig.height=6, fig.width=7.5}
p_tissue = seurat@meta.data %>% 
  group_by(cellType, Tissue) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = Percent, fill = Tissue)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = tissue_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p_origin = seurat@meta.data %>% 
  group_by(cellType, Origin) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = Percent, fill = Origin)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = origin_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p_origin2 = seurat@meta.data %>% 
  group_by(cellType, Origin2_n) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = Percent, fill = Origin2_n)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = origin2_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p_origin3 = seurat@meta.data %>% 
  group_by(cellType, Origin3) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = Percent, fill = Origin3)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = origin3_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p_patient = seurat@meta.data %>% 
  group_by(cellType, Patient) %>% 
  summarise(n = n()) %>% 
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = Percent, fill = Patient)) + 
  geom_bar(stat = 'identity', color = 'white') +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = patient_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p1 = p_origin2 + p_patient + plot_layout(nrow = 1)
p2 = p_tissue + p_origin3 + plot_layout(nrow = 1)
p1/p2
```

### Tissue

```{r cell-num-source-tissue}
DT::datatable(
  seurat@meta.data %>%
    group_by(cellType, Tissue) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Origin2_n

```{r cell-num-source-Origin2_n}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Origin2_n) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Origin3

```{r cell-num-source-origin3}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Origin3) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Patient

```{r cell-num-source-patient}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Patient) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Origin

```{r cell-num-source-origin}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Origin) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

### Source

```{r cell-num-source-source}
DT::datatable(
  seurat@meta.data %>%
    group_by(seurat_clusters, Source) %>%
    summarise(n = n()) %>%
    mutate(Percent = n/sum(n) * 100),
  filter = 'top'
)
```

## N3-N5 {.tabset}



```{r n3-n5, fig.height=5, fig.width=7}
# overall
p1 = seurat@meta.data %>%
  filter(Tissue == 'Tumor') %>%
  group_by(Origin, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin, y = Percent, fill = cellType)) +
  geom_col() +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p2 = seurat@meta.data %>%
  filter(Tissue == 'Tumor') %>%
  group_by(Patient, Origin, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin, y = Percent, fill = cellType)) +
  geom_col() +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(.~Patient) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p1 + p2 +
  plot_layout(guides = "collect", widths = c(1, 3))
```

## LN Metastasis {.tabset}



### Overall

```{r LN-Metastasis, fig.width=4, fig.height=3}
# overall
p1 = seurat@meta.data %>%
  filter(Tissue == 'LN') %>%
  group_by(Metastasis_n, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = Percent, fill = Metastasis_n)) +
  geom_bar(stat="identity", position="dodge", width=0.8) +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = metastasis_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p1
```

### by Origin



```{r LN-Metastasis-from-same-Origin}
seurat@misc$sam_info %>%
  filter(Tissue == 'LN') %>%
  # 1 2 
  mutate(Origin = case_when(
    Origin %in% c("LN-LP1", "LN-LP2") ~ "LN-LP",
    TRUE ~ as.character(Origin)
  )) %>%
  # keep those with LN_N/LN_P
  filter(Origin %in% c("LN-LP", "LN-ITP", "LN-LG", "LN-LTP", "LN-RR")) %>%
  group_by(Origin, Metastasis_n) %>%
  summarise(n = n())
```



```{r LN-Metastasis-from-same-Origin-prop}
dat = seurat@meta.data %>%
  filter(Tissue == 'LN') %>%
  # 1 2 
  mutate(Origin = case_when(
    Origin %in% c("LN-LP1", "LN-LP2") ~ "LN-LP",
    TRUE ~ as.character(Origin)
  )) %>%
  # keep those with LN_N/LN_P
  filter(Origin %in% c("LN-LP", "LN-ITP", "LN-LG", "LN-LTP", "LN-RR")) %>%
  group_by(Origin, Metastasis_n, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100)
DT::datatable(dat,
              filter = "top")
```



```{r LN-Metastasis-from-same-Origin-p, fig.width=8, fig.height=4}
p2 = dat %>%  
  ggplot(aes(x = cellType, y = Percent, fill = Metastasis_n)) +
  #geom_col() +
  geom_bar(stat="identity", position="dodge", width=0.8) +
  labs(x = NULL, fill = NULL) +
  scale_fill_manual(values = metastasis_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_grid(.~Origin) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p2
```

### by Patient

 S0619  LN-LP1  LN-LP2 S1125  LN-RR  LN-LR 

```{r LN-Metastasis-from-same-Origin-by-Patient}
data.frame(
  Source = c("S0619.LN4", "S0619.LN5", "S1125.LN1", "S1125.LN6"),
  Patient = c("S0619", "S0619", "S1125", "S1125"),
  Tissue = c("LN", "LN", "LN", "LN"),
  Origin = c("LN-LP1", "LN-LP2", "LN-RR", "LN-LR"),
  Metastasis = c("N", "P", "N", "P"),
  Note = c("1", "2", "", "")
) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

```{r LN-Metastasis-from-same-Origin-by-Patient-p, fig.width=4, fig.height=4}
# S0619
p1 = seurat@meta.data %>%
  filter(Source %in% c('S0619.LN4', 'S0619.LN5')) %>%
  group_by(Origin, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin, y = Percent, fill = cellType)) +
  geom_col() +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, fill = NULL, title = 'S0619') +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

# S1125
p2 = seurat@meta.data %>%
  filter(Source %in% c('S1125.LN1', 'S1125.LN6')) %>%
  group_by(Origin, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Origin, y = Percent, fill = cellType)) +
  geom_col() +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, fill = NULL, title = 'S1125') +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

p1 + p2 +
  plot_layout(guides = "collect")
```

## Patient Metastasis





```{r metastatic-and-non-metastatic-tissue, fig.height=5, fig.width=7.5}
patient_info = tibble(Patient = c("S0619", "S0730", "S0819", "S0920", "S1125", "S1204"),
                      Metastatic = c('P', 'N', 'N', 'N', 'P', 'P'))

seurat@meta.data %>%
  left_join(patient_info, by = 'Patient') %>%
  filter(Origin2_n != "LN_P") %>%
  group_by(Origin2_n, Metastatic, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = Percent, fill = Metastatic)) +
  #geom_col() +
  geom_bar(stat="identity", position="dodge", width=0.8) +
  scale_fill_manual(values = metastasis_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, fill = NULL) +
  facet_wrap(.~Origin2_n) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## LN Drainage {.tabset}

### by Metastasis



```{r LN-drainage-by-Metastasis, fig.height=3, fig.width=7.5}
seurat@misc$sam_info %>%
  filter(Tissue == 'LN') %>%
  group_by(Metastasis_n, Drainage) %>%
  summarise(n = n())
```

Plot

```{r LN-drainage-by-Metastasis-p, fig.height=3, fig.width=6}
seurat@meta.data %>%
  filter(Tissue == 'LN') %>%
  group_by(Metastasis_n, Drainage, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = cellType, y = Percent, fill = Drainage)) +
  #geom_col() +
  geom_bar(stat="identity", position="dodge", width=0.8) +
  scale_fill_manual(values = drainage_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, fill = NULL) +
  facet_wrap(.~Metastasis_n) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### by Patient



```{r LN-drainage-by-Metastasis-by-patient, fig.height=3, fig.width=7.5}
seurat@misc$sam_info %>%
  filter(Tissue == 'LN') %>%
    filter(Patient %in% c("S0619", "S1125", "S1204")) %>%
  group_by(Patient, Metastasis_n, Drainage) %>%
  summarise(n = n())
```

plot

```{r LN-drainage-by-patient, fig.height=6, fig.width=8}
seurat@meta.data %>%
  filter(Tissue == 'LN') %>%
  group_by(Patient, Drainage, Metastasis_n, cellType) %>%
  summarise(n = n()) %>%
  mutate(Percent = n/sum(n) * 100) %>%
  ggplot(aes(x = Drainage, y = Percent, fill = cellType)) +
  geom_col() +
  scale_fill_manual(values = cell_color_maps) +
  scale_y_continuous(expand = c(0, 0)) +
  #geom_text_repel(aes(label = paste0(round(Percent, 1), "%")),
  #                  position = position_stack(vjust = 0.5)) +
  labs(x = NULL, fill = NULL) +
  facet_grid(Metastasis_n~Patient) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


## HC/Correlation {.tabset}

### Cell type

```{r hc-cellType-cor, cache=FALSE, fig.height=5, fig.width=5}
# Get average expression of variable features across cell types
Idents(seurat) <- 'cellType'
var_ave_byCell = AverageExpression(seurat, assays = "RNA", features = VariableFeatures(seurat), slot = "data")

var_ave_byCell.cor = cor(var_ave_byCell$RNA)
ggcorrplot(round(var_ave_byCell.cor, 2), hc.order = TRUE,
           type = "upper", lab = TRUE, ggtheme = cowplot::theme_cowplot)
```

```{r hc-cellType-heat, fig.width=3, fig.height=4, cache=FALSE}
pheatmap(var_ave_byCell$RNA, scale = "row", show_rownames = F, 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         treeheight_row = 1, treeheight_col = 10)
```

### Origin2_n

```{r hc-Origin2_n-cor, cache=FALSE, fig.height=5, fig.width=5}
# Get average expression of variable features across Origin2_n
# change identity to orig.ident
Idents(seurat) <- "Origin2_n"

# get average expression across original ident
var_ave_byOrigin2 = AverageExpression(seurat, assays = "RNA", features = VariableFeatures(seurat), slot = "data")

# restore cell identity
Idents(seurat) <- "cellType"

var_ave_byOrigin2.cor = cor(var_ave_byOrigin2$RNA)
ggcorrplot(round(var_ave_byOrigin2.cor, 2), hc.order=TRUE, type = "upper", lab = TRUE,
           ggtheme = cowplot::theme_cowplot)
```

```{r hc-Origin2_n-heat, fig.width=3, fig.height=4, cache=FALSE}
pheatmap(var_ave_byOrigin2$RNA, scale = "row", show_rownames = F,
         color = colorRampPalette(c("blue", "white", "red"))(50),
         treeheight_row = 1, treeheight_col = 10)
```

### Tissue

```{r hc-tissue-cor, cache=FALSE, fig.height=4, fig.width=4}
# Get average expression of variable features across tissue
# change identity to orig.ident
Idents(seurat) <- "Tissue"

# get average expression across original ident
var_ave_byTissue = AverageExpression(seurat, assays = "RNA", features = VariableFeatures(seurat), slot = "data")

# restore cell identity
Idents(seurat) <- "cellType"

var_ave_byTissue.cor = cor(var_ave_byTissue$RNA)
ggcorrplot(round(var_ave_byTissue.cor, 2), hc.order=TRUE, type = "upper", lab = TRUE,
           ggtheme = cowplot::theme_cowplot)
```

```{r hc-tissue-heat, fig.width=3, fig.height=4, cache=FALSE}
pheatmap(var_ave_byTissue$RNA, scale = "row", show_rownames = F,
         color = colorRampPalette(c("blue", "white", "red"))(50),
         treeheight_row = 1, treeheight_col = 10)
```

### Origin

```{r hc-Origin-cor, cache=FALSE, fig.width=8, fig.height=8}
# Get average expression of variable features across tissue
# change identity to orig.ident
Idents(seurat) <- "Origin"

# get average expression across original ident
var_ave_byOrigin = AverageExpression(seurat, assays = "RNA", features = VariableFeatures(seurat), slot = "data")

# restore cell identity
Idents(seurat) <- "cellType"

var_ave_byOrigin.cor = cor(var_ave_byOrigin$RNA)
ggcorrplot(round(var_ave_byOrigin.cor, 2), hc.order=TRUE, type = "upper", lab = TRUE,
           ggtheme = cowplot::theme_cowplot)
```

```{r hc-Origin-heat, fig.height=4, fig.width=4, cache=FALSE}
pheatmap(var_ave_byOrigin$RNA, scale = "row", show_rownames = F,
         color = colorRampPalette(c("blue", "white", "red"))(50),
         treeheight_row = 1, treeheight_col = 10)
```

### Source

```{r hc-source-cor, fig.height=8, fig.width=8, cache=FALSE}
# change identity to orig.ident
Idents(seurat) <- "Source"

# get average expression across original ident
var_ave_bySource = AverageExpression(seurat, assays = "RNA", features = VariableFeatures(seurat), slot = "data")

# restore cell identity
Idents(seurat) <- "cellType"

var_ave_bySource.cor = cor(var_ave_bySource$RNA)

# pheatmap
gg_color_hue <- function(n) {
  hues = seq(15, 375, length=n+1)
  grDevices::hcl(h=hues, l=65, c=100)[1:n]
}
# column annotation
col_ann = seurat@misc$sam_info %>%
  dplyr::select(Source, Patient, Tissue, Origin) %>%
  column_to_rownames(var = 'Source')
# color
ann_col = list()
for (col in colnames(col_ann)) {
  if (!is.factor(col_ann[,col])) col_ann[,col] <- factor(col_ann[,col])
  num <- length(levels(col_ann[,col]))
  colors <- 
    if (num <= 9) {
      RColorBrewer::brewer.pal(num, "Set1")
    } else {
      gg_color_hue(num)
    }
  ann_col[[col]] <- stats::setNames(colors, levels(col_ann[,col]))
}

pheatmap(var_ave_bySource.cor, scale = "none", display_numbers = F,
         color = colorRampPalette(c("blue", "white", "red"))(50), 
         show_rownames = F, show_colnames = F,
         annotation_col = col_ann, annotation_colors = ann_col,
         treeheight_row = 1, treeheight_col = 10)
```

```{r hc-source-heat, fig.height=8, fig.width=8, cache=FALSE}
pheatmap(var_ave_bySource$RNA, scale = "row", 
         show_rownames = F, show_colnames = F,
         annotation_col = col_ann, annotation_colors = ann_col, 
         color = colorRampPalette(c("blue", "white", "red"))(50),
         treeheight_col = 10, treeheight_row = 1)
```

## Figures

### Main

UMAP

```{r main-umap, fig.width=7.5, fig.height=2.4, cache=FALSE}
p.umap = p.umap.cellType +
  p.umap.vdj +
  p.umap.Origin2_n +
  #p.umap.Patient +
  plot_layout(nrow = 1)
p.umap
#ggsave2(p.umap, filename = here::here("output", DOCNAME, 'fig.main.umap.pdf'),
#       width = 8, height = 2.4, device = cairo_pdf)
ggsave2(p.umap, filename = here::here("output", DOCNAME, 'fig.main.umap.png'),
       width = 7.5, height = 2.4, type = "cairo")
```

Markers

```{r main-markers, fig.width=7.5, fig.height=1.3, cache=FALSE}
p.marker.main
ggsave2(p.marker.main, filename = here::here("output", DOCNAME, 'fig.main.markers.pdf'),
       width = 7.5, height = 1.5, device = cairo_pdf)
```

Sample composition

```{r main-composition, fig.width=7.5, fig.height=3, cache=FALSE}
p.cell_by_origin2
```

### Supp.

More markers

```{r supp-markers, fig.width=7.5, fig.height=5, cache=FALSE}
p.marker.supp
ggsave2(p.marker.supp, filename = here::here("output", DOCNAME, 'fig.supp.markers.pdf'),
        height = 5, width = 7.5, device = cairo_pdf)
```

Quality

```{r supp-quality, fig.width=7.5, fig.height=5, cache=FALSE}
p.quality
ggsave2(p.quality, filename = here::here("output", DOCNAME, 'fig.supp.quality.pdf'),
        height = 5, width = 7.5, device = cairo_pdf)
```

Composition

```{r supp-cluster-quality, fig.width=7.5, fig.height=2.5, cache=FALSE}
p.composition
ggsave2(p.composition, filename = here::here("output", DOCNAME, 'fig.supp.composition.pdf'),
        height = 2.5, width = 7.5, device = cairo_pdf)
```

## Summary

- We got 7 cell types

