---
title: "CytoTRACE.cd4"
---

```{r knitr, include = FALSE}
DOCNAME = "CytoTRACE.cd4"
#dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = TRUE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

Using [CytoTRACE](https://cytotrace.stanford.edu) to predict the differentiation state of cells. Higher CytoTRACE value implies higher developmental potential.

Paper of CytoTRACE: Gulati, G.S., Sikandar, S.S., Wesche, D.J., Manjunath, A., Bharadwaj, A., Berger, M.J., Ilagan, F., Kuo, A.H., Hsieh, R.W., Cai, S., et al. (2020). Single-cell transcriptional diversity is a hallmark of developmental potential. Science 367, 405–411.

## Setup

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(ggplotify)
library(ggcorrplot)
library(cowplot)
library(patchwork)
library(ggpubr)
theme_set(theme_cowplot())

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# Seurat
library(Seurat)

# CytoTRACE
library(CytoTRACE)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

## random1

Load rds

```{r load-rds-random1, cache=FALSE}
random1 = readRDS(here::here('output/04.rm_cells/seurat_cd4.random1.rds'))
Idents(random1) <- 'seurat_clusters'
random1
```

run CytoTRACE

```{r CytoTRACE-random1}
mat = GetAssayData(random1, assay = 'RNA', slot = 'counts')
mat = Matrix::as.matrix(mat)
result = CytoTRACE(mat, enableFast = F, ncores = 10)
```

Visualizing CytoTRACE results

```{r vis-CytoTRACE-no-run-random1, eval=FALSE, include=FALSE}
# output pdf directly, not run
emb = FetchData(random1, vars = c('UMAP_1', 'UMAP_2'))
phe = FetchData(random1, vars = c('seurat_clusters'))
phe = stats::setNames(as.character(phe[,1]), rownames(phe))
phe = paste('C_', phe, sep = '')

plotCytoTRACE(result, emb = emb, phenotype = phe)
plotCytoGenes(results, numOfGenes = 10)
```

```{r vis-CytoTRACE-random1, fig.height=3, fig.width=4}
meta = FetchData(random1, vars = c('UMAP_1', 'UMAP_2', 'seurat_clusters')) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```

```{r vis-box-CytoTRACE-random1, fig.height=3, fig.width=4}
meta = FetchData(random1, vars = c('UMAP_1', 'UMAP_2', 'seurat_clusters')) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=seurat_clusters, y=value, fill=seurat_clusters)) +
  geom_boxplot() +
  scale_fill_manual(values = cluster_color_maps) +
  labs(y= 'Predicted order') +
  theme(legend.position = 'none')
```

## random2

Load rds

```{r load-rds-random2, cache=FALSE}
random2 = readRDS(here::here('output/04.rm_cells/seurat_cd4.random2.rds'))
Idents(random2) <- 'seurat_clusters'
random2
```

run CytoTRACE

```{r CytoTRACE-random2}
mat = GetAssayData(random2, assay = 'RNA', slot = 'counts')
mat = Matrix::as.matrix(mat)
result = CytoTRACE(mat, enableFast = F, ncores = 10)
```

Visualizing CytoTRACE results

```{r vis-CytoTRACE-no-run-random2, eval=FALSE, include=FALSE}
# output pdf directly, not run
emb = FetchData(random2, vars = c('UMAP_1', 'UMAP_2'))
phe = FetchData(random2, vars = c('seurat_clusters'))
phe = stats::setNames(as.character(phe[,1]), rownames(phe))
phe = paste('C_', phe, sep = '')

plotCytoTRACE(result, emb = emb, phenotype = phe)
plotCytoGenes(results, numOfGenes = 10)
```

```{r vis-CytoTRACE-random2, fig.height=3, fig.width=4}
meta = FetchData(random2, vars = c('UMAP_1', 'UMAP_2', 'seurat_clusters')) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```

```{r vis-box-CytoTRACE-random2, fig.height=3, fig.width=4}
meta = FetchData(random2, vars = c('UMAP_1', 'UMAP_2', 'seurat_clusters')) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=seurat_clusters, y=value, fill=seurat_clusters)) +
  geom_boxplot() +
  scale_fill_manual(values = cluster_color_maps) +
  labs(y= 'Predicted order') +
  theme(legend.position = 'none')
```

## random3

Load rds

```{r load-rds-random3, cache=FALSE}
random3 = readRDS(here::here('output/04.rm_cells/seurat_cd4.random3.rds'))
Idents(random3) <- 'seurat_clusters'
random3
```

run CytoTRACE

```{r CytoTRACE-random3}
mat = GetAssayData(random3, assay = 'RNA', slot = 'counts')
mat = Matrix::as.matrix(mat)
result = CytoTRACE(mat, enableFast = F, ncores = 10)
```

Visualizing CytoTRACE results

```{r vis-CytoTRACE-no-run-random3, eval=FALSE, include=FALSE}
# output pdf directly, not run
emb = FetchData(random3, vars = c('UMAP_1', 'UMAP_2'))
phe = FetchData(random3, vars = c('seurat_clusters'))
phe = stats::setNames(as.character(phe[,1]), rownames(phe))
phe = paste('C_', phe, sep = '')

plotCytoTRACE(result, emb = emb, phenotype = phe)
plotCytoGenes(results, numOfGenes = 10)
```

```{r vis-CytoTRACE-random3, fig.height=3, fig.width=4}
meta = FetchData(random3, vars = c('UMAP_1', 'UMAP_2', 'seurat_clusters')) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point() +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```

```{r vis-box-CytoTRACE-random3, fig.height=3, fig.width=4}
meta = FetchData(random3, vars = c('UMAP_1', 'UMAP_2', 'seurat_clusters')) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=seurat_clusters, y=value, fill=seurat_clusters)) +
  geom_boxplot() +
  scale_fill_manual(values = cluster_color_maps) +
  labs(y= 'Predicted order') +
  theme(legend.position = 'none')
```



