---
title: "CytoTRACE.cd8"
---

```{r knitr, include = FALSE}
DOCNAME = "CytoTRACE.cd8"
#dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

Using [CytoTRACE](https://cytotrace.stanford.edu) to predict the differentiation state of cells. Higher CytoTRACE value implies higher developmental potential.

Paper of CytoTRACE: Gulati, G.S., Sikandar, S.S., Wesche, D.J., Manjunath, A., Bharadwaj, A., Berger, M.J., Ilagan, F., Kuo, A.H., Hsieh, R.W., Cai, S., et al. (2020). Single-cell transcriptional diversity is a hallmark of developmental potential. Science 367, 405–411.

Run `CytoTRACE.cd8.R` first.

## Setup

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(ggplotify)
library(ggcorrplot)
library(cowplot)
library(patchwork)
library(ggpubr)
theme_set(theme_cowplot(font_size = 10))

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# Seurat
library(Seurat)

# CytoTRACE
library(CytoTRACE)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

## random1

Load rds

```{r load-rds-random1, cache=FALSE}
srat = readRDS(here::here('output/04.rm_cells/seurat_cd8.random1.rds'))
Idents(srat) <- 'level_3'
srat
```

Load CytoTRACE results

```{r load-CytoTRACE-random1, cache=FALSE}
result = readRDS(here::here("output", DOCNAME, "CytoTRACE.random1.rds"))
```

Vis

```{r vis-random1, fig.height=3, fig.width=3}
meta = FetchData(srat, vars = c('UMAP_1', 'UMAP_2', 'level_3', "Patient")) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=reorder(level_3, -value, median), y=value, fill=level_3)) +
  geom_boxplot() +
  scale_fill_manual(values = t_level_3_color) +
  labs(y= 'Predicted order', x = NULL) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r vis-CytoTRACE, fig.height=3.5, fig.width=4}
tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point(size = .5) +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```

## random2

Load rds

```{r load-rds-random2, cache=FALSE}
srat = readRDS(here::here('output/04.rm_cells/seurat_cd8.random2.rds'))
Idents(srat) <- 'level_3'
srat
```

Load CytoTRACE results

```{r load-CytoTRACE-random2, cache=FALSE}
result = readRDS(here::here("output", DOCNAME, "CytoTRACE.random2.rds"))
```

Vis

```{r vis-random2, fig.height=3, fig.width=3}
meta = FetchData(srat, vars = c('UMAP_1', 'UMAP_2', 'level_3', "Patient")) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=reorder(level_3, -value, median), y=value, fill=level_3)) +
  geom_boxplot() +
  scale_fill_manual(values = t_level_3_color) +
  labs(y= 'Predicted order', x = NULL) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r vis-CytoTRACE-random2, fig.height=3.5, fig.width=4}
tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point(size = .5) +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```

## random3

Load rds

```{r load-rds-random3, cache=FALSE}
srat = readRDS(here::here('output/04.rm_cells/seurat_cd8.random3.rds'))
Idents(srat) <- 'level_3'
srat
```

Load CytoTRACE results

```{r load-CytoTRACE-random3, cache=FALSE}
result = readRDS(here::here("output", DOCNAME, "CytoTRACE.random3.rds"))
```

Vis

```{r vis-random3, fig.height=3, fig.width=3}
meta = FetchData(srat, vars = c('UMAP_1', 'UMAP_2', 'level_3', "Patient")) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=reorder(level_3, -value, median), y=value, fill=level_3)) +
  geom_boxplot() +
  scale_fill_manual(values = t_level_3_color) +
  labs(y= 'Predicted order', x = NULL) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r vis-CytoTRACE-random3, fig.height=3.5, fig.width=4}
tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point(size = .5) +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```

## random4

Load rds

```{r load-rds-random4, cache=FALSE}
srat = readRDS(here::here('output/04.rm_cells/seurat_cd8.random4.rds'))
Idents(srat) <- 'level_3'
srat
```

Load CytoTRACE results

```{r load-CytoTRACE-random4, cache=FALSE}
result = readRDS(here::here("output", DOCNAME, "CytoTRACE.random4.rds"))
```

Vis

```{r vis-random4, fig.height=3, fig.width=3}
meta = FetchData(srat, vars = c('UMAP_1', 'UMAP_2', 'level_3', "Patient")) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=reorder(level_3, -value, median), y=value, fill=level_3)) +
  geom_boxplot() +
  scale_fill_manual(values = t_level_3_color) +
  labs(y= 'Predicted order', x = NULL) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r vis-CytoTRACE-random4, fig.height=3.5, fig.width=4}
tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point(size = .5) +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```

## random5

Load rds

```{r load-rds-random5, cache=FALSE}
srat = readRDS(here::here('output/04.rm_cells/seurat_cd8.random5.rds'))
Idents(srat) <- 'level_3'
srat
```

Load CytoTRACE results

```{r load-CytoTRACE-random5, cache=FALSE}
result = readRDS(here::here("output", DOCNAME, "CytoTRACE.random5.rds"))
```

Vis

```{r vis-random5, fig.height=3, fig.width=3}
meta = FetchData(srat, vars = c('UMAP_1', 'UMAP_2', 'level_3', "Patient")) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=reorder(level_3, -value, median), y=value, fill=level_3)) +
  geom_boxplot() +
  scale_fill_manual(values = t_level_3_color) +
  labs(y= 'Predicted order', x = NULL) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r vis-CytoTRACE-random5, fig.height=3.5, fig.width=4}
tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point(size = .5) +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```

## random6

Load rds

```{r load-rds-random6, cache=FALSE}
srat = readRDS(here::here('output/04.rm_cells/seurat_cd8.random6.rds'))
Idents(srat) <- 'level_3'
srat
```

Load CytoTRACE results

```{r load-CytoTRACE-random6, cache=FALSE}
result = readRDS(here::here("output", DOCNAME, "CytoTRACE.random6.rds"))
```

Vis

```{r vis-random6, fig.height=3, fig.width=3}
meta = FetchData(srat, vars = c('UMAP_1', 'UMAP_2', 'level_3', "Patient")) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=reorder(level_3, -value, median), y=value, fill=level_3)) +
  geom_boxplot() +
  scale_fill_manual(values = t_level_3_color) +
  labs(y= 'Predicted order', x = NULL) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r vis-CytoTRACE-random6, fig.height=3.5, fig.width=4}
tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point(size = .5) +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```


## Session info

