---
title: "CytoTRACE.epithelia"
---

```{r knitr, include = FALSE}
DOCNAME = "CytoTRACE.epithelia"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

Using [CytoTRACE](https://cytotrace.stanford.edu) to predict the differentiation state of cells. Higher CytoTRACE value implies higher developmental potential.

> Paper of CytoTRACE: Gulati, G.S., Sikandar, S.S., Wesche, D.J., Manjunath, A., Bharadwaj, A., Berger, M.J., Ilagan, F., Kuo, A.H., Hsieh, R.W., Cai, S., et al. (2020). Single-cell transcriptional diversity is a hallmark of developmental potential. Science 367, 405–411.

## Setup

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(ggplotify)
library(ggcorrplot)
library(cowplot)
library(patchwork)
library(ggpubr)
theme_set(theme_cowplot())

# heatmap
library(pheatmap)
library(ComplexHeatmap)

# color
library(ggsci)

# Seurat
library(Seurat)

# CytoTRACE
library(CytoTRACE)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load seurat rds

```{r load-rds, cache=FALSE}
seurat = readRDS(here::here('output/04.rm_cells/seurat_epithelia.rds'))
Idents(seurat) <- 'seurat_clusters'
seurat
```

Load CytoTRACE results

```{r load-rds-CytoTRACE, cache=FALSE}
result = readRDS(here::here("output", DOCNAME, "CytoTRACE.res.rds"))
```

## CytoTRACE

Visualizing CytoTRACE results

```{r vis-CytoTRACE-no-run, eval=FALSE, include=FALSE}
# output pdf directly, not run
emb = FetchData(seurat, vars = c('UMAP_1', 'UMAP_2'))
phe = FetchData(seurat, vars = c('seurat_clusters'))
phe = stats::setNames(as.character(phe[,1]), rownames(phe))
phe = paste('C_', phe, sep = '')

plotCytoTRACE(result, emb = emb, phenotype = phe)
plotCytoGenes(results, numOfGenes = 10)
```

```{r vis-CytoTRACE, fig.height=7, fig.width=8}
meta = FetchData(seurat, vars = c('UMAP_1', 'UMAP_2', 'seurat_clusters', "Patient")) %>%
  rownames_to_column('barcode')

tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=UMAP_1, y=UMAP_2, color=value)) +
  geom_point(size = .5) +
  scale_color_viridis_c() +
  labs(color = 'Predicted\norder')
```

```{r vis-box-CytoTRACE, fig.height=4, fig.width=8}
tibble(barcode = names(result$CytoTRACE),
       value = result$CytoTRACE) %>%
  left_join(meta, by = 'barcode') %>%
  ggplot(aes(x=reorder(seurat_clusters, -value, median), y=value, fill=seurat_clusters)) +
  geom_boxplot() +
  scale_fill_manual(values = cluster_color_maps) +
  labs(y= 'Predicted order') +
  theme(legend.position = 'none')
```

## By patient

```{r fig.width=12, fig.height=9}
#p = 'S0619'
patients = unique(meta$Patient)
lapply(patients, function(p){
  res = readRDS(here::here("output", DOCNAME, paste0("CytoTRACE.res.", paste0(p, ".rds"))))
  tibble(barcode = names(res$CytoTRACE),
         value = res$CytoTRACE) %>%
    left_join(meta, by = 'barcode') %>%
    filter(Patient == p) %>%
    ggplot(aes(x=reorder(seurat_clusters, -value, median), y=value, fill=seurat_clusters)) +
    geom_boxplot() +
    scale_fill_manual(values = cluster_color_maps) +
    labs(y= 'Predicted order', title = p) +
    theme(legend.position = 'none')
}) %>%
  wrap_plots(ncol = 2)
```

## Session info

