---
title: "enrichment.ALL.cellType2"
---

```{r knitr, include = FALSE}
DOCNAME <- "enrichment.ALL.cellType2"

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = TRUE,
  cache.path     = paste0("cache/", DOCNAME, "/"),
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = TRUE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

Enrichment of all cellType2

## Setup

Load packages

```{r libaries, cache = FALSE}
# Tidyverse
library(tidyverse)

# Plotting
library(EnvStats)
library(ggcorrplot)
library(pheatmap)
library(ggpubr)
library(ggplotify)
library(ggrepel)

# patch
library(cowplot)
theme_set(theme_cowplot(
  font_size = 12,
  rel_small = 10 / 12,
  rel_tiny = 8 / 12,
  rel_large = 12 / 12,
  font_family = "Arial"
))
library(patchwork)

# color
library(ggsci)

# fonts
library(extrafont)
# font_import()
# font_import(paths = "/T02Data/niuyw/.local/share/fonts")
loadfonts(device = "pdf", quiet = TRUE)

# Seurat
library(Seurat)
```

```{r source, cache = FALSE}
source(here::here("code/preprocess.R"))
source(here::here("code/color.R"))
source(here::here("code/plot.R"))
```

Load metadata

```{r load-rds, cache=FALSE, message=FALSE, warning=FALSE}
metadata = read_csv(here::here("output/04.rm_cells/seurat.cellmeta.csv"))
sam_info = read_csv(here::here('data/sample_info.2.csv'))
```

## Plot

### Origin2_n (enrichment score + cell number) {.tabset}

#### All cells

With cell number (row order by number)

```{r fig.width=4, fig.height=8, warning=FALSE}
cluster_cellNum <- metadata %>%
  group_by(cellType2) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  group_by(cellType2, Origin2_n) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp <- setdiff(unique(metadata$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x <- tibble(cellType2 = unique(cluster_cellNum$cellType2), Origin2_n = t, o = 0)
    cluster_byOrigin2 <- bind_rows(cluster_byOrigin2, x)
  }
}

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin2_n") %>%
  left_join(cluster_cellNum, by = "cellType2") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType2, Origin2_n, r) %>%
  spread(Origin2_n, r, fill = 0) %>%
  column_to_rownames("cellType2")

# order
tmp.df = r_oe %>%
  rownames_to_column("cellType2") %>%
  arrange(PBMC1, PBMC2, LN_N, LN_P, Normal, Adjacent, Tumor)

# data for plot
d4p <- r_oe %>%
  rownames_to_column("cellType2") %>%
  pivot_longer(cols = !cellType2, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "PBMC1", "PBMC2",
      "LN_N", "LN_P",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType2 = factor(cellType2, levels = tmp.df$cellType2)
  )

# cell Num
df.cell_num <- cluster_byOrigin2

# plot
ggplot(d4p, aes(x = Origin2_n, y = cellType2, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.cell_num,
    aes(x = Origin2_n, y = cellType2, size = o),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "Cells") +
  theme_cowplot(font_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
```

With cell number (row order by hclust)

```{r fig.width=4, fig.height=8, warning=FALSE}
# order
hc <- hclust(dist(r_oe))
# data for plot
d4p <- r_oe %>%
  rownames_to_column("cellType2") %>%
  pivot_longer(cols = !cellType2, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "PBMC1", "PBMC2",
      "LN_N", "LN_P",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType2 = factor(cellType2, levels = rownames(r_oe)[hc$order])
  )

# cell Num
df.cell_num <- cluster_byOrigin2

# plot
ggplot(d4p, aes(x = Origin2_n, y = cellType2, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.cell_num,
    aes(x = Origin2_n, y = cellType2, size = o),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "Cells") +
  theme_cowplot(font_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
```

The original plot

```{r , fig.width=4, fig.height=8, cache=FALSE}
# use ggplot for better control
gg <- plot_origin2_enrich(r_oe)
gg
```

Plot with pheatmap

```{r fig.width=4, fig.height=8, warning=FALSE}
paletteLength <- 50
myColor <- colorRampPalette(c("blue", "white", "red"))(paletteLength)
# length(breaks) == length(paletteLength) + 1
# use floor and ceiling to deal with even/odd length pallettelengths
myBreaks <- c(
  seq(min(r_oe), 1, length.out = ceiling(paletteLength / 2) + 1),
  seq(max(r_oe) / paletteLength, max(r_oe), length.out = floor(paletteLength / 2))
)

pheatmap::pheatmap(r_oe,
  color = myColor, breaks = myBreaks,
  cluster_cols = FALSE, border_color = NA,
  fontsize = 8
)
```

#### Immune

With cell number (row order by number)

```{r fig.width=4, fig.height=6, warning=FALSE}
cluster_cellNum <- metadata %>%
  filter(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets")) %>%
  group_by(cellType2) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  filter(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets")) %>%
  group_by(cellType2, Origin2_n) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp <- setdiff(unique(metadata$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x <- tibble(cellType2 = unique(cluster_cellNum$cellType2), Origin2_n = t, o = 0)
    cluster_byOrigin2 <- bind_rows(cluster_byOrigin2, x)
  }
}

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin2_n") %>%
  left_join(cluster_cellNum, by = "cellType2") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType2, Origin2_n, r) %>%
  spread(Origin2_n, r, fill = 0) %>%
  column_to_rownames("cellType2")

# order
tmp.df = r_oe %>%
  rownames_to_column("cellType2") %>%
  arrange(PBMC1, PBMC2, LN_N, LN_P, Normal, Adjacent, Tumor)

# data for plot
d4p <- r_oe %>%
  rownames_to_column("cellType2") %>%
  pivot_longer(cols = !cellType2, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "PBMC1", "PBMC2",
      "LN_N", "LN_P",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType2 = factor(cellType2, levels = tmp.df$cellType2)
  )

# cell Num
df.cell_num <- cluster_byOrigin2

# plot
ggplot(d4p, aes(x = Origin2_n, y = cellType2, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.cell_num,
    aes(x = Origin2_n, y = cellType2, size = o),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "Cells") +
  theme_cowplot(font_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
```

With cell number (row order by hclust)

```{r fig.width=4, fig.height=6, warning=FALSE}
# order
hc <- hclust(dist(r_oe))
# data for plot
d4p <- r_oe %>%
  rownames_to_column("cellType2") %>%
  pivot_longer(cols = !cellType2, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "PBMC1", "PBMC2",
      "LN_N", "LN_P",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType2 = factor(cellType2, levels = rownames(r_oe)[hc$order])
  )

# cell Num
df.cell_num <- cluster_byOrigin2

# plot
ggplot(d4p, aes(x = Origin2_n, y = cellType2, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.cell_num,
    aes(x = Origin2_n, y = cellType2, size = o),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "Cells") +
  theme_cowplot(font_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
```

The original plot

```{r , fig.width=4, fig.height=6, cache=FALSE}
# use ggplot for better control
gg <- plot_origin2_enrich(r_oe)
gg
```

#### Non-immune

With cell number (row order by number)

```{r fig.width=4, fig.height=4, warning=FALSE}
cluster_cellNum <- metadata %>%
  filter(!(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets"))) %>%
  group_by(cellType2) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  filter(!(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets"))) %>%
  group_by(cellType2, Origin2_n) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp <- setdiff(unique(metadata$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x <- tibble(cellType2 = unique(cluster_cellNum$cellType2), Origin2_n = t, o = 0)
    cluster_byOrigin2 <- bind_rows(cluster_byOrigin2, x)
  }
}

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin2_n") %>%
  left_join(cluster_cellNum, by = "cellType2") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType2, Origin2_n, r) %>%
  spread(Origin2_n, r, fill = 0) %>%
  column_to_rownames("cellType2")

# order
tmp.df = r_oe %>%
  rownames_to_column("cellType2") %>%
  arrange(PBMC1, PBMC2, LN_N, LN_P, Normal, Adjacent, Tumor)

# data for plot
d4p <- r_oe %>%
  rownames_to_column("cellType2") %>%
  pivot_longer(cols = !cellType2, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "PBMC1", "PBMC2",
      "LN_N", "LN_P",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType2 = factor(cellType2, levels = tmp.df$cellType2)
  )

# cell Num
df.cell_num <- cluster_byOrigin2

# plot
ggplot(d4p, aes(x = Origin2_n, y = cellType2, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.cell_num,
    aes(x = Origin2_n, y = cellType2, size = o),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "Cells") +
  theme_cowplot(font_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
```

With cell number (row order by hclust)

```{r fig.width=4, fig.height=4, warning=FALSE}
# order
hc <- hclust(dist(r_oe))
# data for plot
d4p <- r_oe %>%
  rownames_to_column("cellType2") %>%
  pivot_longer(cols = !cellType2, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "PBMC1", "PBMC2",
      "LN_N", "LN_P",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType2 = factor(cellType2, levels = rownames(r_oe)[hc$order])
  )

# cell Num
df.cell_num <- cluster_byOrigin2

# plot
ggplot(d4p, aes(x = Origin2_n, y = cellType2, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.cell_num,
    aes(x = Origin2_n, y = cellType2, size = o),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "Cells") +
  theme_cowplot(font_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
```

The original plot

```{r , fig.width=4, fig.height=4, cache=FALSE}
# use ggplot for better control
gg <- plot_origin2_enrich(r_oe)
gg
```

### Origin2_n (enrichment score + % lineage cell) {.tabset}

#### All cells

With cell number (row order by number)

```{r fig.width=4, fig.height=8, warning=FALSE}
cluster_cellNum <- metadata %>%
  group_by(cellType2) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  group_by(cellType2, Origin2_n) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp <- setdiff(unique(metadata$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x <- tibble(cellType2 = unique(cluster_cellNum$cellType2), Origin2_n = t, o = 0)
    cluster_byOrigin2 <- bind_rows(cluster_byOrigin2, x)
  }
}

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin2_n") %>%
  left_join(cluster_cellNum, by = "cellType2") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType2, Origin2_n, r) %>%
  spread(Origin2_n, r, fill = 0) %>%
  column_to_rownames("cellType2")

# order
tmp.df = r_oe %>%
  rownames_to_column("cellType2") %>%
  arrange(PBMC1, PBMC2, LN_N, LN_P, Normal, Adjacent, Tumor)

# data for plot
d4p <- r_oe %>%
  rownames_to_column("cellType2") %>%
  pivot_longer(cols = !cellType2, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "PBMC1", "PBMC2",
      "LN_N", "LN_P",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType2 = factor(cellType2, levels = tmp.df$cellType2)
  )

# % lineage prop
a = metadata %>%
  group_by(cellType2, cellType, Origin2_n) %>%
  summarise(cellType2.cell_num = n()) %>%
  mutate(cellType2.cell_num = as.numeric(cellType2.cell_num))
b = metadata %>%
  group_by(cellType) %>%
  summarise(cellType.cell_num = n()) %>%
  mutate(cellType.cell_num = as.numeric(cellType.cell_num))
df.lineage_prop = a %>%
  left_join(b, by = "cellType") %>%
  mutate(lineage.prop = cellType2.cell_num*100/cellType.cell_num)

# plot
ggplot(d4p, aes(x = Origin2_n, y = cellType2, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.lineage_prop,
    aes(x = Origin2_n, y = cellType2, size = lineage.prop),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "% lineage") +
  theme_cowplot(font_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
```

#### Immune

With cell number (row order by number)

```{r fig.width=4, fig.height=6, warning=FALSE}
cluster_cellNum <- metadata %>%
  filter(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets")) %>%
  group_by(cellType2) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  filter(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets")) %>%
  group_by(cellType2, Origin2_n) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp <- setdiff(unique(metadata$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x <- tibble(cellType2 = unique(cluster_cellNum$cellType2), Origin2_n = t, o = 0)
    cluster_byOrigin2 <- bind_rows(cluster_byOrigin2, x)
  }
}

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin2_n") %>%
  left_join(cluster_cellNum, by = "cellType2") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType2, Origin2_n, r) %>%
  spread(Origin2_n, r, fill = 0) %>%
  column_to_rownames("cellType2")

# order
tmp.df = r_oe %>%
  rownames_to_column("cellType2") %>%
  arrange(PBMC1, PBMC2, LN_N, LN_P, Normal, Adjacent, Tumor)

# data for plot
d4p <- r_oe %>%
  rownames_to_column("cellType2") %>%
  pivot_longer(cols = !cellType2, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "PBMC1", "PBMC2",
      "LN_N", "LN_P",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType2 = factor(cellType2, levels = tmp.df$cellType2)
  )

# % lineage prop
a = metadata %>%
  filter(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets")) %>%
  group_by(cellType2, cellType, Origin2_n) %>%
  summarise(cellType2.cell_num = n()) %>%
  mutate(cellType2.cell_num = as.numeric(cellType2.cell_num))
b = metadata %>%
  filter(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets")) %>%
  group_by(cellType) %>%
  summarise(cellType.cell_num = n()) %>%
  mutate(cellType.cell_num = as.numeric(cellType.cell_num))
df.lineage_prop = a %>%
  left_join(b, by = "cellType") %>%
  mutate(lineage.prop = cellType2.cell_num*100/cellType.cell_num)

# plot
ggplot(d4p, aes(x = Origin2_n, y = cellType2, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.lineage_prop,
    aes(x = Origin2_n, y = cellType2, size = lineage.prop),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "% lineage") +
  theme_cowplot(font_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
```

#### Non-immune

With cell number (row order by number)

```{r fig.width=4, fig.height=4, warning=FALSE}
cluster_cellNum <- metadata %>%
  filter(!(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets"))) %>%
  group_by(cellType2) %>%
  summarise(cluster.cell_num = n()) %>%
  mutate(cluster.cell_num = as.numeric(cluster.cell_num))

origin2_CellNum <- sam_info %>%
  group_by(Origin2_n) %>%
  summarise(origin.cell_num = sum(cell.filtered)) %>%
  mutate(origin.cell_num = as.numeric(origin.cell_num)) %>%
  mutate(total = sum(origin.cell_num)) %>%
  mutate(total = as.numeric(total))

cluster_byOrigin2 <- metadata %>%
  filter(!(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets"))) %>%
  group_by(cellType2, Origin2_n) %>%
  summarise(o = n()) %>%
  mutate(o = as.numeric(o))

# add origins not existing
tmp <- setdiff(unique(metadata$Origin2_n), cluster_byOrigin2$Origin2_n)
if (length(tmp) > 0) {
  for (t in tmp) {
    x <- tibble(cellType2 = unique(cluster_cellNum$cellType2), Origin2_n = t, o = 0)
    cluster_byOrigin2 <- bind_rows(cluster_byOrigin2, x)
  }
}

r_oe <- cluster_byOrigin2 %>%
  left_join(origin2_CellNum, by = "Origin2_n") %>%
  left_join(cluster_cellNum, by = "cellType2") %>%
  mutate(e = (origin.cell_num * cluster.cell_num) / total) %>%
  mutate(r = o / e) %>%
  dplyr::select(cellType2, Origin2_n, r) %>%
  spread(Origin2_n, r, fill = 0) %>%
  column_to_rownames("cellType2")

# order
tmp.df = r_oe %>%
  rownames_to_column("cellType2") %>%
  arrange(PBMC1, PBMC2, LN_N, LN_P, Normal, Adjacent, Tumor)

# data for plot
d4p <- r_oe %>%
  rownames_to_column("cellType2") %>%
  pivot_longer(cols = !cellType2, names_to = "Origin2_n") %>%
  mutate(
    Origin2_n = factor(Origin2_n, levels = c(
      "PBMC1", "PBMC2",
      "LN_N", "LN_P",
      "Normal", "Adjacent",
      "Tumor"
    )),
    cellType2 = factor(cellType2, levels = tmp.df$cellType2)
  )

# % lineage prop
a = metadata %>%
  filter(!(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets"))) %>%
  group_by(cellType2, cellType, Origin2_n) %>%
  summarise(cellType2.cell_num = n()) %>%
  mutate(cellType2.cell_num = as.numeric(cellType2.cell_num))
b = metadata %>%
  filter(!(cellType %in% c("T cells", "B cells", "Myeloid", "Platelets"))) %>%
  group_by(cellType) %>%
  summarise(cellType.cell_num = n()) %>%
  mutate(cellType.cell_num = as.numeric(cellType.cell_num))
df.lineage_prop = a %>%
  left_join(b, by = "cellType") %>%
  mutate(lineage.prop = cellType2.cell_num*100/cellType.cell_num)

# plot
ggplot(d4p, aes(x = Origin2_n, y = cellType2, fill = value)) +
  geom_tile() +
  geom_point(
    data = df.lineage_prop,
    aes(x = Origin2_n, y = cellType2, size = lineage.prop),
    shape = 1,
    inherit.aes = FALSE
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(
    midpoint = 1,
    low = "blue", mid = "white", high = "red",
    name = expression(R[O / E])
  ) +
  scale_size(range = c(1, 3)) +
  labs(size = "% lineage") +
  theme_cowplot(font_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 0),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    text = element_text(size = 6)
  )
```

## Session

